---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Build network graph for advanced R solutions book cover
Dec 2020
Authors: Henning Bumann, Tobias Stalder

output: edges_df, nodes_df
options:
 - Filter options for "depth" of the network (e.g. hierarchy level)
 - volles Netzwerk vs. selection
 - size (N Aufgaben) und type für jedes Zeile in node_df berechnen

## set working dir and load packages ---------------------------------------
```{r}
packages <- c("tidyverse", "Cairo", "ggpubr",
              # "ggalt",  # läuft bei mir nicht
              "ggbump", "scales", "prismatic", "treemapify","GGally",
              "intergraph", "ggraph", "visNetwork", "threejs", "networkD3",
              #"ndtv",
              "igraph", "gtools")

package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```

## load and clean data to long format
```{r}
df_long_raw <- read.csv2(r"(progress_data.csv)", stringsAsFactors = FALSE, 
                     header = TRUE) %>% 
  as_tibble() %>% 
  mutate(status = "finalised") %>% 
  modify_at("status", factor,
            levels = c("open", "started", "solved", "reviewed", "finalised")) %>% 
  tibble::rowid_to_column("id") %>% 
  separate(chapter, "nr", sep = " ", remove = FALSE, extra = "drop") %>% 
  mutate(nr = as.integer(nr), 
         chapter = factor(chapter),
         chapter = fct_reorder(chapter, nr, .desc = FALSE),
         part = case_when(nr < 9 ~ "Foundations",
                          nr < 12 ~ "Functional\nProgramming",
                          nr < 17 ~ "Object Oriented\nProgramming",
                          nr < 22 ~ "Metaprogramming",
                          TRUE    ~ "Techniques"),
         part = factor(part, levels = c("Foundations", "Functional\nProgramming",
                                        "Object Oriented\nProgramming",
                                        "Metaprogramming", "Techniques"))) %>% 
  add_count(part, name = "n_part") %>% 
  add_count(chapter, name = "n_chapter") %>% 
  add_count(subchapter, name = "n_subchapter")


df_long <- df_long_raw %>% 
  mutate(part = paste0("pa_", part),
         chapter = paste0("ch_", chapter),
         subchapter = paste0("sc_", subchapter),
         exercise = paste0("ex_", id)) %>% 
  select(id, part, chapter, subchapter, exercise, starts_with("n_"))
```

## create nodes_df
```{r}
nodes_df <- bind_rows(
  distinct(df_long, node = part, size = n_part),
  distinct(df_long, node = chapter, size = n_chapter),
  distinct(df_long, node = subchapter, size = n_subchapter),
  distinct(df_long, node = exercise, size = 1)
) %>% 
  mutate(nodetype = case_when(node %>% str_starts("pa_") ~ "part",
                              node %>% str_starts("ch_") ~ "chapter",
                              node %>% str_starts("sc_") ~ "subchapter",
                              node %>% str_starts("ex_") ~ "exercise"))

# mögliche subsetbildung auf diese Weise :)
nodes_wo_ex_df <- nodes_df %>% 
  filter(!nodetype %in% c("exercise"))

## alternative code snippet
# nodes_df <- df_long %>% 
#   select(part, chapter, subchapter, exercise) %>% 
#   modify(as.character) %>% 
#   map(unique) %>% 
#   reduce(c) %>% 
#   tibble(node = .) 
```


## create edges_df
```{r}
edges_full_df <- tidyr::crossing(from = nodes_df$node,
                                 to = nodes_df$node) %>% 
  mutate(edgetype = paste0(str_sub(from, end = 3), str_sub(to, end = 2))) %>% 
  filter(from != to,
         edgetype %in% c("pa_pa", "ch_ch", "sc_sc", "ex_ex"))

# TODO: connect with everything in one hierarchy and with the parent
edges_full_df %>% 
  count(edgetype)

# - kick out all connections from ex_ to ch_ / pa_
# - kick out all edges from sc_ to pa_


# TODO: define interconnections between exercises of a subchapter
df_long %>% 
  filter(subchapter == "sc_Binding basics") %>% 
  pull(exercise) %>% 
  expand_grid(from = ., to = .) %>% 
  filter(from != to)


# do this grouped (via function?)


# define interconnections between subchapters of a chapter


# define interconnections between chapters of a part



## define valid cross hierarchy edges
# find a way to only include edges from ex_ to matching sc_
ex_sc <- df_long %>% 
  distinct(exercise, subchapter) %>% 
  tidyr::unite(ex_sc, exercise, subchapter)

# find way to only include sc to matching c
sc_c <- df_long %>% 
  distinct(subchapter, chapter) %>% 
  tidyr::unite(sc_c, subchapter, chapter)

# find way to only include c to matching p
c_p <- df_long %>% 
  distinct(chapter, part) %>% 
  tidyr::unite(c_p, chapter, part)

# combine
include_cross_hierarchy <- bind_rows(
  ex_sc,
  sc_c,
  c_p
)



# finally do the filtering...
edges_df <- edges_full_df %>% 
  tidyr::unite(from_to, from, to, remove = FALSE) %>% 
  select(from, to, from_to) %>% 
  semi_join(include_cross_hierarchy, by = c("from_to" = "ex_sc"))


# remove "mirrored connections"?




## ALTERNATIVE: explicitly create all the connections we want



# TODO: connect with predecessor and descendant (to reduce number of connections)
```


## test network
```{r}
net_ex <- graph_from_data_frame(d = edges_df, vertices = nodes_df)

plot(net_ex)

ggraph(net_ex, layout = "kk") +
  geom_edge_link(color = "grey") +
  geom_node_point(size = 3) +
  geom_node_text(aes(label = name), color = "blue", nudge_y = .2)

```