# Data structures {#intro}

## Vectors

1. __<span style="color:red">Q</span>__: What are the six types of atomic vector? How does a list differ from an
   atomic vector?  
__<span style="color:green">A</span>__:The six types are logical, integer, double, character, complex and raw. The elements of a list
don't have to be of the same type.

2. __<span style="color:red">Q</span>__: What makes `is.vector()` and `is.numeric()` fundamentally different to
   `is.list()` and `is.character()`?  
__<span style="color:green">A</span>__: The first two tests don't check for a specific type.

3. __<span style="color:red">Q</span>__: Test your knowledge of vector coercion rules by predicting the output of
   the following uses of `c()`:

    ```{r, eval=FALSE}
    c(1, FALSE)      # will be coerced to numeric   -> 1 0
    c("a", 1)        # will be coerced to character -> "a" "1"
    c(list(1), "a")  # will be coerced to a list of lists (types don't change)
    c(TRUE, 1L)      # will be coerced to integer   -> 1 1
    ```

4.  __<span style="color:red">Q</span>__: Why do you need to use `unlist()` to convert a list to an 
    atomic vector? Why doesn't `as.vector()` work?  
__<span style="color:green">A</span>__: To get rid of (flatten) the nested structure.

5. __<span style="color:red">Q</span>__: Why is `1 == "1"` true? Why is `-1 < FALSE` true? Why is `"one" < 2` false?  
__<span style="color:green">A</span>__: These operators are all functions which coerce their arguments (in these cases) to character, double and character. To enlighten the latter case: "one" is behind "2" in ASCI.

6. __<span style="color:red">Q</span>__: Why is the default missing value, `NA`, a logical vector? What's special
   about logical vectors? (Hint: think about `c(FALSE, NA_character_)`.)  
__<span style="color:green">A</span>__: It is a practical concern. When you combine NAs in `c()` with other atomic types it will be coerced
like TRUE and FALSE to integer `(NA_integer)`, double `(NA_real)`, complex `(NA_complex)` and character `(NA_character)`.

## Attributes

1.  __<span style="color:red">Q</span>__: An early draft used this code to illustrate `structure()`:

    ```{r}
    structure(1:5, comment = "my attribute")
    ```

    But when you print that object you don't see the comment attribute.
    Why? Is the attribute missing, or is there something else special about
    it? (Hint: try using help.) \index{attributes!comment}
    
    __<span style="color:green">A</span>__: 

2.  __<span style="color:red">Q</span>__: What happens to a factor when you modify its levels? 
    
    ```{r, results = "none"}
    f1 <- factor(letters)
    levels(f1) <- rev(levels(f1))
    ```
    
    __<span style="color:green">A</span>__: 

3.  __<span style="color:red">Q</span>__: What does this code do? How do `f2` and `f3` differ from `f1`?

    ```{r, results = "none"}
    f2 <- rev(factor(letters))

    f3 <- factor(letters, levels = rev(letters))
    ```
    
    __<span style="color:green">A</span>__: 

## Matrices and arrays

1.  __<span style="color:red">Q</span>__: What does `dim()` return when applied to a vector?  
__<span style="color:green">A</span>__: 

2.  __<span style="color:red">Q</span>__: If `is.matrix(x)` is `TRUE`, what will `is.array(x)` return?  
__<span style="color:green">A</span>__: 

3.  __<span style="color:red">Q</span>__: How would you describe the following three objects? What makes them
    different to `1:5`?

    ```{r}
    x1 <- array(1:5, c(1, 1, 5))
    x2 <- array(1:5, c(1, 5, 1))
    x3 <- array(1:5, c(5, 1, 1))
    ```  
    
    __<span style="color:green">A</span>__: 

## Data frames

1.  __<span style="color:red">Q</span>__: What attributes does a data frame possess?  
__<span style="color:green">A</span>__: 

2.  __<span style="color:red">Q</span>__: What does `as.matrix()` do when applied to a data frame with 
    columns of different types?  
    __<span style="color:green">A</span>__: 

3.  __<span style="color:red">Q</span>__: Can you have a data frame with 0 rows? What about 0 columns?  
__<span style="color:green">A</span>__: 
