# Functions

## Function components

1.  __<span style="color:red">Q</span>__: What function allows you to tell if an object is a function? What function
    allows you to tell if a function is a primitive function?  
    __<span style="color:green">A</span>__: You can test objects with `is.function()` and `is.primitive()`.

2.  __<span style="color:red">Q</span>__: This code makes a list of all functions in the base package. 
    
    ```{r}
    objs <- mget(ls("package:base"), inherits = TRUE)
    funs <- Filter(is.function, objs)
    ```
    
    Use it to answer the following questions:
    
    a. Which base function has the most arguments?
    
    a. How many base functions have no arguments? What's special about those
       functions?
       
    a. How could you adapt the code to find all primitive functions?  
    
    __<span style="color:green">A</span>__: 
    
    a.  First we create a named vector that returns the number of arguments per function and then we subset it with the index of it's maximum entry:
        
    ```{r}
    f_arg_length <- sapply(funs, function(x) length(formals(x)))
    f_arg_length[which.max(f_arg_length)]
    ```
    
    b. We check the number of functions with `formals()` returning `0` or `NULL`.
    Then we will see, that all of these functions have `formals` equal to `NULL`,
    which means, that they should be primitive functions.
    
    ```{r}
    sum(sapply(funs, function(x) is.null(formals(x)) | length(formals(x)) == 0))
    sum(sapply(funs, function(x) !is.null(formals(x)) & length(formals(x)) == 0))
    sum(sapply(funs, function(x) is.null(formals(x))))
    sum(sapply(funs, function(x) is.null(formals(x)) & is.primitive(x)))
    ```
    
    Hence not all functions with `formals` equal to `NULL` are primitive functions, there must be non primitive functions with this property too.  
    c. Change the predicate in `Filter` to `is.primitive`:
    
    ```{r, eval = FALSE}
    funs <- Filter(is.primitive, objs)
    ```  
    
3. __<span style="color:red">Q</span>__: What are the three important components of a function?  
__<span style="color:green">A</span>__: `body()`, `formals()` and `environment()`.
    
    > There is one exception to the rule that functions have three components. Primitive functions, like `sum()`, call C code directly with `.Primitive()` and contain no R code. Therefore their `formals()`, `body()`, and `environment()` are all `NULL`.

4. __<span style="color:red">Q</span>__: When does printing a function not show what environment it was created in?  
__<span style="color:green">A</span>__: When the function is a primitive or it was created in the global environment.

## Lexical Scoping

1. __<span style="color:red">Q</span>__: What does the following code return? Why? What does each of the three `c`'s mean?

    ```{r, eval = FALSE}
    c <- 10
    c(c = c)
    ```  
    
    __<span style="color:green">A</span>__: A named vector `c`, which first field has the value 10 and the name "c". The first "c" is the `c()` function, the second is the name of the first entry and the third is the value of the first entry.
       
2. __<span style="color:red">Q</span>__: What are the four principles that govern how R looks for values?  
__<span style="color:green">A</span>__:   As stated in the book:
    
    > There are four basic principles behind R's implementation of lexical scoping:
    * name masking (Variables are evaluated according to the highest-precedence environment in which they are defined, starting from the local environment and working upwards through each parent environment.)
    * functions vs. variables (For all intents and purposes, function names are evaluated by the same rules as for variables.  If it is implicit that a function is being used, R will ignore objects with the same name that are not functions.)
    * a fresh start (Every time a function is called, a new environment is created to host execution.)
    * dynamic lookup (R looks for values, when a function is run. Not when it is created)

3. __<span style="color:red">Q</span>__: What does the following function return? Make a prediction before 
   running the code yourself.

    ```{r, eval = FALSE}
    f <- function(x) {
      f <- function(x) {
        f <- function(x) {
          x ^ 2
        }
        f(x) + 1
      }
      f(x) * 2
    }
    f(10)
    ```
      
      __<span style="color:green">A</span>__: 202
      
## Function arguments

1.  __<span style="color:red">Q</span>__: Clarify the following list of odd function calls:

    ```{r, eval = FALSE}
    x <- sample(replace = TRUE, 20, x = c(1:10, NA))
    # -> sample(x = c(1:10, NA), size = 20, replace = TRUE)
    y <- runif(min = 0, max = 1, 20)
    # -> runif(n = 20, min = 0, max = 1)
    cor(m = "k", y = y, u = "p", x = x)
    # -> cor(x = x, y = y, use = "pairwise.complete.obs", method = "kendall")
    ```  

2.  __<span style="color:red">Q</span>__: What does this function return? Why? Which principle does it illustrate?
  
    ```{r, eval = FALSE}    
    f1 <- function(x = {y <- 1; 2}, y = 0) {
      x + y
    }
    f1()
    ```  
    
    __<span style="color:green">A</span>__: It returns 3 and illustrates lazy evaluation. As you can see, y becomes 1, but only when `x` is evaluated (before `y`) inside the function (otherwise it is 0):
    
    ```{r}    
    f2 <- function(x = {y <- 1; 2}, y = 0) {
      y
    }
    f2()
    ```  
    
    Note that funny things can happen if we switch the evaluation order (even within one line)
    
    ```{r}    
    f3 <- function(x = {y <- 1; 2}, y = 0) {
      y + x
    }
    f3()
    ```


    or we evaluate `y` once before and once after the evaluation of `x`
    
    ```{r}    
    f4 <- function(x = {y <- 1; 2}, y = 0) {
      y_before_x <- y
      x
      y_after_x <- y
      c(y_before_x, y_after_x)
    }
    f4()
    ```

3.  __<span style="color:red">Q</span>__: What does this function return? Why? Which principle does it illustrate?

    ```{r, eval = FALSE}
    f2 <- function(x = z) {
      z <- 100
      x
    }
    f2()
    ```  
    
    __<span style="color:green">A</span>__: 100, lazy evaluation. If we changed the order
    of `z <- 100` and `x` inside the body of the function, we would get an error, while assigning `z` to `x`, because `z` wouldn`t be defined.
    
## Special calls

1. __<span style="color:red">Q</span>__: Create a list of all the replacement functions found in the base package. 
   Which ones are primitive functions?  
   __<span style="color:green">A</span>__: We can find replacementfunctions by searching for functions that end on "<-":
   
    ```{r, eval = TRUE}
    repls <- funs[grepl("<-$", names(funs))]
    names(repls)
    names(Filter(is.primitive, repls))
    ```

2. __<span style="color:red">Q</span>__: What are valid names for user-created infix functions?  
  
    __<span style="color:green">A</span>__: From the textbook:
    
    > All user-created infix functions must start and end with `%` ... they can contain any sequence of characters (except “%”, of course).

3. __<span style="color:red">Q</span>__: Create an infix `xor()` operator.  
__<span style="color:green">A</span>__: 

    ```{r}
    `%xor_%` <- function(a, b){
      (a | b) & !(a & b)
    }
    ```

4. __<span style="color:red">Q</span>__: Create infix versions of the set functions `intersect()`, `union()`, and 
   `setdiff()`.  
   __<span style="color:green">A</span>__: 
   
    ```{r}
    `%intersect_%` <- function(a, b){
      unique(a[a %in% b])
      }
    
    `%union_%` <- function(a, b){
      unique(c(a, b))
      }

    `%setdiff_%` <- function(a, b){
      unique(a[! a %in% b])
      }
    ```
   
5. __<span style="color:red">Q</span>__: Create a replacement function that modifies a random location in a vector.  
__<span style="color:green">A</span>__: 

    ```{r, eval = FALSE}
    `random_replacement<-` <- function(x, value){
      x[sample(length(x), 1)] <- value
      x
      }
    ```

## Return Values

1.  __<span style="color:red">Q</span>__: How does the `chdir` parameter of `source()` compare to `in_dir()`? Why 
    might you prefer one approach to the other?
    The `in_dir()` approach was given in the book as
    
    ```{r, eval = FALSE}
    in_dir <- function(dir, code) {
      old <- setwd(dir)
      on.exit(setwd(old))
      
      force(code)
      }
    ```
    
    __<span style="color:green">A</span>__: `in_dir()` takes a path to a working directory as an argument. At the beginning of the function the working directory is changed to this specification and with a call to `on.exit` it is guranteed, that when the function finishes the working directory also equals to this specification.
    
    In `source()` you need the `chdir` argument to specify, if the working directory should be changed during the evaluation to the `file` argument, if this is a pathname. The difference in `source()` is, that the actual working directory as output of `getwd()` is saved to set it in `on.exit` before changing the directory to the pathname (given to the `file` argument) for the rest of the execution of the `source()` function.
    
    The methodology used in `in_dir()` is more free, allowing more programmer customization, whereas the methodology in `chdir()` is more constrained, reducing programmer error. This is generally a trade-off that different people will want to take different sides of, usually depending on how much they know what they are doing.

2.  __<span style="color:red">Q</span>__: What function undoes the action of `library()`? How do you save and restore
    the values of `options()` and `par()`?  
    __<span style="color:green">A</span>__: Use `detach()` with `"package:packagename"` as first argument.  
    Type `options()` (or `par()`) to see the actual settings and save them via `old_opt <- options()[["optionname"]]` (or `old_par <- par()[["parname"]]`). For options you could also use `getOption()` with `"optionname"` as argument. To restore (or overwriting) just use `options(optionname = old_opt)` or `par(parname = old_par)`.

3.  __<span style="color:red">Q</span>__: Write a function that opens a graphics device, runs the supplied code, and 
    closes the graphics device (always, regardless of whether or not the plotting code worked).  
    __<span style="color:green">A</span>__: 
    
    ```{r, eval = FALSE}
    plot_pdf <- function(code){
      pdf("test.pdf")
      on.exit(dev.off())
      code
      }
    ```

4.  __<span style="color:red">Q</span>__: We can use `on.exit()` to implement a simple version of `capture.output()`.

    ```{r, warning = FALSE}
    capture.output2 <- function(code) {
      temp <- tempfile()
      on.exit(file.remove(temp), add = TRUE)

      sink(temp)
      on.exit(sink(), add = TRUE)

      force(code)
      readLines(temp)
    }
    capture.output2(cat("a", "b", "c", sep = "\n"))
    ```
    
    Compare `capture.output()` to `capture.output2()`. How do the functions differ? What features have I removed to make the key ideas easier to see? How have I rewritten the key ideas to be easier to understand?  
    __<span style="color:green">A</span>__: Using `body(capture.output)`, we can see the source code for the original `capture.output()` function. `capture.output()` is a good clip longer (39 lines vs. 7 lines). The reason for this is that `capture.output2()` is more modular, since `capture.output()` writes out entire methods like `readLines()` instead of invoking them. This makes `capture.output2` easier to understand if you understand the underlying methods.
    
    However, `capture.output2()` does remove potentially important functionality, as `capture.output()` appears to handle important exceptions not handled in `capture.output2()`, and `capture.output()` offers the ability to chose between overwriting or appending to a file.