# Environments

## Environment basics

1.  __<span style="color:red">Q</span>__: List three ways in which an environment differs from a list.  
__<span style="color:green">A</span>__: The most important differences are:
    * environments have reference semantics
    * environments have parents
    * environments are not ordered
    * elements of environments need to be (uniquely) named

2.  __<span style="color:red">Q</span>__: If you don't supply an explicit environment, where do `ls()` and `rm()`
    look? Where does `<-` make bindings? The  
    __<span style="color:green">A</span>__: `ls()` and `rm` look in their calling environments which they find by `as.environment(-1)`.  
    From the book:
    
    > Assignment is the act of binding (or rebinding) a name to a value in an environment.  
    
    From `` ?`<-` ``:
    
    > The operators `<-` and `=` assign into the environment in which they are evaluated. The operator `<-` can be used anywhere, whereas the operator `=` is only allowed at the top level (e.g., in the complete expression typed at the command prompt) or as one of the subexpressions in a braced list of expressions.

3.  __<span style="color:red">Q</span>__: Using `parent.env()` and a loop (or a recursive function), verify that the 
    ancestors of `globalenv()` include `baseenv()` and `emptyenv()`. Use the 
    same basic idea to implement your own version of `search()`.  
    __<span style="color:green">A</span>__:  We can print the ancestors for example by using a recursive function:  
    
    ```{r, eval = TRUE}
    ancestors <- function(env = globalenv()){
      if (identical(env, emptyenv())) {
        print(environmentName(env))} 
      else {
        print(environmentName(env))
        ancestors(parent.env(env))
      }
    }
    
    ancestors()
    ```
    
    To implement a new version of `search()` we use a while statement:
    
    ```{r, eval = TRUE}
    search2 <- function(env = globalenv()){
      envs <- character()
      while (!identical(env, emptyenv())) {
        ename <- environmentName(env)
        if (ename == "base") ename <- "package:base"
        if (ename == "R_GlobalEnv") ename <- ".GlobalEnv"
        envs <- c(envs, ename)
        env <- parent.env(env)
        }
      return(envs)
    }
    
    search2()
    # visual check that results are identical to the original search() function
    search()
    ```
    
## Recursing over environments

1.  __<span style="color:red">Q</span>__: Modify `where()` to find all environments that contain a binding for
    `name`.  
    __<span style="color:green">A</span>__: We look at the source code of the original `pryr::where()`:  
    
    ```{r, eval = FALSE}
    pryr::where
    function (name, env = parent.frame()) 
      {
      stopifnot(is.character(name), length(name) == 1)
      env <- to_env(env)
      if (identical(env, emptyenv())) { # "base case"
        stop("Can't find ", name, call. = FALSE)
        }
      if (exists(name, env, inherits = FALSE)) { # "success case"
        env
        }
      else { # "recursive case"
        where(name, parent.env(env)) # we will copy this line in the success case
      }
    }
    ```
    
    Since `where()` stops searching when a match appears, we copy the recursive call in the `else` block to the block of the matching ("success") case, so that our new function `where2` will look for a binding within the complete search path. We also need to pay attention to other details. We have to take care to save the bindings in an object, while not overriding it in our recursive calls. So we create a list object for that and define a new function within `where2()` that we call `where2.internal`. `where2.internal()` will do the recursive work and whenever it finds a binding it will write it via `<<-` to the especially created list in its enclosing environment:

    ```{r, eval = FALSE}
    where2 <- function(name, env = parent.frame()){
      # we need to collect all environments where name has a binding
      env_list <- list()
      
      # since our function will be recursive and env_list would be overwritten
      # when it is inside the recursive function, we put it on the outside of
      # the recursive function and concatenate every binding environment
      # that we find via the `<<-` operator on its end.
      # In the following we start by defining the recursive function:
      where2.internal <- function(name, env = parent.frame()) {
        stopifnot(is.character(name), length(name) == 1)
        env <- pryr:::to_env(env) # note that we need to call to_env via pryr:::
        
        # when we reach the empty environment, we return all binding environments
        # (if we found some) if we found no bindings, we give the same error message
        # as pryr::where does
        if (identical(env, emptyenv())) {
          if (length(env_list) != 0){
            return(env_list)
            }
          stop("Can't find ", name, call. = FALSE)
          }
        if (exists(name, env, inherits = FALSE)) {
          # this is a case where we find a binding. the main difference to
          # pryr::where is that we don't return immediately. Instead we save
          # the binding environment to env_list and call where2.internal again
          env_list <<- c(env_list, env)
          where2.internal(name, parent.env(env))
          } else {
            where2.internal(name, parent.env(env))
          }
      }
      
      # as a last step we just call where2.internal() to start the recursion
      where2.internal(name, env = parent.frame())
    }
    ```   
    
    Note that `where2.internal()` still provides the same structure as `pryr::where` does and you can also divide it in "base case", "success case" and "recursive case".

2.  __<span style="color:red">Q</span>__: Write your own version of `get()` using a function written in the style 
    of `where()`.  
    __<span style="color:green">A</span>__: Note that `get()` provides a bit more arguments than our following version, but it should be easy to build up on that. However, we can change `pryr::where` to `get2()` with just changing one line:
    
    ```{r}
    get2 <- function(name, env = parent.frame()) 
      {
      stopifnot(is.character(name), length(name) == 1)
      env <- pryr:::to_env(env)
      if (identical(env, emptyenv())) {
        stop("Can't find ", name, call. = FALSE)
        }
      if (exists(name, env, inherits = FALSE)) {
        # we cancel env and substitute it with the following line, where we subset
        # the environment (like a list) by the name of our object and return it
        return(env[[name]])
        }
      else {
        get2(name, parent.env(env))
      }
    }
    ```

3.  __<span style="color:red">Q</span>__: Write a function called `fget()` that finds only function objects. It 
    should have two arguments, `name` and `env`, and should obey the regular 
    scoping rules for functions: if there's an object with a matching name 
    that's not a function, look in the parent. For an added challenge, also 
    add an `inherits` argument which controls whether the function recurses up 
    the parents or only looks in one environment.  
    __<span style="color:green">A</span>__: 

4.  __<span style="color:red">Q</span>__: Write your own version of `exists(inherits = FALSE)` (Hint: use `ls()`.) 
    Write a recursive version that behaves like `exists(inherits = TRUE)`.  
    __<span style="color:green">A</span>__: 

## Function environments

1.  __<span style="color:red">Q</span>__: List the four environments associated with a function. What does each one
    do? Why is the distinction between enclosing and binding environments
    particularly important?  
    __<span style="color:green">A</span>__:  
    *   Enclosing: where the function is created
    *   Binding: where the function was assigned
    *   Execution: a temporary environment which is created when the function is executed
    *   Calling: the environment from where the function was called  
    
    The difference between binding and enclosing environment is important, because of R's lexical scoping rules. If R can't find an object in the current environment while executing a function, it will look for it in the enclosing environment.
    
2.  __<span style="color:red">Q</span>__: Draw a diagram that shows the enclosing environments of this function:
    
    ```{r, eval = FALSE}
    f1 <- function(x1) {
      f2 <- function(x2) {
        f3 <- function(x3) {
          x1 + x2 + x3
        }
        f3(3)
      }
      f2(2)
    }
    f1(1)
    ```  
    
    __<span style="color:green">A</span>__: 
    
3.  __<span style="color:red">Q</span>__: Expand your previous diagram to show function bindings.  
__<span style="color:green">A</span>__: 

4.  __<span style="color:red">Q</span>__: Expand it again to show the execution and calling environments.  
__<span style="color:green">A</span>__: 

5.  __<span style="color:red">Q</span>__: Write an enhanced version of `str()` that provides more information 
    about functions. Show where the function was found and what environment 
    it was defined in.  
    __<span style="color:green">A</span>__: 

## Binding names to values

1.  __<span style="color:red">Q</span>__: What does this function do? How does it differ from `<<-` and why
    might you prefer it?
    
    ```{r, error = TRUE}
    rebind <- function(name, value, env = parent.frame()) {
      if (identical(env, emptyenv())) {
        stop("Can't find ", name, call. = FALSE)
      } else if (exists(name, envir = env, inherits = FALSE)) {
        assign(name, value, envir = env)
      } else {
        rebind(name, value, parent.env(env))
      }
    }
    rebind("a", 10)
    a <- 5
    rebind("a", 10)
    a
    ```  
    
    __<span style="color:green">A</span>__: 

2.  __<span style="color:red">Q</span>__: Create a version of `assign()` that will only bind new names, never 
    re-bind old names. Some programming languages only do this, and are known 
    as [single assignment languages][single assignment].  
    __<span style="color:green">A</span>__: 

3.  __<span style="color:red">Q</span>__: Write an assignment function that can do active, delayed, and locked 
    bindings. What might you call it? What arguments should it take? Can you 
    guess which sort of assignment it should do based on the input?  
    __<span style="color:green">A</span>__: 