# Functionals

## My first functional: lapply()

1.  __<span style="color:red">Q</span>__: Why are the following two invocations of `lapply()` equivalent?

    ```{r, eval = FALSE}
    trims <- c(0, 0.1, 0.2, 0.5)
    x <- rcauchy(100)
    
    lapply(trims, function(trim) mean(x, trim = trim))
    lapply(trims, mean, x = x)
    ```
    
    __<span style="color:green">A</span>__: In the first statement each element of `trims` is explicitly supplied to `mean()`'s second argument. In the latter statement this happens via 
positional matching, since `mean()`'s first argument is supplied via name
in `lapply()`'s third argument (`...`).

2.  __<span style="color:red">Q</span>__: The function below scales a vector so it falls in the range [0, 1]. How
    would you apply it to every column of a data frame? How would you apply it 
    to every numeric column in a data frame?

    ```{r}
    scale01 <- function(x) {
      rng <- range(x, na.rm = TRUE)
      (x - rng[1]) / (rng[2] - rng[1])
    }
    ```
    
    __<span style="color:green">A</span>__: Since this function needs numeric input, one can check this via an if clause. If one also wants to return non-numeric input columns, these can be supplied to the `else` argument of the `if()` "function":
    
    ```{r, eval = FALSE}
    data.frame(lapply(iris, function(x) if (is.numeric(x)) scale01(x) else x))
    ```

3.  __<span style="color:red">Q</span>__: Use both for loops and `lapply()` to fit linear models to the
    `mtcars` using the formulas stored in this list:

    ```{r}
    formulas <- list(
      mpg ~ disp,
      mpg ~ I(1 / disp),
      mpg ~ disp + wt,
      mpg ~ I(1 / disp) + wt
    )
    ```
    
    __<span style="color:green">A</span>__: Like in the first exercise, we can create two `lapply()` versions:
    
    ```{r, eval = TRUE}
    # lapply (2 versions)
    la1 <- lapply(formulas, lm, data = mtcars)
    la2 <- lapply(formulas, function(x) lm(formula = x, data = mtcars))

    # for loop
    lf1 <- vector("list", length(formulas))
    for (i in seq_along(formulas)){
      lf1[[i]] <- lm(formulas[[i]], data = mtcars)
    }
    ```
    
    Note that all versions return the same content, but they won't be identical, since the values of the "call" element will differ between each version.
    
4.  __<span style="color:red">Q</span>__: Fit the model `mpg ~ disp` to each of the bootstrap replicates of `mtcars` 
    in the list below by using a for loop and `lapply()`. Can you do it 
    without an anonymous function?

    ```{r, eval = TRUE}
    bootstraps <- lapply(1:10, function(i) {
      rows <- sample(1:nrow(mtcars), rep = TRUE)
      mtcars[rows, ]
    })
    ```
    
    __<span style="color:green">A</span>__: 
    
    ```{r, eval = TRUE}
    # lapply without anonymous function
    la <- lapply(bootstraps, lm, formula = mpg ~ disp)
    
    # for loop
    lf <- vector("list", length(bootstraps))
    for (i in seq_along(bootstraps)){
      lf[[i]] <- lm(mpg ~ disp, data = bootstraps[[i]])
    }
    ```

5.  __<span style="color:red">Q</span>__: For each model in the previous two exercises, extract $R^2$ using the
    function below.

    ```{r, eval = TRUE}
    rsq <- function(mod) summary(mod)$r.squared
    ```
    
    __<span style="color:green">A</span>__: For the models in exercise 3:
    
    ```{r, eval = TRUE}
    sapply(la1, rsq)
    sapply(la2, rsq)
    sapply(lf1, rsq)
    ```
    
    And the models in exercise 4:
    
    ```{r, eval = TRUE}
    sapply(la, rsq)
    sapply(lf, rsq)
    ```

## For loops functionals: friends of lapply():

1.  __<span style="color:red">Q</span>__: Use `vapply()` to:
    
    a) Compute the standard deviation of every column in a numeric data frame.
    
    a) Compute the standard deviation of every numeric column in a mixed data
       frame. (Hint: you'll need to use `vapply()` twice.)
       
    __<span style="color:green">A</span>__: As a numeric `data.frame` we choose `cars`:
    
    ```{r, eval = FALSE}
    vapply(cars, sd, numeric(1))
    ```
    
    And as a mixed `data.frame` we choose `iris`:

    ```{r, eval = FALSE}    
    vapply(iris[vapply(iris, is.numeric, logical(1))],
           sd, 
           numeric(1))
    ```

2.  __<span style="color:red">Q</span>__: Why is using `sapply()` to get the `class()` of each element in 
    a data frame dangerous?
    
    __<span style="color:green">A</span>__: Columns of data.frames might have more than one class, so the class of `sapply()`'s output may differ from time to time (silently). If ...
    
    * all columns have one class: `sapply()` returns a character vector
    * one column has more classes than the others: `sapply()` returns a list
    * all columns have the same number of classes, which is more than one: `sapply()` returns a matrix
    
    For example:
    
    ```{r}
    a <- letters[1:3]
    class(a) <- c("class1", "class2")
    df <- data.frame(a = character(3))
    df$a <- a
    df$b <- a
    class(sapply(df, class))
    ```
    
    Note that this case often appears, wile working with the POSIXt types, POSIXct and POSIXlt.
    
3.  __<span style="color:red">Q</span>__: The following code simulates the performance of a t-test for non-normal 
    data. Use `sapply()` and an anonymous function to extract the p-value from 
    every trial.

    ```{r}
    trials <- replicate(
      100, 
      t.test(rpois(10, 10), rpois(7, 10)),
      simplify = FALSE
    )
    ```
    
    Extra challenge: get rid of the anonymous function by using `[[` directly.
    
    __<span style="color:green">A</span>__: 
    
    ```{r, eval = FALSE}
    # anonymous function:
    sapply(trials, function(x) x[["p.value"]])
    # without anonymous function:
    sapply(trials, "[[", "p.value")
    ```

4.  __<span style="color:red">Q</span>__: What does `replicate()` do? What sort of for loop does it eliminate? Why 
    do its arguments differ from `lapply()` and friends?
    
    __<span style="color:green">A</span>__: As stated in `?replicate`:
    
    > replicate is a wrapper for the common use of sapply for repeated evaluation of an expression (which will usually involve random number generation).
    
    We can see this clearly in the source code:
    
    ```{r, echo = FALSE}
    replicate
    ```
    
    Like `sapply()` `replicate()` eliminates a for loop. As explained for `MAP()` in the textbook, also every `replicate()` could have been written via `lapply()`. But using `replicate()` is more concise, and more clearly indicates what you're trying to do.

5.  __<span style="color:red">Q</span>__: Implement a version of `lapply()` that supplies `FUN` with both the name 
    and the value of each component.
    
    __<span style="color:green">A</span>__:
    
    ```{r, eval = TRUE}
    lapply_nms <- function(X, FUN, ...){
      Map(FUN, X, names(X), ...)
    }
    lapply_nms(iris, function(x, y) c(class(x), y))
    ```

6.  __<span style="color:red">Q</span>__: Implement a combination of `Map()` and `vapply()` to create an `lapply()`
    variant that iterates in parallel over all of its inputs and stores its 
    outputs in a vector (or a matrix). What arguments should the function 
    take?

7.  __<span style="color:red">Q</span>__: Implement `mcsapply()`, a multicore version of `sapply()`. Can you
    implement `mcvapply()`, a parallel version of `vapply()`? Why or why not?

## Manipulating matrices and data frames

1.  __<span style="color:red">Q</span>__: How does `apply()` arrange the output? Read the documentation and perform 
    some experiments.
    
    __<span style="color:green">A</span>__:
    
    `apply()` arranges its output columns (or list elements) according to the order of the margin.
The rows are ordered by the other dimensions, starting with the "last" dimension
of the input object. What this means should become clear by looking at the three and four dimensional cases of the following example:

    ```{r, eval = FALSE}
    # for two dimensional cases everything is sorted by the other dimension
    arr2 <- array(1:9, dim = c(3, 3), dimnames = list(paste0("row", 1:3),
                                                  paste0("col", 1:3)))
    arr2
    apply(arr2, 1, head, 1) # Margin is row
    apply(arr2, 1, head, 9) # sorts by col
    
    apply(arr2, 2, head, 1) # Margin is col
    apply(arr2, 2, head, 9) # sorts by row
    
    # 3 dimensional
    arr3 <- array(1:27, dim = c(3,3,3), dimnames = list(paste0("row", 1:3),
                                                    paste0("col", 1:3),
                                                    paste0("time", 1:3)))
    arr3
    apply(arr3, 1, head, 1) # Margin is row
    apply(arr3, 1, head, 27) # sorts by time and col
    
    apply(arr3, 2, head, 1) # Margin is col
    apply(arr3, 2, head, 27) # sorts by time and row
    
    apply(arr3, 3, head, 1) # Margin is time
    apply(arr3, 3, head, 27) # sorts by col and row
    
    # 4 dimensional
    arr4 <- array(1:81, dim = c(3,3,3,3), dimnames = list(paste0("row", 1:3),
                                                      paste0("col", 1:3),
                                                      paste0("time", 1:3),
                                                      paste0("var", 1:3)))
    arr4
    
    apply(arr4, 1, head, 1) # Margin is row
    apply(arr4, 1, head, 81) # sorts by var, time, col
    
    apply(arr4, 2, head, 1) # Margin is col
    apply(arr4, 2, head, 81) # sorts by var, time, row
    
    apply(arr4, 3, head, 1) # Margin is time
    apply(arr4, 3, head, 81) # sorts by var, col, row
    
    apply(arr4, 4, head, 1) # Margin is var
    apply(arr4, 4, head, 81) # sorts by time, col, row
    ```

2.  __<span style="color:red">Q</span>__: There's no equivalent to `split()` + `vapply()`. Should there be? When 
    would it be useful? Implement one yourself.
    
    __<span style="color:green">A</span>__: We can modify the `tapply2()` approach from the book, where `split()` and `sapply()` were combined:
    
    ```{r, eval = FALSE}
    v_tapply <- function(x, group, f, FUN.VALUE, ..., USE.NAMES = TRUE) {
      pieces <- split(x, group)
      vapply(pieces, f, FUN.VALUE, ..., USE.NAMES = TRUE)
    }
    ```
    
    `tapply()` has a `SIMPLIFY` argument. When you set it to `FALSE`, `tapply()` will always return a list. It is easy to create cases where the length and the types/classes of the list elements vary depending on the input. The `vapply()` version could be useful, ff you want to control the structure of the output to get an error according to some logic of a specific usecase or you want typestable output to build up other functions on top of the output.
    
3.  __<span style="color:red">Q</span>__: Implement a pure R version of `split()`. (Hint: use `unique()` and 
    subsetting.) Can you do it without a for loop?
    
    __<span style="color:green">A</span>__: 
    
    ```{r, eval = FALSE}
    split2 <- function(x, f, drop = FALSE, ...){
      levs <- if (drop) {as.character(unique(f))} else {unique(levels(f))}
      lapply(levs, function(lv) x[f == lv, , drop = FALSE])
    }
    ```

4.  __<span style="color:red">Q</span>__: What other types of input and output are missing? Brainstorm before you 
    look up some answers in the [plyr paper](http://www.jstatsoft.org/v40/i01/).
    
## Manipulating lists

1.  __<span style="color:red">Q</span>__: Why isn't `is.na()` a predicate function? What base R function is closest
    to being a predicate version of `is.na()`?

2.  __<span style="color:red">Q</span>__: Use `Filter()` and `vapply()` to create a function that applies a summary 
    statistic to every numeric column in a data frame.
    
    __<span style="color:green">A</span>__: 
    
    ```{r, eval = FALSE}
    vapply_num <- function(X, FUN, FUN.VALUE){
      vapply(Filter(is.numeric, X), FUN, FUN.VALUE)
    }
    ```

3.  __<span style="color:red">Q</span>__: What's the relationship between `which()` and `Position()`? What's
    the relationship between `where()` and `Filter()`?

4.  __<span style="color:red">Q</span>__: Implement `Any()`, a function that takes a list and a predicate function, 
    and returns `TRUE` if the predicate function returns `TRUE` for any of 
    the inputs. Implement `All()` similarly.
    
    __<span style="color:green">A</span>__: `Any()`:
    
    ```{r, eval = FALSE}
    Any <- function(l, pred){
      stopifnot(is.list(l))
      
      for (i in seq_along(l)){
        if (pred(l[[i]])) return(TRUE)
      }
      
      return(FALSE)
    }
    ```
    
    `All()`:
    
    ```{r, eval = FALSE}
    All <- function(l, pred){
      stopifnot(is.list(l))
      
      for (i in seq_along(l)){
        if (!pred(l[[i]])) return(FALSE)
      }
      
      return(TRUE)
    }
    ```

5.  __<span style="color:red">Q</span>__: Implement the `span()` function from Haskell: given a list `x` and a 
    predicate function `f`, `span` returns the location of the longest 
    sequential run of elements where the predicate is true. (Hint: you 
    might find `rle()` helpful.)
    
    __<span style="color:green">A</span>__: Our `span_r()` function returns the first index of     the longest sequential run of elements where the predicate is true. In case of more than one longest sequenital, more than one first_index is returned.
    
    ```{r, eval = FALSE}
    span_r <- function(l, pred){
      # We test if l is a list
      stopifnot(is.list(l))
    
      # we preallocate a logical vector and save the result
      # of the predicate function applied to each element of the list
      test <- vector("logical", length(l))
      for (i in seq_along(l)){
        test[i] <- (pred(l[[i]]))
      }
      # we return NA, if the output of pred is always FALSE
      if(!any(test)) return(NA_integer_)
      
      # Otherwise we look at the length encoding of TRUE and FALSE values.
      rle_test <- rle(test)
      # Since it might happen, that more than one maximum series of TRUE's appears,
      # we have to implement some logic, which might be easier, if we save the rle 
      # output in a data.frmame
      rle_test <- data.frame(lengths = rle_test[["lengths"]],
                              values = rle_test[["values"]],
                              cumsum = cumsum(rle_test[["lengths"]]))
      rle_test[["first_index"]] <- rle_test[["cumsum"]] - rle_test[["lengths"]] + 1
      # In the last line we calculated the first index in the original list for every encoding
      # In the next line we calculate a column, which gives the maximum 
      # encoding length among all encodings with the value TRUE
      rle_test[["max"]] <-  max(rle_test[rle_test[, "values"] == TRUE, ][,"lengths"])
      # Now we just have to subset for maximum length among all TRUE values and return the
      # according "first index":
      rle_test[rle_test$lengths == rle_test$max & rle_test$values == TRUE, ]$first_index
    }
    ```

## Mathematical functionals

1.  __<span style="color:red">Q</span>__: Implement `arg_max()`. It should take a function and a vector of inputs, 
    and return the elements of the input where the function returns the highest 
    value. For example, `arg_max(-10:5, function(x) x ^ 2)` should return -10.
    `arg_max(-5:5, function(x) x ^ 2)` should return `c(-5, 5)`.
    Also implement the matching `arg_min()` function.
    
    __<span style="color:green">A</span>__: `arg_max()`:
    
    ```{r, eval = FALSE}
    arg_max <- function(x, f){
      x[f(x) == max(f(x))]
    }
    ```
    
    `arg_min()`:
    
    ```{r, eval = FALSE}
    arg_min <- function(x, f){
      x[f(x) == min(f(x))]
    }
    ```

2.  __<span style="color:red">Q</span>__: Challenge: read about the 
    [fixed point algorithm](http://mitpress.mit.edu/sicp/full-text/book/book-Z-H-12.html#%_sec_1.3). 
    Complete the exercises using R.

## A family of functions

1.  __<span style="color:red">Q</span>__: Implement `smaller` and `larger` functions that, given two inputs, return 
    either the smaller or the larger value. Implement `na.rm = TRUE`: what 
    should the identity be? (Hint: 
    `smaller(x, smaller(NA, NA, na.rm = TRUE), na.rm = TRUE)` must be `x`, so 
    `smaller(NA, NA, na.rm = TRUE)` must be bigger than any other value of x.) 
    Use `smaller` and `larger` to implement equivalents of `min()`, `max()`,
    `pmin()`, `pmax()`, and new functions `row_min()` and `row_max()`.

2.  __<span style="color:red">Q</span>__: Create a table that has _and_, _or_, _add_, _multiply_, _smaller_, and 
    _larger_ in the columns and _binary operator_, _reducing variant_, 
    _vectorised variant_, and _array variants_ in the rows.

    a) Fill in the cells with the names of base R functions that perform each of
       the roles.

    a) Compare the names and arguments of the existing R functions. How
       consistent are they? How could you improve them?

    a) Complete the matrix by implementing any missing functions.

3.  __<span style="color:red">Q</span>__: How does `paste()` fit into this structure? What is the scalar binary 
    function that underlies `paste()`? What are the `sep` and `collapse` 
    arguments to `paste()` equivalent to? Are there any `paste` variants 
    that don't have existing R implementations?