# Function operators

## Behavioural FOs

1.  __<span style="color:red">Q</span>__: Write a FO that logs a time stamp and message to a file every time a 
    function is run.
    
    __<span style="color:green">A</span>__: Note that the example will create a file file in your current working directory:
    
    ```{r, eval = FALSE}
    logger <- function(f, filename){
      force(f)
      filename_tmp <- paste(filename, basename(tempfile()), sep = "_")
      write(paste("created at:", Sys.time()), filename_tmp, append = TRUE)
      function(..., message = "you can add a message at each call") {
        write(paste0("used at: ", Sys.time(), ", ", message), filename_tmp, append = TRUE)
        f(...)
      }
    }
    
    # the following line creates a file, which name starts with "mean_log_"
    mean2 <- logger(mean, "mean_log") 
    mean2(1:4, message = "first time") 
    mean2(1:4, message = "second_time")
    ```

2.  __<span style="color:red">Q</span>__: What does the following function do? What would be a good name for it?

    ```{r}
    f <- function(g) {
      force(g)
      result <- NULL
      function(...) {
        if (is.null(result)) {
          result <<- g(...)
        }
        result
      }
    }
    runif2 <- f(runif)
    runif2(5)
    runif2(10)
    ```
    
    __<span style="color:green">A</span>__: It returns a new version of the inputfunction. That version will always return the result of it's first run (in case this not `NULL`), no matter how the input changes. Good names could be `first_run()` or `initial_return()`. 

3.  __<span style="color:red">Q</span>__: Modify `delay_by()` so that instead of delaying by a fixed amount of time, 
    it ensures that a certain amount of time has elapsed since the function 
    was last called. That is, if you called 
    `g <- delay_by(1, f); g(); Sys.sleep(2); g()` there shouldn't be an 
    extra delay.
    
    __<span style="color:green">A</span>__: We can do this with three little tricks (and the help
    of 42):
    
    ```{r, eval = FALSE}
    delay_by_v2 <- function(delay, f) {
      force(f)
      # we initialise the timestamp for the last run. We set a specific default value,
      # to ensure that the first run of the returned function will never be delayed
      last_runtime <- Sys.time() - (delay + 42)
      function(...) {
        # we continually check if enough time passed with an (empty) while statement.
        while (Sys.time() < last_runtime + delay) {}
        # we override the start for the next waiting interval.
        # Note that this is done on exit (after the function is evaluated)
        on.exit(last_runtime <<- Sys.time()) 
        return(f(...))
      }
    }
    ```
    
    Alternatively to the empty while statement we could have used `Sys.sleep()`. I would not recommend this solution, since `?Sys.sleep` indicates that `Sys.sleep()` might have some overhead and seems not to be as exact as we need.

4.  __<span style="color:red">Q</span>__: Write `wait_until()` which delays execution until a specific time.

    __<span style="color:green">A</span>__: 

    ```{r, eval = FALSE}
    wait_until <- function(time, f) {
      force(f)
      function(...) {
        while (Sys.time() < time) {}
        return(f(...))
      }
    }
    
    # a little test
    ptm <- proc.time()
    m <- wait_until(Sys.time() + 10, mean)
    m(1:3)
    proc.time() - ptm
    ```

5.  __<span style="color:red">Q</span>__: There are three places we could have added a memoise call: why did we 
    choose the one we did?

    ```{r, eval = FALSE}
    download <- memoise(dot_every(10, delay_by(1, download_file)))
    download <- dot_every(10, memoise(delay_by(1, download_file)))
    download <- dot_every(10, delay_by(1, memoise(download_file)))
    ```
    
    __<span style="color:green">A</span>__: The second was chosen. It's easy to see why, if we eliminate the other two options:
    
    * The first version only prints a dot at every tenth `download()` call with a new input.
    This is because `dot_every()` is inside of `memoise()` and the counter created by
    `dot_every()` is not "activated" if the input is known.
    
    * The third version takes one second for every call. Even if we already know the result and
    don't download anything again.
    
6.  __<span style="color:red">Q</span>__: Why is the `remember()` function inefficient? How could you implement it 
    in more efficient way?

7.  __<span style="color:red">Q</span>__: Why does the following code, from 
    [stackoverflow](http://stackoverflow.com/questions/8440675), not do what you expect?

    ```{r}
    # return a linear function with slope a and intercept b.
    f <- function(a, b) function(x) a * x + b

    # create a list of functions with different parameters.
    fs <- Map(f, a = c(0, 1), b = c(0, 1))

    fs[[1]](3)
    # should return 0 * 3 + 0 = 0
    ```

    How can you modify `f` so that it works correctly?
    
    __<span style="color:green">A</span>__: You can read in the [stackoverflow](http://stackoverflow.com/questions/8440675) link that the question arose, because the original return of 
    `fs[[1]](3)` was `4`, which is due to lazy evaluation and could be solved by two users via `force()`:
    
    ```{r, eval = FALSE}
    f <- function(a, b) {force(a); force(b); function(x) a * x + b}
    ```
    
    However you can see in the result within the question that **R**'s behaviour was changed in this case and as Jan Kislinger points out on [twitter](https://twitter.com/JanKislinger/status/794433891486547968):
    
    > The real question should be: "How did they modify #rstats so that it works correctly?" otherwise it's a tricky question :D
    
    Note that the same issue appears in the [textbook](http://adv-r.had.co.nz/Function-operators.html#behavioural-fos):
    
    > In the following example, we take a list of functions and delay each one. But when we try to evaluate the mean, we get the sum instead. 
    
    ```{r, eval = FALSE}
    funs <- list(mean = mean, sum = sum)
    funs_m <- lapply(funs, delay_by, delay = 0.1)

    funs_m$mean(1:10)
    #> [1] 5.5
    ```
    
    Which (as one can see) is not true anymore...actually it changed in R version [**3.2**](https://stat.ethz.ch/pipermail/r-announce/2015/000583.html):
    
    > Higher order functions such as the apply functions and Reduce()
     now force arguments to the functions they apply in order to
     eliminate undesirable interactions between lazy evaluation and
     variable capture in closures.  This resolves PR#16093.
     
     For further interested: [PR#16093](https://bugs.r-project.org/bugzilla3/show_bug.cgi?id=16093#c1) will lead you to the subject "iterated lapply" within the 
     [R-devel Archives](https://stat.ethz.ch/pipermail/r-devel/2015-March/subject.html#start). Note that the behaviour in for loops is still as "the old `lapply()`" behaviour.
    
## Output FOs

1.  __<span style="color:red">Q</span>__: Create a `negative()` FO that flips the sign of the output of the 
    function to which it is applied.
    
    __<span style="color:green">A</span>__: 
    
    ```{r, eval = FALSE}
    negative <- function(f){
      force(f)
      function(...){
        -f(...)
      }
    }
    ```

2.  __<span style="color:red">Q</span>__: The `evaluate` package makes it easy to capture all the outputs (results, 
    text, messages, warnings, errors, and plots) from an expression. Create a 
    function like `capture_it()` that also captures the warnings and errors 
    generated by a function.
    
    __<span style="color:green">A</span>__: One way is just to capture the output of  `tryCatch()` with identity handlers for errors and warnings:
    
    ```{r, eval = TRUE}
    capture_trials <- function(f){
      force(f)
      function(...){
        capture.output(tryCatch(f(...),
                                error = function(e) e,
                                warning = function(w) w)
        )
      }
    }

    # we test the behaviour
    log_t <- capture_trials(log)
    elements <- list(1:10, c(-1, 10), c(TRUE, FALSE), letters)
    results <- lapply(elements, function(x) log_t(x))
    results
    
    # further
    # results_detailed <- lapply(elements, function(x) lapply(x, function(y))log2(x))
    # results_detailed
    ```

3.  __<span style="color:red">Q</span>__: Create a FO that tracks files created or deleted in the working directory 
    (Hint: use `dir()` and `setdiff()`.) What other global effects of 
    functions might you want to track?
    
    __<span style="color:green">A</span>__: We start with a short version to show the idea:
    
    ```{r, eval = FALSE}
    track_dir <- function(f){
      force(f)
      function(...){
        dir_old <- dir()
        on.exit(if(!setequal(dir(), dir_old)){
          message("files in your working directory were deleted or added by this function")})
        f(...)
      }
    }
    
    # the following test will create the file "delete_me" in your working directory
    td <- track_dir(dir.create)
    td("delete_me")
    ```
    
    Of course we can provide more information on the type of changes:
    
    ```{r, eval = FALSE}
    track_dir <- function(f){
      force(f)
      function(...){
        dir_old <- dir()
        
        on.exit(if(!setequal(dir(), dir_old)){
          message("Files in your working directory were deleted or added by this
                  function.")}, add = TRUE)
        on.exit(if(length(setdiff(dir_old, dir()) != 0)){
          message(paste0("The following files were deleted: ",
                         paste(setdiff(dir_old, dir()), collapse = ", ")
                         ))}, add = TRUE)
        on.exit(if(length(setdiff(dir(), dir_old) != 0)){
          message(paste0("The following files were added: ", 
                         paste(setdiff(dir(), dir_old), collapse = ", ")
                         ))}, add = TRUE)
        
        f(...)
      }
    }

    # the following test will again create two files in your working directory
    td <- track_dir(sapply)
    td(c("delete_me", "me_too"), dir.create)
    ```
    
    Other global effects that might be worth tracking include changes regarding:
    
    * the search path and/or introduced `conflicts()`
    * `options()` and `par()` which modify global settings
    * the path of the working directory
    * environment variables
    * the locale.

## Input FOs

1.  __<span style="color:red">Q</span>__: Our previous `download()` function only downloads a single file. How can 
    you use `partial()` and `lapply()` to create a function that downloads 
    multiple files at once? What are the pros and cons of using `partial()` vs. 
    writing a function by hand?

2.  __<span style="color:red">Q</span>__: Read the source code for `plyr::colwise()`. How does the code work? What 
    are `colwise()`'s three main tasks? How could you make `colwise()` simpler 
    by implementing each task as a function operator? (Hint: think about 
    `partial()`.)
    
    __<span style="color:orange">A</span>__: We describe how it works by commenting the source code:
    
    ```{r, eval = FALSE}
    function (.fun, .cols = true, ...) 
      {
      # We check if .cols is not a function, since it is possible to supply a
      # predicate function.
      # if so, the .cols arguments will be "quoted", and filter() will 
      # be a function that checks and evaluates these .cols within its other argument
      if (!is.function(.cols)) {
        .cols <- as.quoted(.cols)
        filter <- function(df) eval.quoted(.cols, df)
      }
      # otherwise, filter will be be Filter(), which applies the function 
      # in .cols to every element of its other argument
      else {
        filter <- function(df) Filter(.cols, df)
      }
      # the ... arguments are caught in the list dots
      dots <- list(...)
      # a function is created, which will also be the return value.
      # it checks if its input is a data frame
      function(df, ...) {
        stopifnot(is.data.frame(df))
        # if df is split (in "plyr" speaking), this will be taken into account...
        df <- strip_splits(df)
        # now the columns of the data frame are chosen, depending on the input of .cols
        # this can chosen directly, via a predicate function, or all columns (default)
        filtered <- filter(df)
        # if this means, that no columns are selected, an empty data frame will be returned
        if (length(filtered) == 0) 
          return(data.frame())
        # otherwise lapply will be called on all filtered columns, with 
        # the .fun argument, which has to be provided by the user, and some other
        # arguments provided by the user, when calling the function (...) and
        # when defining the function (dots)
        out <- do.call("lapply", c(list(filtered, .fun, ...), 
                               dots))
        # the output will be named and converted from list into a data frame again
        names(out) <- names(filtered)
        quickdf(out)
      }
    }
    
    <environment: namespace:plyr>
    ```

3.  __<span style="color:red">Q</span>__: Write FOs that convert a function to return a matrix instead of a data 
    frame, or a data frame instead of a matrix. If you understand S3, 
    call them `as.data.frame.function()` and `as.matrix.function()`.
    
    __<span style="color:green">A</span>__: 
    
    ```{r, eval = FALSE}
    as.matrix.function <- function(f){
      force(f)
      function(...){
        as.matrix(f(...))
      }
    }
    
    as.data.frame.function <- function(f){
      force(f)
      function(...){
        as.data.frame(f(...))
      }
    }
    ```

4.  __<span style="color:red">Q</span>__: You've seen five functions that modify a function to change its output 
    from one form to another. What are they? Draw a table of the various 
    combinations of types of outputs: what should go in the rows and what 
    should go in the columns? What function operators might you want to write 
    to fill in the missing cells? Come up with example use cases.

5.  __<span style="color:red">Q</span>__: Look at all the examples of using an anonymous function to partially 
    apply a function in this and the previous chapter. Replace the anonymous 
    function with `partial()`. What do you think of the result? Is it easier or 
    harder to read?
    
    __<span style="color:green">A</span>__: The results are easy to read. Especially the `Map()` examples profit in readability:
    
    ```{r}
    library(pryr)
    ## From Functionals
    # 1
    trims <- c(0, 0.1, 0.2, 0.5)
    x <- rcauchy(1000)
    unlist(lapply(trims, function(trim) mean(x, trim = trim)))
    unlist(lapply(trims, partial(mean, x)))
    
    # 2
    xs <- replicate(5, runif(10), simplify = FALSE)
    ws <- replicate(5, rpois(10, 5) + 1, simplify = FALSE)
    unlist(Map(function(x, w) weighted.mean(x, w, na.rm = TRUE), xs, ws))
    unlist(Map(partial(weighted.mean, na.rm = TRUE), xs, ws))
    
    # 3
    add <- function(x, y, na.rm = FALSE) {
      if (na.rm && (is.na(x) || is.na(y))) rm_na(x, y, 0) else x + y
    }
  
    r_add <- function(xs, na.rm = TRUE) {
      Reduce(function(x, y) add(x, y, na.rm = na.rm), xs)
    }
    
    r_add_compact <- function(xs, na.rm = TRUE) {
      Reduce(partial(add, na.rm = na.rm), xs)
    }
    
    r_add(1:4)
    r_add_compact(1:4)
    
    # 4
    v_add1 <- function(x, y, na.rm = FALSE) {
      stopifnot(length(x) == length(y), is.numeric(x), is.numeric(y))
      if (length(x) == 0) return(numeric())
      simplify2array(
        Map(function(x, y) add(x, y, na.rm = na.rm), x, y)
      )
    }
  
    v_add1_compact <- function(x, y, na.rm = FALSE) {
      stopifnot(length(x) == length(y), is.numeric(x), is.numeric(y))
      if (length(x) == 0) return(numeric())
      simplify2array(
        Map(partial(add, na.rm = na.rm), x, y)
      )
    }
    
    v_add1(1:3, 2:4)
    v_add1_compact(1:3, 2:4)
    
    # 5
    c_add <- function(xs, na.rm = FALSE) {
      Reduce(function(x, y) add(x, y, na.rm = na.rm), xs,
             accumulate = TRUE)
    }
    
    c_add_compact <- function(xs, na.rm = FALSE) {
      Reduce(partial(add, na.rm = na.rm), xs,
             accumulate = TRUE)
    }
    
    c_add(1:3)
    c_add_compact(1:3)
    
    ## From Function operators
    # 6
    f <- function(x) x ^ 2
    partial(f)
    
    # 7
    # Map(function(x, y) f(x, y, zs), xs, ys)
    # Map(partial(f, zs = zs), xs, yz)
    
    # 8
    # f <- function(a) g(a, b = 1)
    # f <- partial(g, b = 1)
    
    # 9
    compact <- function(x) Filter(Negate(is.null), x)
    compact <- partial(Filter, Negate(is.null))
    
    # 10
    # Map(function(x, y) f(x, y, zs), xs, ys)
    # Map(partial(f, zs = zs), xs, ys)
    
    # 11
    funs2 <- list(
      sum = function(...) sum(..., na.rm = TRUE),
      mean = function(...) mean(..., na.rm = TRUE),
      median = function(...) median(..., na.rm = TRUE)
    )
    
    funs2 <- list(
      sum = partial(sum, na.rm = TRUE),
      mean = partial(mean, na.rm = TRUE),
      median = partial(median, na.rm = TRUE)
    )
    ```

## Combining FOs

1.  __<span style="color:red">Q</span>__: Implement your own version of `compose()` using `Reduce` and `%o%`. For 
    bonus points, do it without calling `function`.
    
    __<span style="color:green">A</span>__: We use the definition from the textbook:
    
    ```{r}
    compose <- function(f, g) {
      function(...) f(g(...))
    }
    
    "%o%" <- compose
    ```
    
    And then we build two versions. One via an anonymous function and one via `partial()`:
    
    ```{r, eval}
    compose_red <- function(fs) {
      Reduce(function(f, g) function(...) f(g(...)), fs)
    }
    compose_red(c(mean, length, unique))(1:10)
    
    compose_red_bonus <- function(fs) {
      Reduce(partial(partial(`%o%`)), fs)
    }
    compose_red_bonus(c(mean, length, unique))(1:10)
    ```

2.  __<span style="color:red">Q</span>__: Extend `and()` and `or()` to deal with any number of input functions. Can 
    you do it with `Reduce()`? Can you keep them lazy (e.g., for `and()`, the 
    function returns once it sees the first `FALSE`)?
    
    __<span style="color:green">A</span>__: We use `and()` and `or()` as defined in the textbook. They are lazy, since they are build up on `&&` and `||`. Also their reduced versions stay lazy, as we will show at the end of the code
    
    ```{r}
    and <- function(f1, f2) {
      force(f1); force(f2)
      function(...) {
        f1(...) && f2(...)
      }
    }
    
    and_red <- function(fs){
      Reduce(function(f, g) and(f, g), fs)
    }
    
    or <- function(f1, f2) {
      force(f1); force(f2)
      function(...) {
        f1(...) || f2(...)
      }
    }
    
    or_red <- function(fs){
      Reduce(function(f, g) or(f, g), fs)
    }
    
    # Errors before the first TRUE will be returned
    tryCatch(
      or_red(c(is.logical, is.logical, stop, is.character))("a"), 
      error = function(e) e
    )
    
    # Errors after the first TRUE won't be returned
    or_red(c(is.logical, is.logical, is.character, stop))("a")
    ```

3.  __<span style="color:red">Q</span>__: Implement the `xor()` binary operator. Implement it using the existing 
    `xor()` function. Implement it as a combination of `and()` and `or()`. What 
    are the advantages and disadvantages of each approach? Also think about 
    what you'll call the resulting function to avoid a clash with the existing
    `xor()` function, and how you might change the names of `and()`, `not()`, 
    and `or()` to keep them consistent.

4.  __<span style="color:red">Q</span>__: Above, we implemented boolean algebra for functions that return a logical 
    function. Implement elementary algebra (`plus()`, `minus()`, `multiply()`, 
    `divide()`, `exponentiate()`, `log()`) for functions that return numeric 
    vectors.
    
    __<span style="color:green">A</span>__:
    
    ```{r, eval = FALSE}
    plus <- function(f1, f2) {
      force(f1); force(f2)
      function(...) {
        f1(...) + f2(...)
      }
    }
    
    minus <- function(f1, f2) {
      force(f1); force(f2)
      function(...) {
        f1(...) - f2(...)
      }
    }
    
    multiply <- function(f1, f2) {
      force(f1); force(f2)
      function(...) {
        f1(...) * f2(...)
      }
    }
    
    divide <- function(f1, f2) {
      force(f1); force(f2)
      function(...) {
        f1(...) / f2(...)
      }
    }
    
    exponentiate <- function(f1, f2) {
      force(f1); force(f2)
      function(...) {
        f1(...) ^ f2(...)
      }
    }
    
    # we rename log to log_ since log() already exists
    log_ <- function(f1, f2) {
      force(f1); force(f2)
      function(...) {
        log(f1(...), f2(...))
      }
    }
    
    # Test
    mns <- minus(mean, function(x) x^2)
    mns(1:5)
    ```