# Non standard evaluation

## Capturing expressions

1.  __<span style="color:red">Q</span>__: One important feature of `deparse()` to be aware of when programming is that 
    it can return multiple strings if the input is too long. For example, the 
    following call produces a vector of length two:

    ```{r, eval = FALSE}
    g(a + b + c + d + e + f + g + h + i + j + k + l + m +
      n + o + p + q + r + s + t + u + v + w + x + y + z)
    ```

    Why does this happen? Carefully read the documentation for `?deparse`. Can you write a
    wrapper around `deparse()` so that it always returns a single string?

2.  __<span style="color:red">Q</span>__: Why does `as.Date.default()` use `substitute()` and `deparse()`?
    Why does `pairwise.t.test()` use them? Read the source code.

3.  __<span style="color:red">Q</span>__: `pairwise.t.test()` assumes that `deparse()` always returns a length one 
    character vector. Can you construct an input that violates this expectation? 
    What happens?

4.  __<span style="color:red">Q</span>__: `f()`, defined above, just calls `substitute()`. Why can't we use it
    to define `g()`? In other words, what will the following code return?
    First make a prediction. Then run the code and think about the results.

    ```{r, eval = FALSE}
    f <- function(x) substitute(x)
    g <- function(x) deparse(f(x))
    g(1:10)
    g(x)
    g(x + y ^ 2 / z + exp(a * sin(b)))
    ```

## Non standard evaluation in subset

1.  __<span style="color:red">Q</span>__: Predict the results of the following lines of code:

    ```{r, eval = FALSE}
    eval(quote(eval(quote(eval(quote(2 + 2))))))
    eval(eval(quote(eval(quote(eval(quote(2 + 2)))))))
    quote(eval(quote(eval(quote(eval(quote(2 + 2)))))))
    ```

2.  __<span style="color:red">Q</span>__: `subset2()` has a bug if you use it with a single column data frame.
    What should the following code return? How can you modify `subset2()`
    so it returns the correct type of object?

    ```{r}
    subset2 <- function(x, condition) {
  condition_call <- substitute(condition)
  r <- eval(condition_call, x)
  x[r, ]
}
    sample_df2 <- data.frame(x = 1:10)
    subset2(sample_df2, x > 8)
    ```

3.  __<span style="color:red">Q</span>__: The real subset function (`subset.data.frame()`) removes missing
    values in the condition. Modify `subset2()` to do the same: drop the 
    offending rows.

4.  __<span style="color:red">Q</span>__: What happens if you use `quote()` instead of `substitute()` inside of
    `subset2()`?

5.  __<span style="color:red">Q</span>__: The second argument in `subset()` allows you to select variables. It
    treats variable names as if they were positions. This allows you to do 
    things like `subset(mtcars, , -cyl)` to drop the cylinder variable, or
    `subset(mtcars, , disp:drat)` to select all the variables between `disp`
    and `drat`. How does this work? I've made this easier to understand by
    extracting it out into its own function.

    ```{r, eval = FALSE}
    select <- function(df, vars) {
      vars <- substitute(vars)
      var_pos <- setNames(as.list(seq_along(df)), names(df))
      pos <- eval(vars, var_pos)
      df[, pos, drop = FALSE]
    }
    select(mtcars, -cyl)
    ```

6.  __<span style="color:red">Q</span>__: What does `evalq()` do? Use it to reduce the amount of typing for the
    examples above that use both `eval()` and `quote()`.

## Scoping issues

1.  __<span style="color:red">Q</span>__: `plyr::arrange()` works similarly to `subset()`, but instead of selecting
    rows, it reorders them. How does it work? What does
    `substitute(order(...))` do? Create a function that does only that
    and experiment with it.

2.  __<span style="color:red">Q</span>__: What does `transform()` do? Read the documentation. How does it work?
    Read the source code for `transform.data.frame()`. What does
    `substitute(list(...))` do?

3.  __<span style="color:red">Q</span>__: `plyr::mutate()` is similar to `transform()` but it applies the
    transformations sequentially so that transformation can refer to columns
    that were just created:

    ```{r, eval = FALSE}
    df <- data.frame(x = 1:5)
    transform(df, x2 = x * x, x3 = x2 * x)
    plyr::mutate(df, x2 = x * x, x3 = x2 * x)
    ```

    How does mutate work? What's the key difference between `mutate()` and
    `transform()`?

4.  __<span style="color:red">Q</span>__: What does `with()` do? How does it work? Read the source code for
    `with.default()`. What does `within()` do? How does it work? Read the
    source code for `within.data.frame()`. Why is the code so much more
    complex than `with()`?

## Calling from another function

1.  __<span style="color:red">Q</span>__: The following R functions all use NSE. For each, describe how it uses NSE,
    and read the documentation to determine its escape hatch.
    * `rm()`
    * `library()` and `require()`
    * `substitute()`
    * `data()`
    * `data.frame()`

2.  __<span style="color:red">Q</span>__: Base functions `match.fun()`, `page()`, and `ls()` all try to
    automatically determine whether you want standard or non-standard
    evaluation. Each uses a different approach. Figure out the essence
    of each approach then compare and contrast.

3.  __<span style="color:red">Q</span>__: Add an escape hatch to `plyr::mutate()` by splitting it into two functions.
    One function should capture the unevaluated inputs. The other should take a 
    data frame and list of expressions and perform the computation.

4.  __<span style="color:red">Q</span>__: What's the escape hatch for `ggplot2::aes()`? What about `plyr::.()`?
    What do they have in common? What are the advantages and disadvantages
    of their differences?

5.  __<span style="color:red">Q</span>__: The version of `subset2_q()` I presented is a simplification of real
    code. Why is the following version better?

    ```{r}
    subset2_q <- function(x, cond, env = parent.frame()) {
      r <- eval(cond, x, env)
      x[r, ]
    }
    ```

    Rewrite `subset2()` and `subscramble()` to use this improved version.

## Substitute

1.  __<span style="color:red">Q</span>__: Use `subs()` to convert the LHS to the RHS for each of the following pairs:
    * `a + b + c` -> `a * b * c`
    * `f(g(a, b), c)` -> `(a + b) * c`
    * `f(a < b, c, d)` -> `if (a < b) c else d`

2.  __<span style="color:red">Q</span>__: For each of the following pairs of expressions, describe why you can't
    use `subs()` to convert one to the other.
    * `a + b + c` -> `a + b * c`
    * `f(a, b)` -> `f(a, b, c)`
    * `f(a, b, c)` -> `f(a, b)`

3.  __<span style="color:red">Q</span>__: How does `pryr::named_dots()` work? Read the source.

## The downsides of non-standard evaluation

1. __<span style="color:red">Q</span>__: What does the following function do? What’s the escape hatch? Do you think that this is an appropriate use of NSE?

    nl <- function(...) {
      dots <- named_dots(...)
      lapply(dots, eval, parent.frame())
    }

2. __<span style="color:red">Q</span>__: Instead of relying on promises, you can use formulas created with ~ to explicitly capture an expression and its environment. What are the advantages and disadvantages of making quoting explicit? How does it impact referential transparency?

3. __<span style="color:red">Q</span>__: Read the standard non-standard evaluation rules found at http://developer.r-project.org/nonstandard-eval.pdf.