
```{r, include=FALSE}
source("common.R")
```

# Function factories

## Prerequisites

For most of this chapter base R is sufficient. Just a few exercises require the *rlang*, *dplyr*, *purrr* and *ggplot2* packages.

```{r, message=FALSE}
library(rlang)
library(dplyr)
library(purrr)
library(ggplot2)
```

## Factory fundamentals

__[Q1]{.Q}__: The definition of `force()` is simple:

```{r}
force
```

Why is it better to `force(x)` instead of just `x`?

__[A]{.solved}__: As you can see `force(x)` is just syntactic sugar for `x`. As being mentioned in the textbook we prefer this explicit form, because

> using this function clearly indicates that you’re forcing evaluation, not that you’ve accidentally typed `x`."

__[Q2]{.Q}__: Base R contains two function factories, `approxfun()` and `ecdf()`. Read their documentation and experiment to figure out what the functions do and what they return.

__[A]{.solved}__: Let's begin with `approxfun()` as it is used within `ecdf()` as well:

`approxfun()` takes a combination of data points (x and y values) as input and returns a stepwise linear (or constant) interpolation function. To find out what this means exactly, we first create a few random data points.

```{r}
x <- runif(10)
y <- runif(10)
plot(x, y, lwd = 10)
```

Next, we use approxfun to construct the linear and constant interpolation functions for our x and y values.

```{r}
f_lin <- approxfun(x, y)
f_con <- approxfun(x, y, method = "constant")

# Both functions exactly reproduce their input y values
identical(f_lin(x), y)
identical(f_con(x), y)
```

When we apply these functions to new x values, these are mapped to the lines connecting the initial y values (linear case) or to the same y value as for the next smallest initial x value (constant case).

```{r}
x_new <- runif(1000)

plot(x, y, lwd = 10)
points(x_new, f_lin(x_new), col = "red", pch = 16)
points(x_new, f_con(x_new), col = "blue", pch = 16)
```

However, both functions are only defined within `range(x)`.

```{r}
f_lin(range(x))
f_con(range(x))

(eps <- .Machine$double.neg.eps)

f_lin(c(min(x) - eps, max(x) + eps))
f_con(c(min(x) - eps, max(x) + eps))
```

To change this behaviour, one can set `rule = 2`. This leads to the result that for values outside of `range(x)` the boundary values of the function are returned.

```{r}
f_lin <- approxfun(x, y, rule = 2)
f_con <- approxfun(x, y, method = "constant", rule = 2)

f_lin(c(-Inf, Inf))
f_con(c(-Inf, Inf))
```

Another option is to customize the return values as individual constants for each side via `yleft` and/or `yright`.

```{r}
f_lin <- approxfun(x, y, yleft = 10)
f_con <- approxfun(x, y, method = "constant", yleft = 10, yright = -10)

f_lin(c(-Inf, Inf))
f_con(c(-Inf, Inf))
```

Further, `approxfun()` provides the option to shift the y values for `method = "constant"` between their left and right values. According to the documentation this indicates a compromise between left- and right-continuous steps.

```{r}
f_con <- approxfun(x, y, method = "constant", f = .5)

plot(x, y, lwd = 10)
points(x_new, f_con(x_new), pch = 16)
```

Finally, the `ties` argument allows to aggregate y values if multiple ones were provided for the same x value. E.g. in the following line we use `mean()` to aggregate these y values before they are used for the interpolation `approxfun(x = c(1,1,2), y = 1:3, ties = mean)`.

Next, we focus on `ecdf`. `ecdf()` is an acronym for empirical cumulative distribution function. For a numeric vector of density values, `ecdf()` initially creates the (x, y) pairs for the nodes of the density function and then passes these pairs to `approxfun()`, which gets called with specifically adapted settings (`approxfun(vals, cumsum(tabulate(match(x, vals)))/n, method = "constant", yleft = 0, yright = 1, f = 0, ties = "ordered")`). 

```{r}
x <- runif(10)
f_ecdf <- ecdf(x)
class(f_ecdf)

plot(x, f_ecdf(x), lwd = 10, ylim = 0:1)
```

New values are then mapped on the y value of the next smallest x value from within the initial input.

```{r}
x_new <- runif(1000)

plot(x, f_ecdf(x), lwd = 10, ylim = 0:1)
points(x_new, f_ecdf(x_new), ylim = 0:1)
```

__[Q3]{.Q}__: Create a function `pick()` that takes an index, `i`, as an argument and returns a function with an argument `x` that subsets `x` with `i`.

```{r, eval = FALSE}
pick(1)(x)
# should be equivalent to
x[[1]]

lapply(mtcars, pick(5))
# should be equivalent to
lapply(mtcars, function(x) x[[5]])
```

__[A]{.solved}__: In this exercise `pick(i)` acts as a function factory, which returns the required subsetting function.

```{r}
pick <- function(i) {
  force(i)
  
  function(x) x[[i]]
}

x <- 1:3
identical(x[[1]], pick(1)(x))
identical(lapply(mtcars, function(x) x[[5]]),
          lapply(mtcars, pick(5)))
```

__[Q4]{.Q}__: Create a function that creates functions that compute the i^th^ [central moment](http://en.wikipedia.org/wiki/Central_moment) of a numeric vector. You can test it by running the following code:

```{r, eval = FALSE}
m1 <- moment(1)
m2 <- moment(2)

x <- runif(100)
stopifnot(all.equal(m1(x), 0))
stopifnot(all.equal(m2(x), var(x) * 99 / 100))
```

__[A]{.solved}__: The first moment is closely related to the mean and describes the average deviation from the mean, which is 0 (within numerical margin of error). The second moment describes the variance of the input data. If we want compare it to `var`, we need to undo [Bessel's correction](https://en.wikipedia.org/wiki/Bessel%27s_correction) correction by multiplying with $\frac{N-1}{N}$.

```{r}
moment <- function(i){
  force(i)
  
  function(x) sum((x - mean(x)) ^ i) / length(x)
}

m1 <- moment(1)
m2 <- moment(2)

x <- runif(100)
all.equal(m1(x), 0)  # removed stopifnot() for clarity
all.equal(m2(x), var(x) * 99 / 100)
```

__[Q5]{.Q}__: What happens if you don't use a closure? Make predictions, then verify with the code below.

```{r}
i <- 0
new_counter2 <- function() {
  i <<- i + 1
  i
}
```

__[A]{.solved}__: Without the captured and encapsulated environment of a closure the counts will be stored in the global environment. Here they can be overwritten or deleted as well as interfere with other counters.

```{r, error = TRUE}
new_counter2()
i
new_counter2()
i

i <- 0
new_counter2()
i
```

__[Q6]{.Q}__: What happens if you use `<-` instead of `<<-`? Make predictions, then verify with the code below.

```{r}
new_counter3 <- function() {
  i <- 0
  function() {
    i <- i + 1
    i
  }
}
```

__[A]{.solved}__: Without the super assignment `<<-`, the counter will always return 1. The counter always starts in a new execution environment within the same enclosing environment, which contains an unchanged value for `i` (in this case it remains 0).

```{r}
new_counter_3 <- new_counter3()

new_counter_3()
new_counter_3()
```


## Graphical factories

__[Q1]{.Q}__: Compare and contrast `ggplot2::label_bquote()` with `scales::number_format()`.

__[A]{.solved}__: Both functions will help you in styling your output, e.g. in your plots and they do this by returning the desired formatting function to you.

`ggplot2::label_bquote()` takes relatively straightforward [plotmath](https://stat.ethz.ch/R-manual/R-patched/library/grDevices/html/plotmath.html) expressions and uses them for faceting labels in ggplot2. Because this function is used in ggplot2 it needs to return a function of `class = "labeller"`.

`scales::number_format()` initially `force()`s the computation of all parameters. It's essentially a parametrized wrapper around `scales::number()` and will help you format numbers appropriately. It will return a simple function.

## Statistical factories

__[Q1]{.Q}__: In `boot_model()`, why don't I need to force the evaluation of `df` or `model`?

__[A]{.solved}__: `boot_model()` ultimately returns a function, and whenever you return a function you need to make sure all the inputs are explicitly evaluated. Here that happens automatically because we use `df` and `formula` in `lm()`.

```{r}
boot_model <- function(df, formula) {
  mod <- lm(formula, data = df)
  fitted <- unname(fitted(mod))
  resid <- unname(resid(mod))
  rm(mod)
  
  function() {
    fitted + sample(resid)
  }
} 
```

__[Q2]{.Q}__: Why might you formulate the Box-Cox transformation like this?

```{r}
boxcox3 <- function(x) {
  function(lambda) {
    if (lambda == 0) {
      log(x)
    } else {
      (x ^ lambda - 1) / lambda
    }
  }  
}
```

__[A]{.solved}__: `boxcox3()` returns a function where `x` is fixed (though it is not forced, so it may manipulated later). This allows us to apply and test different transformations for different inputs and give them a descriptive name.

```{r, out.width = "49%", fig.show = "hold"}
boxcox_airpassengers <- boxcox3(AirPassengers)

plot(boxcox_airpassengers(0))
plot(boxcox_airpassengers(1))
plot(boxcox_airpassengers(2))
plot(boxcox_airpassengers(3))
```

__[Q3]{.Q}__: Why don't you need to worry that `boot_permute()` stores a copy of the data inside the function that it generates?

__[A]{.solved}__: Because it doesn't actually store a copy; it's just a name that points to the same underlying object in memory.

```{r}
boot_permute <- function(df, var) {
  n <- nrow(df)
  force(var)
  
  function() {
    col <- df[[var]]
    col[sample(n, replace = TRUE)]
  }
}
boot_mtcars1 <- boot_permute(mtcars, "mpg")

lobstr::obj_size(mtcars)
lobstr::obj_size(boot_mtcars1)
lobstr::obj_sizes(mtcars, boot_mtcars1)
```

__[Q4]{.Q}__: How much time does `ll_poisson2()` save compared to `ll_poisson1()`? Use `bench::mark()` to see how much faster the optimisation occurs. How does changing the length of `x` change the results?

__[A]{.solved}__: Let us recall the definitions of `ll_poisson1()`, `ll_poisson2()` and the test data `x1`:

```{r}
ll_poisson1 <- function(x) {
  n <- length(x)
  
  function(lambda) {
    log(lambda) * sum(x) - n * lambda - sum(lfactorial(x))
  }
}

ll_poisson2 <- function(x) {
  n <- length(x)
  sum_x <- sum(x)
  c <- sum(lfactorial(x))
  
  function(lambda) {
    log(lambda) * sum_x - n * lambda - c
  }
}

x1 <- c(41, 30, 31, 38, 29, 24, 30, 29, 31, 38)
```

A benchmark on `x1` reveals a performance improvement of factor 2 for `ll_poisson2()` over `ll_poisson1()`:

```{r}
bench::mark(
  llp1 = optimise(ll_poisson1(x1), c(0, 100), maximum = TRUE),
  llp2 = optimise(ll_poisson2(x1), c(0, 100), maximum = TRUE)
)
```

As the redundant calculations within `ll_poisson1()` become more expensive with growing length of `x1`, we expect even further relative performance improvements for `ll_poisson2()`. The following benchmark reveals a relative performance improvement of factor 20 for `ll_poisson2()` when `x1` is of length 100000:

```{r, message = FALSE, warning = FALSE}
bench_poisson <- function(x_length){
  x <- rpois(x_length, 100L)
  
  rel_advantage <- bench::mark(
    llp1 = optimise(ll_poisson1(x), c(0, 100), maximum = TRUE),
    llp2 = optimise(ll_poisson2(x), c(0, 100), maximum = TRUE),
    relative = TRUE)$median %>% `[`(1)
  
  rel_advantage
}

bench_df <- tibble(
  x_length = 10^(1:5),
  relative_performance = map_dbl(x_length, bench_poisson)
)

bench_df %>% 
  ggplot(aes(x_length, relative_performance)) +
  geom_point() +
  geom_line() +
  ggtitle(
    "Relative performance of ll_poisson2() over \nll_poision1() increases with length of x"
  ) +
  scale_x_log10()
```


## Function factories + functionals
__[Q1]{.Q}__: Which of the following commands is equivalent to `with(x, f(z))`?

(a) `x$f(x$z)`.
(b) `f(x$z)`.
(c) `x$f(z)`.
(d) `f(z)`.
(e) It depends.

__[A]{.solved}__: (e) "It depends" is the correct answer. Usually `with()` is used with a data frame, so you'd usually expect (b), but if `x` is a list, it could be any of the options.

```{r}
f <- mean
z <- 1
x <- list(f = mean, z = 1)

identical(with(x, f(z)), x$f(x$z))
identical(with(x, f(z)), f(x$z))
identical(with(x, f(z)), x$f(z))
identical(with(x, f(z)), f(z))
```

__[Q2]{.Q}__: Compare and contrast the effects of `env_bind()` vs. `attach()` for the following code.

```{r}
funs <- list(
  mean = function(x) mean(x, na.rm = TRUE),
  sum = function(x) sum(x, na.rm = TRUE)
)

attach(funs)
mean <- function(x) stop("Hi!")
detach(funs)

env_bind(globalenv(), !!!funs)
mean <- function(x) stop("Hi!") 
env_unbind(globalenv(), names(funs))
```

__[A]{.solved}__: `attach()` adds `funs` to the search path. Therefore, the provided functions are found before their respective versions from the base package. Further, they can not get accidently overwritten by similar named functions in the global environment. One annoying downsinde of using `attach()` is the possibility to attach the same object multiple times, making it necessary to call `detach()` equally often.

```{r}
attach(funs)
attach(funs)

head(search())
detach(funs)
detach(funs)
```

In contrast `rlang::env_bind()` just adds the functions in `fun` to the global environment. No further side effects are introduced and the functions are overwritten when similarly named functions are defined.

```{r}
env_bind(globalenv(), !!!funs)
head(search())
```
