```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE
)
```
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE
)
```
# Expressions

## Structure of expressions

1.  __<span style="color:red">Q</span>__: There's no existing base function that checks if an element is
    a valid component of an expression (i.e., it's a constant, name,
    call, or pairlist). Implement one by guessing the names of the "is"
    functions for calls, names, and pairlists.
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    is.expression_component <- function(x) {
      (is.atomic(x) & length(x) == 1) | is.call(x) | is.name(x) | is.pairlist(x)
    }
    ```

2.  __<span style="color:red">Q</span>__: `pryr::ast()` uses non-standard evaluation. What's its escape hatch to
    standard evaluation?
    
    __<span style="color:green">A</span>__: You can call `pryr::call_tree` directly.

3.  __<span style="color:red">Q</span>__: What does the call tree of an if statement with multiple else conditions
    look like?
    
    __<span style="color:green">A</span>__: It depends a little bit how it is written.
    Here the infix version:
    
    ```{r}
    pryr::ast(`if`(FALSE, "first",
                   `if`(TRUE, "second",
                        `if`(TRUE, "third", "fourth"))))
    ```
    
    And here the "normal" version:
    
    ```{r}
    pryr::ast(if (FALSE) {
      "first"
      } else (
        if (TRUE) {
          "second"
          } else (
            if (TRUE) {
              "third"
              } else (
                "fourth"
                )
            )
        ))
    ```
    
    However, under the hood the language will call another base `if` statement. So `else if` seems to be for human readibility.

4.  __<span style="color:red">Q</span>__: Compare `ast(x + y %+% z)` to `ast(x ^ y %+% z)`. What do they tell you about the precedence of custom infix functions?
    
    __<span style="color:green">A</span>__: Comparison of the syntax trees:
    
    ```{r}
    # for ast(x + y %+% z)
    # y %+% z will be calculated first and the result will be added to x
    pryr::ast(x + y %+% z)
    
    # for ast(x ^ y %+% z)
    # x^y will be calculated first, and the result will be used as first argument of `%+%()`
    pryr::ast(x ^ y %+% z)
    ```
    
    So we can conclude that custom infix functions must have a precedence between addition and exponentiation. The general precedence rules can be found for example [here](https://cran.r-project.org/doc/manuals/r-release/R-lang.html).

5.  __<span style="color:red">Q</span>__: Why can't an expression contain an atomic vector of length greater than one?
    Which one of the six types of atomic vector can't appear in an expression?
    Why?
    
    __<span style="color:green">A</span>__: Because you can't type an expression that evaluates to an atomic of greater length than one without using a function (in particular, the function `c`), which means that these expressions would be calls, which are not part of expressions.
    
    Also raws can't appear in expressions, because of a similar reason. We think they are impossible to construct without using `as.raw`, which would mean that we will also end up with a call.
    
    Note also that imaginary parts of complex numbers work:

    ```{r, eval = TRUE}
    pryr::ast(1i)
    ```
  
    Doubles work, but get coerced to integers when possible:

    ```{r, eval = TRUE}
    pryr::ast(1.1)
    ```

    ```{r, eval = TRUE}
    pryr::ast(1.0)
    ```

## Names

1.  __<span style="color:red">Q</span>__: You can use `formals()` to both get and set the arguments of a function.
    Use `formals()` to modify the following function so that the default value
    of `x` is missing and `y` is 10.

    ```{r,collapse = TRUE}
    g <- function(x = 20, y) {
      x + y
    }
    ```
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    formals(g) <- alist(x = , y = 10)
    ```
    
    Similarly one can change the body of the function through `body<-()` and also the
    environment via `environment<-()`.

2.  __<span style="color:red">Q</span>__: Write an equivalent to `get()` using `as.name()` and `eval()`. Write an
    equivalent to `assign()` using `as.name()`, `substitute()`, and `eval()`.
    (Don't worry about the multiple ways of choosing an environment; assume
    that the user supplies it explicitly.)
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    get3 <- function(x, env = parent.frame()) {
      eval(as.name(x), env)
    }
    
    assign3 <- function(x, value, env = parent.frame()) {
      eval(substitute(x <- value,list(x = as.name(x), value = value)), env)
      if (length(x) > 1) {
        warning("Only the first element is used as variable name.")
      }
    }
    ```
    
## Calls

1.  __<span style="color:red">Q</span>__: The following two calls look the same, but are actually different:

    ```{r, eval = TRUE}
    (a <- call("mean", 1:10))
    (b <- call("mean", quote(1:10)))
    identical(a, b)
    ```

    What's the difference? Which one should you prefer?

    __<span style="color:orange">A</span>__: `call` evalulates its `...` arguments. So in the first call `1:10` will be evaluated to an integer (1, 2, 3, ..., 10) and in the second call `quote()` compensates the effect of the evaluation, so that `b`'s second element will be the expression `1:10` (which is again a call):
    
    ```{r, eval = TRUE}
    as.list(a)
    as.list(b)
    ```
    
    We can create an example, where we can see the consequences directly:
    
    ```{r, eval = TRUE}
    # h <- call("mean", z)
    ##> Error: object 'z' not found
    
    h <- call("mean", quote(z))
    # eval(h)
    ##> Error in mean(z) : object 'z' not found
    z <- 1:10
    eval(h)
    ```
    
    From my intuition I would prefer the second version, since it behaves more like lazy evaluation, but I am not sure if this is the correct answer as indended by the exercise.
    
2.  __<span style="color:red">Q</span>__: Implement a pure R version of `do.call()`.
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    do.call2 <- function(what, args, quote = FALSE, env = parent.frame()){
      if(!is.list(args))
        stop("second argument must be a list")
      if (quote) 
        args <- lapply(args, enquote)
      eval(as.call(c(what, args)), env)
    }
    ```

3.  __<span style="color:red">Q</span>__: Concatenating a call and an expression with `c()` creates a list. Implement
    `concat()` so that the following code works to combine a call and
    an additional argument.

    ```{r,collapse = TRUE, eval = FALSE}
    concat(quote(f), a = 1, b = quote(mean(a)))
    #> f(a = 1, b = mean(a))
    ```
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    concat <- function(f, ...){
      as.call(c(f, list(...)))
    }
    concat(quote(f), a = 1, b = quote(mean(a)))
    ```

4.  __<span style="color:red">Q</span>__: Since `list()`s don't belong in expressions, we could create a more
    convenient call constructor that automatically combines lists into the
    arguments. Implement `make_call()` so that the following code works.

    ```{r, eval = FALSE}
    make_call(quote(mean), list(quote(x), na.rm = TRUE))
    #> mean(x, na.rm = TRUE)
    make_call(quote(mean), quote(x), na.rm = TRUE)
    #> mean(x, na.rm = TRUE)
    ```
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    make_call <- function(x, ...){
      as.call(c(x, ...))
    }
    
    make_call(quote(mean), list(quote(x), na.rm = TRUE))
    make_call(quote(mean), quote(x), na.rm = TRUE)
    ```

5.  __<span style="color:red">Q</span>__: How does `mode<-` work? How does it use `call()`?
    
    __<span style="color:green">A</span>__: We can explain it best, when we comment the source code:
    
    ```{r, eval = FALSE}
    function (x) 
      {
      # when x is an expression, mode will(x) return "expression"
      if (is.expression(x)) 
        return("expression")
      # when x is a call (or language, which is exactly the same), the first element 
      # of the call will be coerced to character. 
      # If the call is an autoprinting (like in quote((1))), mode will return "(".
      # For any other call, mode will return "call" 
      if (is.call(x)) 
        return(switch(deparse(x[[1L]])[1L], `(` = "(", "call"))
      # if x is a name (or a symbol, which is exactly the same), then mode will return "name"
      if (is.name(x)) 
        "name"
      # otherwise, mode will return dependent on typeof(x). If typeof(x) is double or integer, 
      # mode will return "numeric". If typeof(x) is closure, builtin or special, mode(x) will
      # return "function". And in all other cases, mode will just return typeof(x)
      else switch(tx <- typeof(x), double = , integer = "numeric", 
                  closure = , builtin = , special = "function", tx)
      }
    <bytecode: 0x000000000c4e66e0>
      <environment: namespace:base>
    ```
      
    As commented above, `mode()` uses `is.call()` to distinguish autoprint- and "normal" calls
with the help of a separate `switch()`.

6.  __<span style="color:red">Q</span>__: Read the source for `pryr::standardise_call()`. How does it work?
    Why is `is.primitive()` needed?
    
    __<span style="color:green">A</span>__: It evaluates the first element of the call,
    which is usually the name of a function, but can also be another call. Then is uses `match.call()` to get the standard names for all the arguments.
    
    `is.primitive()` is used as an escape to just return the call instead of using `match.call()` if the function passed is a primitive. This is done because `match.call()` does not work for primitives.

7.  __<span style="color:red">Q</span>__: `standardise_call()` doesn't work so well for the following calls.
    Why?

    ```{r}
    library(pryr)
    standardise_call(quote(mean(1:10, na.rm = TRUE)))
    standardise_call(quote(mean(n = T, 1:10)))
    standardise_call(quote(mean(x = 1:10, , TRUE)))
    ```
    
    __<span style="color:green">A</span>__: 

8.  __<span style="color:red">Q</span>__: Read the documentation for `pryr::modify_call()`. How do you think
    it works? Read the source code.
    
    __<span style="color:green">A</span>__: Again, we explain by commenting the source
    
    ```{r, eval = FALSE}
    function (call, new_args) {
      # check if call is a call and new_args is a list
      stopifnot(is.call(call), is.list(new_args))
      # standardise the call
      call <- standardise_call(call)
      # check if the supplied new_args list has any unnamed elements.
      # if so, an error occurs.
      nms <- names(new_args) %||% rep("", length(new_args))
      if (any(nms == "")) {
        stop("All new arguments must be named", call. = FALSE)
        }
      # every name element of the call, for which a new argument was supplied by the user,
      # becomes overwritten
      for (nm in nms) {
        call[[nm]] <- new_args[[nm]]
        }
      # finally the modified call is returned
      call
    }
    <environment: namespace:pryr>
    ```

9.  __<span style="color:red">Q</span>__: Use `ast()` and experimentation to figure out the three arguments in an
    `if()` call. Which components are required? What are the arguments to
    the `for()` and `while()` calls?
    
    __<span style="color:green">A</span>__: 
    
    **if:**
    
    ```{r, eval = TRUE}
    ## All these return an error
    # pryr::ast(if)
    # pryr::ast(if())
    # pryr::ast(if{})
    # pryr::ast(if(){})
    # pryr::ast(if(TRUE))
    
    ## This is the minimum required
    pryr::ast(if(TRUE){1})
    
    ## One can also supply an alternative expression
    pryr::ast(if(TRUE){}
              else {3})
    
    ## However, one has to supply the compound expression (the first one),
    ## otherwise we get an error
    # pryr::ast(if(TRUE)
    #           else {3})

    # So this is how if basically works
    pryr::ast(if(cond)expr)
    
    # and here within a call
    eval(call("if", TRUE, 1))
    eval(call("if", TRUE, 1, 2))
    eval(call("if", FALSE, 1, 2))
    ```
    
    **for:**
    
    ```{r, eval = TRUE}
    ## All these return an error
    # pryr::ast(for)
    # pryr::ast(for{})
    # pryr::ast(for())
    # pryr::ast(for(){})
    # pryr::ast(for(in){})
    # pryr::ast(for(var in){})
    # pryr::ast(for(var in 10))
    # pryr::ast(for(in 10){})
    
    ## This is the minimum required
    pryr::ast(for(var in 10){})

    ## So this is how for basically works
    pryr::ast(for(var in seq)expr)

    ## And here within a call (note that we need quote, since var has to be a nanme
    ## and expr has to be an expression)
    eval(call("for", var = quote(i), seq = 1:3, expr = quote(print(i))))
    
    ## as infix function it looks a little bit easier
    `for`(i, 1:3, print(i))
    ```
    
    **while:**
    
    ```{r, eval = TRUE}
    ## All these return an error
    # pryr::ast(while)
    # pryr::ast(while())
    # pryr::ast(while(TRUE))
    # pryr::ast(while(){})
    # pryr::ast(while()1)
    # pryr::ast(while(TRUE){})

    ## This is the minimum required
    pryr::ast(while(TRUE)1)

    ## So this is how while basically works
    pryr::ast(while(cond)expr)
    
    ## And here within a call (infinite loop in this case)
    # eval(call("while", TRUE , 1))
    ```

## Capturing the current call

1.  __<span style="color:red">Q</span>__: Compare and contrast `update_model()` with `update.default()`.
    
    __<span style="color:green">A</span>__: 

2.  __<span style="color:red">Q</span>__: Why doesn't `write.csv(mtcars, "mtcars.csv", row = FALSE)` work?
    What property of argument matching has the original author forgotten?
    
    __<span style="color:green">A</span>__: 

3.  __<span style="color:red">Q</span>__: Rewrite `update.formula()` to use R code instead of C code.
    
    __<span style="color:green">A</span>__: 

4.  __<span style="color:red">Q</span>__: Sometimes it's necessary to uncover the function that called the
    function that called the current function (i.e., the grandparent, not
    the parent). How can you use `sys.call()` or `match.call()` to find
    this function?
    
    __<span style="color:green">A</span>__: 

## Pairlists

1.  __<span style="color:red">Q</span>__: How are `alist(a)` and `alist(a = )` different? Think about both the
    input and the output.
    
    __<span style="color:green">A</span>__: 

2.  __<span style="color:red">Q</span>__: Read the documentation and source code for `pryr::partial()`. What does it
    do? How does it work? Read the documentation and source code for
    `pryr::unenclose()`. What does it do and how does it work?
    
    __<span style="color:green">A</span>__: 

3.  __<span style="color:red">Q</span>__: The actual implementation of `curve()` looks more like

    ```{r curve3}
    curve3 <- function(expr, xlim = c(0, 1), n = 100,
                       env = parent.frame()) {
      env2 <- new.env(parent = env)
      env2$x <- seq(xlim[1], xlim[2], length = n)

      y <- eval(substitute(expr), env2)
      plot(env2$x, y, type = "l", 
        ylab = deparse(substitute(expr)))
    }
    ```

    How does this approach differ from `curve2()` defined above?
    
    __<span style="color:green">A</span>__: 

## Parsing and deparsing

1.  __<span style="color:red">Q</span>__: What are the differences between `quote()` and `expression()`?
    
    __<span style="color:green">A</span>__: 

2.  __<span style="color:red">Q</span>__: Read the help for `deparse()` and construct a call that `deparse()`
    and `parse()` do not operate symmetrically on.
    
    __<span style="color:green">A</span>__: 

3.  __<span style="color:red">Q</span>__: Compare and contrast `source()` and `sys.source()`.
    
    __<span style="color:green">A</span>__: 

4.  __<span style="color:red">Q</span>__: Modify `simple_source()` so it returns the result of _every_ expression,
    not just the last one.
    
    __<span style="color:green">A</span>__: 

5.  __<span style="color:red">Q</span>__: The code generated by `simple_source()` lacks source references. Read
    the source code for `sys.source()` and the help for `srcfilecopy()`,
    then modify `simple_source()` to preserve source references. You can
    test your code by sourcing a function that contains a comment. If
    successful, when you look at the function, you'll see the comment and
    not just the source code.
    
    __<span style="color:green">A</span>__: 

## Walking the AST with recursive functions

1.  __<span style="color:red">Q</span>__: Why does `logical_abbr()` use a for loop instead of a functional
    like `lapply()`?
    
    __<span style="color:green">A</span>__: 

2.  __<span style="color:red">Q</span>__: `logical_abbr()` works when given quoted objects, but doesn't work when
    given an existing function, as in the example below. Why not? How could
    you modify `logical_abbr()` to work with functions? Think about what
    components make up a function.

    ```{r, eval = FALSE}
    f <- function(x = TRUE) {
      g(x + T)
    }
    logical_abbr(f)
    ```
    
    __<span style="color:green">A</span>__: 

3.  __<span style="color:red">Q</span>__: Write a function called `ast_type()` that returns either "constant",
    "name", "call", or "pairlist". Rewrite `logical_abbr()`, `find_assign()`,
    and `bquote2()` to use this function with `switch()` instead of nested if
    statements.
    
    __<span style="color:green">A</span>__: 

4.  __<span style="color:red">Q</span>__: Write a function that extracts all calls to a function. Compare your
    function to `pryr::fun_calls()`.
    
    __<span style="color:green">A</span>__: 

5.  __<span style="color:red">Q</span>__: Write a wrapper around `bquote2()` that does non-standard evaluation
    so that you don't need to explicitly `quote()` the input.
    
    __<span style="color:green">A</span>__: 

6.  __<span style="color:red">Q</span>__: Compare `bquote2()` to `bquote()`. There is a subtle bug in `bquote()`:
    it won't replace calls to functions with no arguments. Why?

    ```{r}
    bquote(.(x)(), list(x = quote(f)))
    bquote(.(x)(1), list(x = quote(f)))
    ```
    
    __<span style="color:green">A</span>__: 

7.  __<span style="color:red">Q</span>__: Improve the base `recurse_call()` template to also work with lists of
    functions and expressions (e.g., as from `parse(path_to_file))`.
    
    __<span style="color:green">A</span>__: 
