```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE
)
```
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE
)
```
# Expressions

## Structure of expressions

1.  __<span style="color:red">Q</span>__: There's no existing base function that checks if an element is
    a valid component of an expression (i.e., it's a constant, name,
    call, or pairlist). Implement one by guessing the names of the "is"
    functions for calls, names, and pairlists.
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    is.expression_component <- function(x) {
      (is.atomic(x) & length(x) == 1) | is.call(x) | is.name(x) | is.pairlist(x)
    }
    ```

2.  __<span style="color:red">Q</span>__: `pryr::ast()` uses non-standard evaluation. What's its escape hatch to
    standard evaluation?
    
    __<span style="color:green">A</span>__: You can call `pryr::call_tree` directly.

3.  __<span style="color:red">Q</span>__: What does the call tree of an if statement with multiple else conditions
    look like?
    
    __<span style="color:green">A</span>__: It depends a little bit how it is written.
    Here the infix version:
    
    ```{r}
    pryr::ast(`if`(FALSE, "first",
                   `if`(TRUE, "second",
                        `if`(TRUE, "third", "fourth"))))
    ```
    
    And here the "normal" version:
    
    ```{r}
    pryr::ast(if (FALSE) {
      "first"
      } else (
        if (TRUE) {
          "second"
          } else (
            if (TRUE) {
              "third"
              } else (
                "fourth"
                )
            )
        ))
    ```
    
    However, under the hood the language will call another base `if` statement. So `else if` seems to be for human readibility.

4.  __<span style="color:red">Q</span>__: Compare `ast(x + y %+% z)` to `ast(x ^ y %+% z)`. What do they tell you about the precedence of custom infix functions?
    
    __<span style="color:green">A</span>__: Comparison of the syntax trees:
    
    ```{r}
    # for ast(x + y %+% z)
    # y %+% z will be calculated first and the result will be added to x
    pryr::ast(x + y %+% z)
    
    # for ast(x ^ y %+% z)
    # x^y will be calculated first, and the result will be used as first argument of `%+%()`
    pryr::ast(x ^ y %+% z)
    ```
    
    So we can conclude that custom infix functions must have a precedence between addition and exponentiation. The general precedence rules can be found for example [here](https://cran.r-project.org/doc/manuals/r-release/R-lang.html).

5.  __<span style="color:red">Q</span>__: Why can't an expression contain an atomic vector of length greater than one?
    Which one of the six types of atomic vector can't appear in an expression?
    Why?
    
    __<span style="color:green">A</span>__: Because you can't type an expression that evaluates to an atomic of greater length than one without using a function (in particular, the function `c`), which means that these expressions would be calls, which are not part of expressions.
    
    Also raws can't appear in expressions, because of a similar reason. We think they are impossible to construct without using `as.raw`, which would mean that we will also end up with a call.
    
    Note also that imaginary parts of complex numbers work:

    ```{r, eval = TRUE}
    pryr::ast(1i)
    ```
  
    Doubles work, but get coerced to integers when possible:

    ```{r, eval = TRUE}
    pryr::ast(1.1)
    ```

    ```{r, eval = TRUE}
    pryr::ast(1.0)
    ```

## Names

1.  __<span style="color:red">Q</span>__: You can use `formals()` to both get and set the arguments of a function.
    Use `formals()` to modify the following function so that the default value
    of `x` is missing and `y` is 10.

    ```{r,collapse = TRUE}
    g <- function(x = 20, y) {
      x + y
    }
    ```
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    formals(g) <- alist(x = , y = 10)
    ```
    
    Similarly one can change the body of the function through `body<-()` and also the
    environment via `environment<-()`.

2.  __<span style="color:red">Q</span>__: Write an equivalent to `get()` using `as.name()` and `eval()`. Write an
    equivalent to `assign()` using `as.name()`, `substitute()`, and `eval()`.
    (Don't worry about the multiple ways of choosing an environment; assume
    that the user supplies it explicitly.)
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    get3 <- function(x, env = parent.frame()) {
      eval(as.name(x), env)
    }
    
    assign3 <- function(x, value, env = parent.frame()) {
      eval(substitute(x <- value,list(x = as.name(x), value = value)), env)
      if (length(x) > 1) {
        warning("Only the first element is used as variable name.")
      }
    }
    ```
    
## Calls

1.  __<span style="color:red">Q</span>__: The following two calls look the same, but are actually different:

    ```{r, eval = TRUE}
    (a <- call("mean", 1:10))
    (b <- call("mean", quote(1:10)))
    identical(a, b)
    ```

    What's the difference? Which one should you prefer?

    __<span style="color:orange">A</span>__: `call` evalulates its `...` arguments. So in the first call `1:10` will be evaluated to an integer (1, 2, 3, ..., 10) and in the second call `quote()` compensates the effect of the evaluation, so that `b`'s second element will be the expression `1:10` (which is again a call):
    
    ```{r, eval = TRUE}
    as.list(a)
    as.list(b)
    ```
    
    We can create an example, where we can see the consequences directly:
    
    ```{r, eval = TRUE}
    arg <- seq(10)
    call1 <- call("mean", arg)
    print(call1)
    call2 <- call("mean", quote(arg))
    print(call2)
    eval(call1)
    eval(call2)
    ```
    
    I would prefer the second version, since it behaves more like lazy evaluation. It's better to have call args depends on the calling environment rather than the enclosing environment,that's more similar to normal function behavior.
    
2.  __<span style="color:red">Q</span>__: Implement a pure R version of `do.call()`.
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    do.call2 <- function(what, args, quote = FALSE, env = parent.frame()) {
      if (!is.list(args)) {
        stop("second argument must be a list")
      }
      if (isTRUE(quote)) {
        args <- lapply(args, enquote)
      }
      eval(as.call(c(what, args)), env)
    }
    ```

3.  __<span style="color:red">Q</span>__: Concatenating a call and an expression with `c()` creates a list. Implement `concat()` so that the following code works to combine a call and an additional argument.

    ```{r,collapse = TRUE, eval = FALSE}
    concat(quote(f), a = 1, b = quote(mean(a)))
    #> f(a = 1, b = mean(a))
    ```
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    concat <- function(f, ...) {
      as.call(c(f, list(...)))
    }
    concat(quote(f), a = 1, b = quote(mean(a)))
    ```

4.  __<span style="color:red">Q</span>__: Since `list()`s don't belong in expressions, we could create a more convenient call constructor that automatically combines lists into the arguments. Implement `make_call()` so that the following code works.

    ```{r, eval = FALSE}
    make_call(quote(mean), list(quote(x), na.rm = TRUE))
    #> mean(x, na.rm = TRUE)
    make_call(quote(mean), quote(x), na.rm = TRUE)
    #> mean(x, na.rm = TRUE)
    ```
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    make_call <- function(x, ...) {
      as.call(c(x, ...))
    }
    
    make_call(quote(mean), list(quote(x), na.rm = TRUE))
    make_call(quote(mean), quote(x), na.rm = TRUE)
    ```

5.  __<span style="color:red">Q</span>__: How does `mode<-` work? How does it use `call()`?
    
    __<span style="color:green">A</span>__: We can explain it best, when we comment the source code:
    
    ```{r, eval = FALSE}
    function (x) {
      # when x is an expression, mode(x) will return "expression"
      if (is.expression(x)) 
        return("expression")
      # when x is a call (or language, which is exactly the same), the first element 
      # of the call will be coerced to character. 
      # If the call is an autoprinting (like in quote((1))), mode will return "(".
      # For any other call, mode will return "call" 
      if (is.call(x)) 
        return(switch(deparse(x[[1L]])[1L], `(` = "(", "call"))
      # if x is a name (or a symbol, which is exactly the same), then mode will return "name"
      if (is.name(x)) 
        "name"
      # otherwise, mode will return dependent on typeof(x). If typeof(x) is double or integer, 
      # mode will return "numeric". If typeof(x) is closure, builtin or special, mode(x) will
      # return "function". And in all other cases, mode will just return typeof(x)
      else switch(tx <- typeof(x), double = , integer = "numeric", 
                  closure = , builtin = , special = "function", tx)
      }
    <bytecode: 0x000000000c4e66e0>
      <environment: namespace:base>
    ```
      
    As commented above, `mode()` uses `is.call()` to distinguish autoprint- and "normal" calls
with the help of a separate `switch()`.

6.  __<span style="color:red">Q</span>__: Read the source for `pryr::standardise_call()`. How does it work?  Why is `is.primitive()` needed?
    
    __<span style="color:green">A</span>__: It evaluates the first element of the call, which is usually the name of a function, but can also be another call. Then is uses `match.call()` to get the standard names for all the arguments.
    
    `is.primitive()` is used as an escape to just return the call instead of using `match.call()` if the function passed is a primitive. This is done because `match.call()` does not work for primitives.

7.  __<span style="color:red">Q</span>__: `standardise_call()` doesn't work so well for the following calls. Why?

    ```{r}
    library(pryr)
    standardise_call(quote(mean(1:10, na.rm = TRUE)))
    standardise_call(quote(mean(n = T, 1:10)))
    standardise_call(quote(mean(x = 1:10, , TRUE)))
    ```
    
    __<span style="color:green">A</span>__: The reason these don't work is not that `mean` is a primitive (as seen in exercise 6) -- it's not -- but because `mean` uses S3 dispatch (i.e., `UseMethod`) and therefore does not store its formals on `mean`, but rather `mean.default`. For example, `pryr::standardize_call` can do much better when the S3 dispatch is explicit.

    For example, this works:

    ```{r}
    pryr::standardise_call(quote(mean.default(n = T, 1:10)))
    ```

8.  __<span style="color:red">Q</span>__: Read the documentation for `pryr::modify_call()`. How do you think it works? Read the source code.
    
    __<span style="color:green">A</span>__: Again, we explain by commenting the source.
    
    ```{r, eval = FALSE}
    function (call, new_args) {
      # check if call is a call and new_args is a list
      stopifnot(is.call(call), is.list(new_args))
      # standardise the call
      call <- standardise_call(call)
      # check if the supplied new_args list has any unnamed elements.
      # if so, an error occurs.
      nms <- names(new_args) %||% rep("", length(new_args))
      if (any(nms == "")) {
        stop("All new arguments must be named", call. = FALSE)
        }
      # every name element of the call, for which a new argument was supplied by the user,
      # becomes overwritten
      for (nm in nms) {
        call[[nm]] <- new_args[[nm]]
        }
      # finally the modified call is returned
      call
    }
    <environment: namespace:pryr>
    ```

9.  __<span style="color:red">Q</span>__: Use `ast()` and experimentation to figure out the three arguments in an `if()` call. Which components are required? What are the arguments to the `for()` and `while()` calls?
    
    __<span style="color:green">A</span>__: 
    
    **if:**
    
    ```{r, eval = TRUE}
    ## All these return an error
    # pryr::ast(if)
    # pryr::ast(if())
    # pryr::ast(if{})
    # pryr::ast(if(){})
    # pryr::ast(if(TRUE))
    
    ## This is the minimum required
    pryr::ast(if(TRUE){1})
    
    ## One can also supply an alternative expression
    pryr::ast(if(TRUE){}
              else {3})
    
    ## However, one has to supply the compound expression (the first one),
    ## otherwise we get an error
    # pryr::ast(if(TRUE)
    #           else {3})

    # So this is how if basically works
    pryr::ast(if(cond)expr)
    
    # and here within a call
    eval(call("if", TRUE, 1))
    eval(call("if", TRUE, 1, 2))
    eval(call("if", FALSE, 1, 2))
    ```
    
    **for:**
    
    ```{r, eval = TRUE}
    ## All these return an error
    # pryr::ast(for)
    # pryr::ast(for{})
    # pryr::ast(for())
    # pryr::ast(for(){})
    # pryr::ast(for(in){})
    # pryr::ast(for(var in){})
    # pryr::ast(for(var in 10))
    # pryr::ast(for(in 10){})
    
    ## This is the minimum required
    pryr::ast(for(var in 10){})

    ## So this is how for basically works
    pryr::ast(for(var in seq)expr)

    ## And here within a call (note that we need quote, since var has to be a nanme
    ## and expr has to be an expression)
    eval(call("for", var = quote(i), seq = 1:3, expr = quote(print(i))))
    
    ## as infix function it looks a little bit easier
    `for`(i, 1:3, print(i))
    ```
    
    **while:**
    
    ```{r, eval = TRUE}
    ## All these return an error
    # pryr::ast(while)
    # pryr::ast(while())
    # pryr::ast(while(TRUE))
    # pryr::ast(while(){})
    # pryr::ast(while()1)
    # pryr::ast(while(TRUE){})

    ## This is the minimum required
    pryr::ast(while(TRUE)1)

    ## So this is how while basically works
    pryr::ast(while(cond)expr)
    
    ## And here within a call (infinite loop in this case)
    # eval(call("while", TRUE , 1))
    ```

## Capturing the current call

1.  __<span style="color:red">Q</span>__: Compare and contrast `update_model()` with `update.default()`.
    
    __<span style="color:green">A</span>__: 

	```{r}
	update_call <- function (object, formula_, ...) {
		call <- object$call
		if (!missing(formula_)) {
			call$formula <- update.formula(formula(object), formula_)
		}
		pryr::modify_call(call, pryr::dots(...))
	}

	update_model <- function(object, formula_, ...) {
		call <- update_call(object, formula_, ...)
		eval(call, environment(formula(object)))
	}
	```

	`update_model` always evaluates the resulting call, whereas `update.default` can return the call without evaluation if `evaluate = FALSE`.

	`update.default` evaluates the call in the global environment, whereas `update_model` evaluates the call within the environment of the object passed.

	`update.default` handles some extras whereas `update_model` does not.

	`update_model`'s syntax is less verbose and easier to read.


2.  __<span style="color:red">Q</span>__: Why doesn't `write.csv(mtcars, "mtcars.csv", row = FALSE)` work?  What property of argument matching has the original author forgotten?
    
    __<span style="color:green">A</span>__: `write.csv` rewrites the call, which does not allow R to use it's natural `match.call` to expand `row` to `row.names`.

3.  __<span style="color:red">Q</span>__: Rewrite `update.formula()` to use R code instead of C code.

    __<span style="color:green">A</span>__: 

	```{r}
	modify_formula <- function(old, new) {
    old_lhs <- old[[2]]
    old_rhs <- old[[3]]
    if (length(new) == 2) {
      new_lhs <- old_lhs
      new_rhs <- new[[2]]
    } else {
      new_lhs <- new[[2]]
      new_rhs <- new[[3]]
    }
    new_lhs_ <- gsub(".", deparse(old_lhs), deparse(new_lhs), fixed = TRUE)
    new_rhs_ <- gsub(".", deparse(old_rhs), deparse(new_rhs), fixed = TRUE)
    as.formula(paste(new_lhs_, "~", new_rhs_))
	}
  modify_formula(y ~ x, ~ . + x2)
  update.formula(y ~ x, ~ . + x2)
  update.formula(y ~ x, log(.) ~ .)
  modify_formula(y ~ x, log(.) ~ .)
  update.formula(. ~ u+v, res ~ .)
  modify_formula(. ~ u+v, res ~ .)
  ```

4.  __<span style="color:red">Q</span>__: Sometimes it's necessary to uncover the function that called the function that called the current function (i.e., the grandparent, not the parent). How can you use `sys.call()` or `match.call()` to find this function?
    
    __<span style="color:green">A</span>__: You can use `sys.call(-2)`.

    ```{r}
    child_fn <- function() {
      message("I am the child_fn")
      message("My parent is ", sys.call(-1))
      message("My grandparent is ", sys.call(-2))
    }
    daddy_fn <- function() { child_fn() }
    grandpa_fn <- function() { daddy_fn() }
    grandpa_fn()
    ```

## Pairlists

1.  __<span style="color:red">Q</span>__: How are `alist(a)` and `alist(a = )` different? Think about both the input and the output.
    
    __<span style="color:green">A</span>__: `alist(a)` returns an unnamed list of length 1 with the first element being the name `a` (note that this refers to the `name` class, which is distinct from being named `a`), so it is unsuitable for use in a function. `alist(a = )` returns a named list with the first element having name `a` and the first element being empty.

2.  __<span style="color:red">Q</span>__: Read the documentation and source code for `pryr::partial()`. What does it do? How does it work? Read the documentation and source code for `pryr::unenclose()`. What does it do and how does it work?
    
    __<span style="color:green">A</span>__: `pryr::partial` takes a function and arguments and then constructs a call of that function with those args. `pryr::unenclose` takes a closure and substitutes the variables in that closure for its values found in its environment, which results in the explicit function.

3.  __<span style="color:red">Q</span>__: The actual implementation of `curve()` looks more like

    ```{r curve3}
    curve3 <- function(expr, xlim = c(0, 1), n = 100, env = parent.frame()) {
      env2 <- new.env(parent = env)
      env2$x <- seq(xlim[1], xlim[2], length = n)
      y <- eval(substitute(expr), env2)
      plot(env2$x, y, type = "l", ylab = deparse(substitute(expr)))
    }
    ```

    ```{r curve2}
    curve2 <- function(expr, xlim = c(0, 1), n = 100, env = parent.frame()) {
      f <- pryr::make_function(alist(x = ), substitute(expr), env)
      x <- seq(xlim[1], xlim[2], length = n)
      y <- f(x)
      plot(x, y, type = "l", ylab = deparse(substitute(expr)))
    }

    ```

    How does this approach differ from `curve2()` defined above?
    
    __<span style="color:green">A</span>__: `curve2` uses pryr::make_function instead of creating an env and evaluating within it.

## Parsing and deparsing

1.  __<span style="color:red">Q</span>__: What are the differences between `quote()` and `expression()`?
    
    __<span style="color:green">A</span>__: An expression is a list of calls, whereas a quote is a single call. See:

    ```{r}
    as.list(expression( 2*3))
    as.list(quote(2*3))
    ```

2.  __<span style="color:red">Q</span>__: Read the help for `deparse()` and construct a call that `deparse()` and `parse()` do not operate symmetrically on.
    
    __<span style="color:green">A</span>__: `parse` and `deparse` handle length > 1 vectors differently.
    
    ```{r}
    parse(text = c(1, 2, 3))
    deparse(c(1, 2, 3))
    ```

3.  __<span style="color:red">Q</span>__: Compare and contrast `source()` and `sys.source()`.
    
    __<span style="color:green">A</span>__: `source` is a standardGeneric created from the base package, whereas `sys.source` is a function exported from base. `source` has many more options than `sys.source`.  `source` can accept data from connections other than files, whereas `sys.source` cannot.  `sys.source` which is a streamlined version to source a file into an environment.

4.  __<span style="color:red">Q</span>__: Modify `simple_source()` so it returns the result of _every_ expression, not just the last one.
    
    __<span style="color:green">A</span>__: 

		```{r}
		simple_source <- function(file, envir = new.env()) {
			stopifnot(file.exists(file))
			stopifnot(is.environment(envir))
			lines <- readLines(file, warn = FALSE)
			exprs <- parse(text = lines)
			if (length(exprs) == 0L) return(invisible())
			invisible(lapply(exprs, function(expr) { eval(expr, envir) }))
		}
		```

5.  __<span style="color:red">Q</span>__: The code generated by `simple_source()` lacks source references. Read the source code for `sys.source()` and the help for `srcfilecopy()`, then modify `simple_source()` to preserve source references. You can test your code by sourcing a function that contains a comment. If successful, when you look at the function, you'll see the comment and not just the source code.
    
    __<span style="color:green">A</span>__: 

		```{r}
		simple_source <- function(file, envir = new.env()) {
			stopifnot(file.exists(file))
			stopifnot(is.environment(envir))
			lines <- readLines(file, warn = FALSE)
			srcfile <- srcfilecopy(file, lines, file.mtime(file), isFile = TRUE)
			exprs <- parse(text = lines, srcfile = srcfile, keep.source = TRUE)
			if (length(exprs) == 0L) return(invisible())
			invisible(lapply(exprs, function(expr) { eval(expr, envir) }))
		}
		```

## Walking the AST with recursive functions

1.  __<span style="color:red">Q</span>__: Why does `logical_abbr()` use a for loop instead of a functional
    like `lapply()`?
    
    __<span style="color:green">A</span>__: 

2.  __<span style="color:red">Q</span>__: `logical_abbr()` works when given quoted objects, but doesn't work when
    given an existing function, as in the example below. Why not? How could
    you modify `logical_abbr()` to work with functions? Think about what
    components make up a function.

    ```{r, eval = FALSE}
    f <- function(x = TRUE) {
      g(x + T)
    }
    logical_abbr(f)
    ```
    
    __<span style="color:green">A</span>__: 

3.  __<span style="color:red">Q</span>__: Write a function called `ast_type()` that returns either "constant",
    "name", "call", or "pairlist". Rewrite `logical_abbr()`, `find_assign()`,
    and `bquote2()` to use this function with `switch()` instead of nested if
    statements.
    
    __<span style="color:green">A</span>__: 

4.  __<span style="color:red">Q</span>__: Write a function that extracts all calls to a function. Compare your
    function to `pryr::fun_calls()`.
    
    __<span style="color:green">A</span>__: 

5.  __<span style="color:red">Q</span>__: Write a wrapper around `bquote2()` that does non-standard evaluation
    so that you don't need to explicitly `quote()` the input.
    
    __<span style="color:green">A</span>__: 

6.  __<span style="color:red">Q</span>__: Compare `bquote2()` to `bquote()`. There is a subtle bug in `bquote()`:
    it won't replace calls to functions with no arguments. Why?

    ```{r}
    bquote(.(x)(), list(x = quote(f)))
    bquote(.(x)(1), list(x = quote(f)))
    ```
    
    __<span style="color:green">A</span>__: 

7.  __<span style="color:red">Q</span>__: Improve the base `recurse_call()` template to also work with lists of
    functions and expressions (e.g., as from `parse(path_to_file))`.
    
    __<span style="color:green">A</span>__: 
