# Performance

## Microbenchmarking

1. __<span style="color:red">Q</span>__: Instead of using `microbenchmark()`, you could use the built-in function
   `system.time()`. But `system.time()` is much less precise, so you'll
   need to repeat each operation many times with a loop, and then divide
   to find the average time of each operation, as in the code below.

    ```{r, eval = FALSE}
    n <- 1:1e6
    system.time(for (i in n) sqrt(x)) / length(n)
    system.time(for (i in n) x ^ 0.5) / length(n)
    ```
    
    How do the estimates from `system.time()` compare to those from
    `microbenchmark()`? Why are they different?

1.  __<span style="color:red">Q</span>__: Here are two other ways to compute the square root of a vector. Which
    do you think will be fastest? Which will be slowest? Use microbenchmarking
    to test your answers.

    ```{r, eval = FALSE}
    x ^ (1 / 2)
    exp(log(x) / 2)
    ```
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    x <- runif(100)
    microbenchmark::microbenchmark(
      sqrt(x),
      x ^ 0.5,
      x ^ (1 / 2),
      exp(log(x) / 2)
    )
    ```

1.  __<span style="color:red">Q</span>__: Use microbenchmarking to rank the basic arithmetic operators (`+`, `-`,
    `*`, `/`, and `^`) in terms of their speed. Visualise the results. Compare
    the speed of arithmetic on integers vs. doubles.

1.  __<span style="color:red">Q</span>__: You can change the units in which the microbenchmark results are
    expressed with the `unit` parameter. Use `unit = "eps"` to show
    the number of evaluations needed to take 1 second. Repeat the benchmarks
    above with the eps unit. How does this change your intuition for performance?

## Language performance

1.  __<span style="color:red">Q</span>__: `scan()` has the most arguments (21) of any base function. About how
    much time does it take to make 21 promises each time scan is called?
    Given a simple input (e.g., `scan(text = "1 2 3", quiet = T)`) what
    proportion of the total run time is due to creating those promises?

1.  __<span style="color:red">Q</span>__: Read ["Evaluating the Design of the R Language"](http://r.cs.purdue.edu/pub/ecoop12.pdf). What other aspects of the R-language slow it
    down? Construct microbenchmarks to illustrate. 

1.  __<span style="color:red">Q</span>__: How does the performance of S3 method dispatch change with the length
    of the class vector? How does performance of S4 method dispatch change
    with number of superclasses? How about RC?

1.  __<span style="color:red">Q</span>__: What is the cost of multiple inheritance and multiple dispatch on
    S4 method dispatch?

1.  __<span style="color:red">Q</span>__: Why is the cost of name lookup less for functions in the base package?

## Implementations performance

1.  __<span style="color:red">Q</span>__: The performance characteristics of `squish_ife()`, `squish_p()`, and
   `squish_in_place()` vary considerably with the size of `x`. Explore the
   differences. Which sizes lead to the biggest and smallest differences?

1.  __<span style="color:red">Q</span>__: Compare the performance costs of extracting an element from a list, a
    column from a matrix, and a column from a data frame. Do the same for rows.