# Profiling

## Has somebody already solved the problem?

1.  __<span style="color:red">Q</span>__: What are faster alternatives to `lm`? Which are specifically designed 
    to work with larger datasets?
        
    __<span style="color:green">A</span>__: Within the [Cran task view](https://cran.rstudio.com/web/views/) for [HighPerformanceComputing](https://cran.rstudio.com/web/views/HighPerformanceComputing.html) we can find for example the `speedglm` package and it's `speedlm()` function. We might not gain any performanceimprovements on small datasets:
    
    ```{r}
    stopifnot(all.equal(
      coef(speedglm::speedlm(Sepal.Length ~ Sepal.Width + Species, data = iris)),
      coef(lm(Sepal.Length ~ Sepal.Width + Species, data = iris))))

    microbenchmark::microbenchmark(
      speedglm::speedlm(Sepal.Length ~ Sepal.Width + Species, data = iris),
      lm(Sepal.Length ~ Sepal.Width + Species, data = iris)
    )
    ```
    
    However on bigger datasets it can make a difference:
    
    ```{r}
    eps <- rnorm(100000)
    x1 <- rnorm(100000, 5, 3)
    x2 <- rep(c("a", "b"), 50000)
    y <- 7 * x1 + (x2 == "a") + eps
    td <- data.frame(y = y, x1 = x1, x2 = x2, eps = eps)
    
    stopifnot(all.equal(
      coef(speedglm::speedlm(y ~ x1 + x2, data = td)),
      coef(lm(y ~ x1 + x2, data = td))))
    
    microbenchmark::microbenchmark(
      speedglm::speedlm(y ~ x1 + x2, data = td),
      lm(y ~ x1 + x2, data = td)
    )
    ```
    
    For further speedinprovements, you might consider switching your linear algebra libraries as stated in `?speedglm::speedlm`
    
    > The functions of class 'speedlm' may speed up the fitting of LMs to large data sets. High performances can be obtained especially if R is linked against an optimized BLAS, such as ATLAS.
    
    Note that there are many other opportunities mentioned in the task view, also some that make it possible to handle data which is not in memory. 
    
    When it comes to pure speed a quick google search on [r fastest lm](http://stackoverflow.com/questions/25416413/is-there-a-faster-lm-function) provides a stackoverflow thread where someone already solved this problem for us...

1.  __<span style="color:red">Q</span>__: What package implements a version of `match()` that's faster for
    repeated lookups? How much faster is it?
    
    __<span style="color:green">A</span>__: Again google gives a good recommendation for the searchterm [r faster match](http://stackoverflow.com/questions/32934933/faster-in-operator):
    
    ```{r}
    set.seed(1)
    table <- 1L:100000L
    x <- sample(table, 10000, replace=TRUE)
    
    stopifnot(all.equal(match(x, table), fastmatch::fmatch(x, table)))
    
    microbenchmark::microbenchmark(
      match(x, table),
      fastmatch::fmatch(x, table)
    )
    ```
    
    On my laptop `fastmatch::fmatch()` is around 25 times as fast as `match()`.

1.  __<span style="color:red">Q</span>__: List four functions (not just those in base R) that convert a string into a
    date time object. What are their strengths and weaknesses?
    
    __<span style="color:orange">A</span>__: At least these functions will do the trick: `as.POSIXct()`, `as.POSIXlt()`, `strftime()`, `strptime()`, `lubridate::ymd_hms()`. There might also be some in the timeseries packages `xts` or `zoo` and in `anytime`. An update on this will follow...

1.  __<span style="color:red">Q</span>__: How many different ways can you compute a 1d density estimate in R?
    
    __<span style="color:green">A</span>__: According to [Deng and Wickham (2011)](https://www.google.de/url?sa=t&rct=j&q=&esrc=s&source=web&cd=3&ved=0ahUKEwim0sTI9JLRAhVScFAKHdLuBOIQFggxMAI&url=http%3A%2F%2Fvita.had.co.nz%2Fpapers%2Fdensity-estimation.pdf&usg=AFQjCNFBdOT4DvSTXtGoawIFkgf6JlXV9Q&sig2=APnTeIL55Zj_0oATw15mBA&cad=rja) density estimation is implemented in over 20 R packages. 

1.  __<span style="color:red">Q</span>__: Which packages provide the ability to compute a rolling mean?
    
    __<span style="color:green">A</span>__: Again google [r rolling mean](http://stackoverflow.com/questions/743812/calculating-moving-average-in-r)
provides us with enough information and guides our attention on solutions in the following packages:
    * `zoo`
    ```{r}
    zoo::rollmean(1:10, 2, na.pad = TRUE, align = "left")
    zoo::rollapply(1:10, 2, mean, fill = NA, align = "left")
    ```
    * `TTR`
    ```{r}
    TTR::SMA(1:10, 2)
    ```
    * `RcppRoll`
    ```{r}
    RcppRoll::roll_mean(1:10, n = 2, fill = NA, align = "left")
    ```
    * `caTools`
    ```{r}
    caTools::runmean(1:10, k = 2, endrule = "NA", align = "left")
    ```

    Note that an exhaustive example on how to create a rolling mean function is provided in the [textbook](http://adv-r.had.co.nz/Functionals.html).

1.  __<span style="color:red">Q</span>__: What are the alternatives to `optim()`?
    
    __<span style="color:green">A</span>__: Depending on the usecase a lot of different options might be considered. For a general overview we would suggest the corresponding taskview on [Optimization](https://cran.r-project.org/web/views/Optimization.html).

## Do as little as possible

1.  __<span style="color:red">Q</span>__: How do the results change if you compare `mean()` and `mean.default()`
    on 10,000 observations, rather than on 100?
    
    __<span style="color:green">A</span>__: We start with 100 observations as shown in the textbook:
    
    ```{r}
    x <- runif(1e2)
    microbenchmark::microbenchmark(
      mean(x),
      mean.default(x)
    )
    ```
    
    In case of 10000 observations we can observe that using `mean.default()` preserves only a small advantage over the use of `mean()`:
    
    ```{r}
    x <- runif(1e4)
    microbenchmark::microbenchmark(
      mean(x),
      mean.default(x),
      unit = "ns"
    )
    ```
    
    When using even more observations - like in the next lines - it seems that `mean.default` doesn't preserve anymore any advantage at all:
    
    ```{r}
    x <- runif(1e6)
    microbenchmark::microbenchmark(
      mean(x),
      mean.default(x),
      unit = "ns"
    )
    ```

1.  __<span style="color:red">Q</span>__: The following code provides an alternative implementation of `rowSums()`.
    Why is it faster for this input?

    ```{r}
    rowSums2 <- function(df) {
      out <- df[[1L]]
      if (ncol(df) == 1) return(out)
      
      for (i in 2:ncol(df)) {
        out <- out + df[[i]]
      }
      out
    }
    
    df <- as.data.frame(
      replicate(1e3, sample(100, 1e4, replace = TRUE))
    )
    system.time(rowSums(df))
    system.time(rowSums2(df))
    ```
    
    __<span style="color:green">A</span>__: 

1.  __<span style="color:red">Q</span>__: What's the difference between `rowSums()` and `.rowSums()`?
    
    __<span style="color:green">A</span>__: `.rowSums()` is defined as
    
    ```{r}
    .rowSums
    ```
    
    this means, that the internal `rowSums()` function is called via `.Internal()`.
    
    > .Internal performs a call to an internal code which is built in to the R
    interpreter.
    
    The internal `rowSums()` is a complete different function than the "normal"
    `rowSums()` function.
    
    Of course (since they have the same name) in this case these functions are heavily related with each other: If we look into the source code of `rowSums()`, we see that it is a wrapper around the internal `rowSums()`. Just some input checkings, conversions and the special cases (complex numbers) are added:
    
    ```{r}
    rowSums
    ```

1.  __<span style="color:red">Q</span>__: Make a faster version of `chisq.test()` that only computes the chi-square
    test statistic when the input is two numeric vectors with no missing
    values. You can try simplifying `chisq.test()` or by coding from the 
    [mathematical definition](http://en.wikipedia.org/wiki/Pearson%27s_chi-squared_test).
    
    __<span style="color:green">A</span>__: Since `chisq.test()` has a relatively long source code, we try a new implementation from scratch:
    
    ```{r}
    chisq.test2 <- function(x, y){
      
      # Input
      if(!is.numeric(x)){
        stop("x must be numeric")}
      if(!is.numeric(y)){
        stop("y must be numeric")}
      if(length(x) != length(y)){
        stop("x and y must have the same length")}
      if(length(x) <= 1){
        stop("length of x must be greater one")}
      if(any(c(x,y) < 0)){
        stop("all entries of x and y must be greater or equal zero")}
      if(sum(complete.cases(x, y)) != length(x)){
        stop("there must be no missing values in x and y")}
      if(any(is.null(c(x,y)))){
        stop("entries of x and y must not be NULL")}
      
      # Help variables
      m <- rbind(x, y)
      margin1 <- rowSums(m)
      margin2 <- colSums(m)
      n <- sum(m)
      me <- tcrossprod(margin1, margin2) / n
      
      # Output
      x_stat = sum((m - me)^2 / me)
      dof <- (length(margin1) - 1) * (length(margin2) - 1)
      p <- pchisq(x_stat, df = dof, lower.tail = FALSE)
      
      return(list(x_stat = x_stat, df = dof, `p-value` = p))
    }
    ```
    
    We check if our new implementation returns the same results
    
    ```{r}
    a <- 21:25
    b <- c(21,23,25,27,29)
    m_test <- cbind(a, b)
    
    chisq.test(m_test)
    chisq.test2(a, b)
    ```
    
    Finally we benchmark this implementation against a compiled version of itself and the original `stats::chisq.test()`:
    
    ```{r}
    chisq.test2c <- compiler::cmpfun(chisq.test2)
    
    microbenchmark::microbenchmark(
      chisq.test(m_test),
      chisq.test2(a, b),
      chisq.test2c(a, b)
    )
    ```

1.  __<span style="color:red">Q</span>__: Can you make a faster version of `table()` for the case of an input of
    two integer vectors with no missing values? Can you use it to
    speed up your chi-square test?
    
    __<span style="color:green">A</span>__: 

1.  __<span style="color:red">Q</span>__: Imagine you want to compute the bootstrap distribution of a sample
    correlation using `cor_df()` and the data in the example below. Given that you 
    want to run this many times, how can you make this code faster? (Hint: the 
    function has three components that you can speed up.)

    ```{r, eval = FALSE}
    n <- 1e6
    df <- data.frame(a = rnorm(n), b = rnorm(n))

    cor_df <- function() {
      # note that we omit the inputparameter i, which appears in the textbook, since 
      # this seems to be unnecessary.
      i <- sample(seq(n), n * 0.01)
      cor(df[i, , drop = FALSE])[2,1]
      # note also that in the last line the textbook says q[] instead of df[]. Since
      # this is probably just a typo, we changed this to df[].
    }
    ```

    Is there a way to vectorise this procedure?
    
    __<span style="color:green">A</span>__: The three components (mentioned in the questions hint) are:

    1. sampling of indices
    2. subsetting the data frame/conversion to matrix (or vector input)
    3. the `cor()` function itself.
    
    Since a run of lineprof like shown in the textbook suggests that `as.matrix()` within the `cor()` function is the biggest bottleneck, we start with that:
    
    ```{r}
    n <- 1e6
    df <- data.frame(a = rnorm(n), b = rnorm(n))
    ```

    Remember the outgoing function:
    
    ```{r}
    cor_df <- function() {
      i <- sample(seq(n), n * 0.01)
      cor(df[i, , drop = FALSE])[2,1]
    }
    ```
    
    First we want to optimise the second line (without attention to the `cor()` function itself). Therefore we exclude the first line from our optimisation approaches and define `i` within the global environment:
    
    ```{r}
    i <- sample(seq(n), n * 0.01)
    ```
    
    Then we define our approaches, check that their behaviour is correct and do the first benchmark:
    
    ```{r}
    # old version
    cor_v1 <- function() {
      cor(df[i, , drop = FALSE])[2,1]
    }
    
    # cbind instead of internal as.matrix
    cor_v2 <- function() {
      m <- cbind(df$a[i], df$b[i])
      cor(m)[2, 1]
    }

    # cbind + vector subsetting of the output matrix
    cor_v3 <- function() {
      m <- cbind(df$a[i], df$b[i])
      cor(m)[2]
    }

    # Use vector input within the cor function, so that no conversion is needed
    cor_v4 <- function() {
      cor(df$a[i], df$b[i])
    }
    
    # check if all return the same result
    cor_list <- list(cor_v1, cor_v2, cor_v3, cor_v4)
    ulapply <- function(X, FUN, ...) unlist(lapply(X, FUN, ...))
    ulapply(cor_list, function(x) identical(x(), cor_v1()))
    
    # benchmark
    set.seed(1)
    microbenchmark::microbenchmark(
      cor_v1(),
      cor_v2(),
      cor_v3(),
      cor_v4()
    )
    ```
    
    According to the resulting medians, lower and upper quartiles of our benchmark all three new versions seem to provide more or less the same speed benefit (note that the maximum and mean can vary a lot for these approaches). Since the second version is most similar to the code we started, we implement this line into a second version of `cor_df()` (if this sounds too arbitrary, note that in the final solution we will come back to the vector input version anyway) and do a benchmark to get the overall speedup:
    
    ```{r}
    cor_df2 <- function() {
      i <- sample(seq(n), n * 0.01)
      m <- cbind(df$a[i], df$b[i])
      cor(m)[2, 1]
    }
    
    microbenchmark::microbenchmark(
      cor_df(),
      cor_df2()
    )
    ```
    
    Now we can focus on a speedup for the random generation of indices. (Note that a run of linepfrof suggests to optimize `cbind()`. However, after rewriting `cor()` to a version that only works with vector input, this step will be unnecessary anyway). We could try differnt approaches for the sequence generation within `sample()` (like `seq(n)`, `seq.int(n)`, `seq_len(n)`, `a:n`) and a direct call of `sample.int()`. In the following, we will see, that `sample.int()` is always faster (since we don't include the generation of the sequence into our benchmark). When we look into `sample.int()` we see that it calls two different internal sample versions depending on the input. Since in our usecase always the second version will be called, we also provide this version in our benchmark: 
    
    ```{r}
    seq_n <- seq(n)
    
    microbenchmark::microbenchmark(
      sample(seq_n, n * 0.01),
      sample.int(n, n * 0.01),
      .Internal(sample(n, n * 0.01, replace = FALSE, prob = NULL))
    )
    ```
    
    The `sample.int()` versions give clearly the biggest improvement. Since the internal version doesn't provide any clear improvement, but restricts the general scope of our function, we choose to implement `sample.int()` in a third version of `cor_df()` and benchmark our actual achievements:
    
    ```{r}
    cor_df3 <- function() {
      i <- sample.int(n, n * 0.01)
      m <- cbind(df$a[i], df$b[i])
      cor(m)[2, 1]
    }

    microbenchmark::microbenchmark(
      cor_df(),
      cor_df2(),
      cor_df3()
    )
    ```
    
    As a last step, we try to speedup the calculation of the pearson correlation coefficient. Since quite a lot of functionality is build into the `stats::cor()` function this seems like a reasonable approach. We try this by working with another `cor()` function from the `WGNA` package and an own implementation which should give a small improvement, because we use `sum(x) / length(x)` instead of `mean(x)` for internal calculations:
    
    ```{r}
    #WGCNA version (matrix and vector). Note that I don't use a local setup which uses
    #the full potential of this function. For furter information see ?WGCNA::cor
    cor_df4m <- function() {
      i <- sample.int(n, n * 0.01)
      m <- cbind(df$a[i], df$b[i])
      WGCNA::cor(m)[2]
    }
    
    cor_df4v <- function() {
      i <- sample.int(n, n * 0.01)
      WGCNA::cor(df$a[i], df$b[i], quick = 1)[1]
    }
    
    #New implementation of underlying cor function
    #A definition can be found for example here
    #http://www.socscistatistics.com/tests/pearson/
    cor2 <- function(x, y){
      xm <- sum(x) / length(x)
      ym <- sum(y) / length(y)
      x_xm <- x - xm
      y_ym <- y - ym
      numerator <- sum((x_xm) * (y_ym))
      denominator <- sqrt(sum(x_xm^2)) * sqrt(sum(y_ym^2))
      return(numerator / denominator)
    }
    
    cor2 <- compiler::cmpfun(cor2)
    
    cor_df5 <- function() {
      i <- sample.int(n, n * 0.01)
      cor2(df$a[i], df$b[i])
    }
    ```
    
    In our final benchmark, we also include compiled verions of all our attempts:
    ```{r}
    cor_df_c <- compiler::cmpfun(cor_df)
    cor_df2_c <- compiler::cmpfun(cor_df2)
    cor_df3_c <- compiler::cmpfun(cor_df3)
    cor_df4m_c <- compiler::cmpfun(cor_df4m)
    cor_df4v_c <- compiler::cmpfun(cor_df4v)
    cor_df5_c <- compiler::cmpfun(cor_df5)

    microbenchmark::microbenchmark(
      cor_df(),
      cor_df2(),
      cor_df3(),
      cor_df4m(),
      cor_df4v(),
      cor_df5(),
      cor_df_c(),
      cor_df2_c(),
      cor_df3_c(),
      cor_df4m_c(),
      cor_df4v_c(),
      cor_df5_c()
    )
    ```
    
    Our final solution benefits most from the switch from data frames to vectors. Working with `sample.int` gives only little improvement. Reimplementing and compiling a new correlation function adds only minimal speedup.
    
    To trust our final result we include a last check for similar return values:
    
    ```{r}
    set.seed(1)
    cor_df()
    set.seed(1)
    cor_df5_c()
    ```
    
    Vectorization of this problem seems rather difficult, since attempts of using matrix calculus, always depend on building and handling big matrices in the first place.
    
    We can for example rewrite our correlation function to work with matrices and build a new (vectorized) version of `cor_df()` on top of that
    
    ```{r}
    cor2m <- function(x, y){
      n_row <- nrow(x)
      xm <- colSums(x) / n_row
      ym <- colSums(y) / n_row
      x_xm <- t(t(x) - xm)
      y_ym <- t(t(y) - ym)
      numerator <- colSums((x_xm) * (y_ym))
      denominator <- sqrt(colSums(x_xm^2)) * sqrt(colSums(y_ym^2))
      return(numerator / denominator)
    }
    
    cor_df_v <- function(i){
      indices <- replicate(i, sample.int(n, n * 0.01), simplify = "array")
      x <- matrix(df$a[indices], ncol = i)
      y <- matrix(df$b[indices], ncol = i)
      cor2m(x, y)
    }
    cor_df_v <- compiler::cmpfun(cor_df_v)
    ```
    
    However this still doesn't provide any improvement over the use of `lapply()`:
    
    ```{r}
    ulapply2 <- function(X, FUN, ...) unlist(lapply(X, FUN, ...), use.names = FALSE)
    
    microbenchmark::microbenchmark(
      cor_df5_c(),
      ulapply2(1:100, function(x) cor_df5_c()),
      cor_df_v(100)
    )
    
    ```
    
    Further improvements can be achieved using parallelisation (for example via `parallel::parLapply()`)
    
## Vectorise

1.  __<span style="color:red">Q</span>__: The density functions, e.g., `dnorm()`, have a common interface. Which 
    arguments are vectorised over? What does `rnorm(10, mean = 10:1)` do?
    
    __<span style="color:green">A</span>__: 

1.  __<span style="color:red">Q</span>__: Compare the speed of `apply(x, 1, sum)` with `rowSums(x)` for varying sizes
    of `x`.
    
    __<span style="color:green">A</span>__: 
  
1.  __<span style="color:red">Q</span>__: How can you use `crossprod()` to compute a weighted sum? How much faster is
    it than the naive `sum(x * w)`?
    
    __<span style="color:green">A</span>__: 