# Profiling

## Has somebody already solved the problem?

1.  What are faster alternatives to `lm`? Which are specifically designed 
    to work with larger datasets?

1.  What package implements a version of `match()` that's faster for
    repeated lookups? How much faster is it?

1.  List four functions (not just those in base R) that convert a string into a
    date time object. What are their strengths and weaknesses?

1.  How many different ways can you compute a 1d density estimate in R?

1.  Which packages provide the ability to compute a rolling mean?

1.  What are the alternatives to `optim()`?

## Do as little as possible

1.  How do the results change if you compare `mean()` and `mean.default()`
    on 10,000 observations, rather than on 100?

1.  The following code provides an alternative implementation of `rowSums()`.
    Why is it faster for this input?

    ```{r}
    rowSums2 <- function(df) {
      out <- df[[1L]]
      if (ncol(df) == 1) return(out)
      
      for (i in 2:ncol(df)) {
        out <- out + df[[i]]
      }
      out
    }
    
    df <- as.data.frame(
      replicate(1e3, sample(100, 1e4, replace = TRUE))
    )
    system.time(rowSums(df))
    system.time(rowSums2(df))
    ```

1.  What's the difference between `rowSums()` and `.rowSums()`?

1.  Make a faster version of `chisq.test()` that only computes the chi-square
    test statistic when the input is two numeric vectors with no missing
    values. You can try simplifying `chisq.test()` or by coding from the 
    [mathematical definition](http://en.wikipedia.org/wiki/Pearson%27s_chi-squared_test).

1.  Can you make a faster version of `table()` for the case of an input of
    two integer vectors with no missing values? Can you use it to
    speed up your chi-square test?

1.  Imagine you want to compute the bootstrap distribution of a sample
    correlation using `cor_df()` and the data in the example below. Given that you 
    want to run this many times, how can you make this code faster? (Hint: the 
    function has three components that you can speed up.)

    ```{r, eval = FALSE}
    n <- 1e6
    df <- data.frame(a = rnorm(n), b = rnorm(n))

    cor_df <- function(i) {
      i <- sample(seq(n), n * 0.01)
      cor(q[i, , drop = FALSE])[2,1]
    }
    ```

    Is there a way to vectorise this procedure?

## Vectorise

1.  The density functions, e.g., `dnorm()`, have a common interface. Which 
    arguments are vectorised over? What does `rnorm(10, mean = 10:1)` do?

1.  Compare the speed of `apply(x, 1, sum)` with `rowSums(x)` for varying sizes
    of `x`.
  
1.  How can you use `crossprod()` to compute a weighted sum? How much faster is
    it than the naive `sum(x * w)`?