# Memory

## Object size

1.  Repeat the analysis above for numeric, logical, and complex vectors.

1.  If a data frame has one million rows, and three variables (two numeric, and 
    one integer), how much space will it take up? Work it out from theory, 
    then verify your work by creating a data frame and measuring its size.

1.  Compare the sizes of the elements in the following two lists. Each 
    contains basically the same data, but one contains vectors of small 
    strings while the other contains a single long string.

    ```{r}
    vec <- lapply(0:50, function(i) c("ba", rep("na", i)))
    str <- lapply(vec, paste0, collapse = "")
    ```

1.  Which takes up more memory: a factor (`x`) or the equivalent character 
    vector (`as.character(x)`)? Why?

1.  Explain the difference in size between `1:5` and `list(1:5)`.

## Memory profiling with lineprof

1. When the input is a list, we can make a more efficient `as.data.frame()` 
   by using special knowledge. A data frame is a list with class `data.frame` 
   and `row.names` attribute. `row.names` is either a character vector or 
   vector of sequential integers, stored in a special format created by 
   `.set_row_names()`. This leads to an alternative `as.data.frame()`:

    ```{r}
    to_df <- function(x) {
      class(x) <- "data.frame"
      attr(x, "row.names") <- .set_row_names(length(x[[1]]))
      x
    }
    ```

    What impact does this function have on `read_delim()`?  What are the 
    downsides of this function?

1.  Line profile the following function with `torture = TRUE`. What is 
    surprising? Read the source code of `rm()` to figure out what's going on.

    ```{r}
    f <- function(n = 1e5) {
      x <- rep(1, n)
      rm(x)
    }
    ```

## Modification in place

1.  The code below makes one duplication. Where does it occur and why? 
    (Hint: look at `refs(y)`.)

    ```{r}
    x <- data.frame(matrix(runif(100 * 1e4), ncol = 100))
    medians <- vapply(x, median, numeric(1))
    y <- as.list(x)
    for(i in seq_along(medians)) {
      y[[i]] <- y[[i]] - medians[i]
    }
    ```

1.  The implementation of `as.data.frame()` in the previous section has one 
    big downside. What is it and how could you avoid it?


[long-vectors]: http://cran.r-project.org/doc/manuals/R-ints.html#Long-vectors