# Memory

## Object size

1.  __<span style="color:red">Q</span>__: Repeat the analysis above for numeric, logical, and complex vectors.
    
    __<span style="color:green">A</span>__: 
    
    **numeric:**
    
    ```{r}
    library(pryr)
    sizes <- sapply(0:50, function(n) object_size(vector("numeric", n)))
    plot(0:50, sizes, xlab = "Length", ylab = "Size (bytes)",
         type = "s")
    
    plot(0:50, sizes - 40, xlab = "Length", 
         ylab = "Bytes excluding overhead", type = "n")
    abline(h = 0, col = "grey80")
    abline(h = c(8, 16, 32, 48, 64, 128), col = "grey80")
    abline(a = 0, b = 8, col = "grey90", lwd = 4)
    lines(sizes - 40, type = "s")
    
    x <- numeric(1e6)
    object_size(x)
    y <- list(x, x, x)
    object_size(y)
    ```
    
    **logical:**
    
    ```{r}
    sizes <- sapply(0:50, function(n) object_size(vector("logical", n)))
    plot(0:50, sizes, xlab = "Length", ylab = "Size (bytes)", 
         type = "s")
    
    plot(0:50, sizes - 40, xlab = "Length", 
         ylab = "Bytes excluding overhead", type = "n")
    abline(h = 0, col = "grey80")
    abline(h = c(8, 16, 32, 48, 64, 128), col = "grey80")
    abline(a = 0, b = 4, col = "grey90", lwd = 4)
    lines(sizes - 40, type = "s")
    
    x <- logical(1e6)
    object_size(x)
    y <- list(x, x, x)
    object_size(y)
    ```
    
    **complex:**
    
    ```{r}
    sizes <- sapply(0:50, function(n) object_size(vector("complex", n)))
    plot(0:50, sizes, xlab = "Length", ylab = "Size (bytes)", 
         type = "s")
    
    plot(0:50, sizes - 40, xlab = "Length", 
         ylab = "Bytes excluding overhead", type = "n")
    abline(h = 0, col = "grey80")
    abline(h = c(8, 16, 32, 48, 64, 128), col = "grey80")
    abline(a = 0, b = 16, col = "grey90", lwd = 4)
    lines(sizes - 40, type = "s")
    
    x <- complex(1e6)
    object_size(x)
    y <- list(x, x, x)
    object_size(y)
    ```

1.  __<span style="color:red">Q</span>__: If a data frame has one million rows, and three variables (two numeric, and 
    one integer), how much space will it take up? Work it out from theory, 
    then verify your work by creating a data frame and measuring its size.
    
    __<span style="color:green">A</span>__: From the textbook we know that 
    
    * an integer's size is 40 bytes plus 4 bytes per allocated entry,
    * a numerics's size is 40 bytes plus 8 bytes per allocated entry.
    
    So we can calculate the size via:
    
    `object_size(df) = 1 * (40 + 4 * 1,000,000) + 2 * (40 + 8 * 1,000,000)
                           = 20,000,120` bytes.
    
    And test this via:
    
    ```{r}
    df <- data.frame(int1 = integer(1000000),
                     num1 = numeric(1000000),
                     num2 = numeric(1000000))
    as.integer(object_size(df))
    ```
    
    Note that we observe a small difference, because we didn't include the costs for creating the `data.frame()` (560 bytes) in our previous calculations.
    
1.  __<span style="color:red">Q</span>__: Compare the sizes of the elements in the following two lists. Each 
    contains basically the same data, but one contains vectors of small 
    strings while the other contains a single long string.

    ```{r}
    vec <- lapply(0:50, function(i) c("ba", rep("na", i)))
    str <- lapply(vec, paste0, collapse = "")
    ```
    
    __<span style="color:green">A</span>__: 
    
    ```{r}
    vec <- lapply(0:50, function(i) c("ba", rep("na", i)))
    str <- lapply(vec, paste0, collapse = "")
    object_size(vec)
    object_size(str)
    object_size(vec, str)
    ```

1.  __<span style="color:red">Q</span>__: Which takes up more memory: a factor (`x`) or the equivalent character 
    vector (`as.character(x)`)? Why?
    
    __<span style="color:green">A</span>__: To be exact: it depends on the length of unique elements in relation to the overall length of the vector.
    
    * In case of a long vector with only a few levels, the character takes approximately twice the memory of a factor:
    
    ```{r}
    object_size(rep(letters[1:20], 1000))
    object_size(factor(rep(letters[1:20], 1000)))
    ```
    
    That is, because a character allocates 8 bytes per entry (if the entry has less than 8 signs, otherwise roughly one byte per sign) and a factor equals an integer (allocates only 4 bytes per entry) with a character vector attribute that contains the levels (unique elements) of the vector:
    
    ```{r}
    a <- rep(1:20, 1000)
    object_size(a)
    
    attr(a, "levels") <- letters[1:20]
    object_size(a)
    
    class(a) <- "factor"
    object_size(a)
    ```
    
    * Of course the factor will allocate more memory, if all entries are unique:
    
    ```{r}
    object_size(letters[1:20])
    object_size(factor(letters[1:20]))
    ```

1.  __<span style="color:red">Q</span>__: Explain the difference in size between `1:5` and `list(1:5)`.

    __<span style="color:green">A</span>__: An empty list needs 40 bytes. For each entry 8 bytes are added. We can see this via:
    
    ```{r}
    object_size(vector("list",0))
    object_size(vector("list",1))
    ```
    
    Since `1:5` needs 72 bytes (note that for memory for short integers is allocated in chunks, as explained in the textbook), `list(1:5)` takes 120 bytes (48 + 72). So in general the cost for saving atomics within a list is 40 bytes for the list plus 8 bytes per atomic/list entry.

## Memory profiling with lineprof

1. __<span style="color:red">Q</span>__: When the input is a list, we can make a more efficient `as.data.frame()` 
   by using special knowledge. A data frame is a list with class `data.frame` 
   and `row.names` attribute. `row.names` is either a character vector or 
   vector of sequential integers, stored in a special format created by 
   `.set_row_names()`. This leads to an alternative `as.data.frame()`:

    ```{r}
    to_df <- function(x) {
      class(x) <- "data.frame"
      attr(x, "row.names") <- .set_row_names(length(x[[1]]))
      x
    }
    ```

    What impact does this function have on `read_delim()`?  What are the 
    downsides of this function?

1.  __<span style="color:red">Q</span>__: Line profile the following function with `torture = TRUE`. What is 
    surprising? Read the source code of `rm()` to figure out what's going on.

    ```{r}
    f <- function(n = 1e5) {
      x <- rep(1, n)
      rm(x)
    }
    ```

## Modification in place

1.  __<span style="color:red">Q</span>__: The code below makes one duplication. Where does it occur and why? 
    (Hint: look at `refs(y)`.)

    ```{r}
    x <- data.frame(matrix(runif(100 * 1e4), ncol = 100))
    medians <- vapply(x, median, numeric(1))
    y <- as.list(x)
    for(i in seq_along(medians)) {
      y[[i]] <- y[[i]] - medians[i]
    }
    ```
    
    __<span style="color:green">A</span>__: It occurs in the first iteration of the for loop. `refs(y)` is 2 before the for loop, because `y` is created via `as.list()`, which is not a primitive and so sets the refs counter up to two. Therefore **R** makes a copy, `refs(y)` becomes 1 and the following modifications will occur in place. The following code illustrates this behaviour, when run in the **R**Gui. (Note that `refs()` will always return 2, when run in RStudio, as stated in the textbook. Note also, that you could detect this behaviour with the `tracemem()` function).
    
    ```{r, eval = FALSE}
    library(pryr)
    rm(list = ls(all = TRUE))

    x <- data.frame(matrix(runif(100 * 1e4), ncol = 4))
    medians <- vapply(x, median, numeric(1))

    y <- as.list(x)
    is.primitive(as.list)
    # [1] FALSE

    for(i in seq_along(medians)) {
      print(c(address(y), refs(y)))
      y[[i]] <- y[[i]] - medians[i]
    }
    # [1] "0x46c4a98" "2"
    # [1] "0x11de30c8" "1" 
    # [1] "0x11de30c8" "1" 
    # [1] "0x11de30c8" "1" 
    ```

1.  __<span style="color:red">Q</span>__: The implementation of `as.data.frame()` in the previous section has one 
    big downside. What is it and how could you avoid it?


[long-vectors]: http://cran.r-project.org/doc/manuals/R-ints.html#Long-vectors
