# Rcpp

## Getting started with C++

1.  __<span style="color:red">Q</span>__: With the basics of C++ in hand, it's now a great time to practice by reading and writing some simple C++ functions. For each of the following functions, read the code and figure out what the corresponding base R function is. You might not understand every part of the code yet, but you should be able to figure out the basics of what the function does.

    ```cpp
    double f1(NumericVector x) {
      int n = x.size();
      double y = 0;
    
      for(int i = 0; i < n; ++i) {
        y += x[i] / n;
      }
      return y;
    }
    
    NumericVector f2(NumericVector x) {
      int n = x.size();
      NumericVector out(n);
    
      out[0] = x[0];
      for(int i = 1; i < n; ++i) {
        out[i] = out[i - 1] + x[i];
      }
      return out;
    }
    
    bool f3(LogicalVector x) {
      int n = x.size();
    
      for(int i = 0; i < n; ++i) {
        if (x[i]) return true;
      }
      return false;
    }
    
    int f4(Function pred, List x) {
      int n = x.size();
    
      for(int i = 0; i < n; ++i) {
        LogicalVector res = pred(x[i]);
        if (res[0]) return i + 1;
      }
      return 0;
    }
    
    NumericVector f5(NumericVector x, NumericVector y) {
      int n = std::max(x.size(), y.size());
      NumericVector x1 = rep_len(x, n);
      NumericVector y1 = rep_len(y, n);
    
      NumericVector out(n);
    
      for (int i = 0; i < n; ++i) {
        out[i] = std::min(x1[i], y1[i]);
      }
    
      return out;
    }
    ```

__<span style="color:green">A</span>__:  The R equivalents are:  

*   f1: `mean()`  
*   f2: `cumsum()`  
*   f3: `any()`
*   f4:  
*   f5: `pmin()`

1.  __<span style="color:red">Q</span>__:  To practice your function writing skills, convert the following functions into C++. For now, assume the inputs have no missing values.  
  
    1. `all()`
  
    ```{r, eval = FALSE}
    cppFunction('
    bool allC(LogicalVector x) {
    int n = x.size();
    
    for(int i = 0; i < n; ++i) {
      if (!x[i]) return false;
    }
    return true;
    }')
    ```
  
    2. `cumprod()`, `cummin()`, `cummax()`.
  
    ```{r, eval = FALSE}
    NumericVector cumprodC(NumericVector x) {
    int n = x.size();
    NumericVector out(n);
    
    out[0] = x[0];
    for(int i = 1; i < n; ++i) {
      out[i]  = out[i - 1] * x[i];
    }
    return out;
    }')
    
    cppFunction('
    NumericVector cumminC(NumericVector x) {
    
    int n = x.size();
    NumericVector out(n);
    
    out[0] = x[0];
    for(int i = 1; i < n; ++i) {
      out[i]  = std::min(out[i - 1], x[i]);
    }
    return out;
    }')
    
    cppFunction('
    NumericVector cummaxC(NumericVector x) {
    int n = x.size();
    NumericVector out(n);
    
    out[0] = x[0];
    for(int i = 1; i < n; ++i) {
      out[i]  = std::max(out[i - 1], x[i]);
    }
    return out;
    }')
    ```

    3. `diff()`. Start by assuming lag 1, and then generalise for lag `n`.  
    
    ```{r, eval = FALSE}
    cppFunction('
    NumericVector diffC(NumericVector x){
    int n = x.size();
    NumericVector out(n - 1);
    
    for(int i = 1; i < n; i++){
        out[i - 1] = x[i] - x[i - 1];
    }
    return out ;
    }')

    cppFunction('
    NumericVector difflagC(NumericVector x, int lag){
    int n = x.size();
    NumericVector out(n - lag);
            
    for(int i = lag; i < n; i++){
        out[i - lag] = x[i] - x[i - lag];
    }
    return out;
    }')
    ```
  
    4. `range`.  
    
    ```{r, eval = FALSE}
    cppFunction('NumericVector rangeC(NumericVector x){
    double omin, omax;  
    int n = x.size();
    NumericVector out(2);

    omin = x[0];
    omax = x[0];

    for(int i = 1; i < n; i++){
        omin = std::min(x[i], omin);
        omax = std::max(x[i], omax);
    }

    out[0] = omin;
    out[1] = omax;
    return out;
    }')
    ```
  
    5. `var`. Read about the approaches you can take on 
     [wikipedia](http://en.wikipedia.org/wiki/Algorithms_for_calculating_variance).
    Whenever implementing a numerical algorithm, it's always good to check what 
    is already known about the problem.

## Missing values

1. Rewrite any of the functions from the first exercise to deal with missing 
   values. If `na.rm` is true, ignore the missing values. If `na.rm` is false, 
   return a missing value if the input contains any missing values. Some 
   good functions to practice with are `min()`, `max()`, `range()`, `mean()`, 
   and `var()`.

1. Rewrite `cumsum()` and `diff()` so they can handle missing values. Note that 
   these functions have slightly more complicated behaviour.

## The STL

To practice using the STL algorithms and data structures, implement the following using R functions in C++, using the hints provided:

1. `median.default()` using `partial_sort`.

1. `%in%` using `unordered_set` and the `find()` or `count()` methods.

1. `unique()` using an `unordered_set` (challenge: do it in one line!).

1. `min()` using `std::min()`, or `max()` using `std::max()`.

1. `which.min()` using `min_element`, or `which.max()` using `max_element`.

1. `setdiff()`, `union()`, and `intersect()` for integers using sorted ranges 
   and `set_union`, `set_intersection` and `set_difference`.