```{r, include=FALSE}
source("common.R")
```

# Quasiquotation
## Prerequisites {-}

To continue computing on the language, we keep using the rlang package in this chapter.

```{r setup}
library(rlang)
```

## Motivation

1. __[Q]{.Q}__: For each function in the following base R code, identify which arguments are quoted and which are evaluated.

    ```{r, eval = FALSE}
    library(MASS)

    mtcars2 <- subset(mtcars, cyl == 4)

    with(mtcars2, sum(vs))
    sum(mtcars2$am)

    rm(mtcars2)
    ```
    
   __[A]{.solved}__: For each argument we first follow the advice from the textbook and execute the argument outside of the respective function. Since `MASS`, `cyl`, `vs` and `am` are not objects contained in the global environment, their execution raises an "Object not found" error. This way we confirm that the respective function arguments are quoted. For the other arguments, we may inspect the source code (and the documentation) to check if any quoting mechanisms are applied or the arguments are evaluated.

    ```{r, eval = FALSE}
    library(MASS)  # MASS -> quoted
    ```

   `library()` also accepts character vectors and doesn't quote when `character.only` is set to `TRUE`, so `library(MASS, character.only = TRUE)` will raise an error.

    ```{r, eval = FALSE}
    mtcars2 <- subset(mtcars, cyl == 4)  # mtcars -> evaluated
                                         # cyl -> quoted
                                         
    with(mtcars2, sum(vs))  # mtcars2 -> evaluated
                            # sum(vs) -> quoted
                            
    sum(mtcars2$am)  # matcars$am -> evaluated
                     # am -> quoted by $()    
    ```
    
   When we inspect the source code of `rm()`, we notice that `rm()` catches its `...` argument as an unevaluated call (in this case a pairlist) via `match.call()`. This call is then converted into a string for further evaluation.

    ```{r, eval = FALSE}
    rm(mtcars2)  # mtcars2 -> quoted
    ```
   
2. __[Q]{.Q}__: For each function in the following tidyverse code, identify which arguments are quoted and which are evaluated.

    ```{r, eval = FALSE}
    library(dplyr)
    library(ggplot2)

    by_cyl <- mtcars %>%
      group_by(cyl) %>%
      summarise(mean = mean(mpg))

    ggplot(by_cyl, aes(cyl, mean)) + geom_point()
    ```

   __[A]{.solved}__: From the previous exercise we've already learned that `library()` quotes its first argument.
    
    ```{r, eval = FALSE}
    library(dplyr)    # dplyr   -> quoted
    library(ggplot2)  # ggplot2 -> quoted
    ```
    
   In similar fashion, it becomes clear that `cyl` is quoted by `group_by()`.
    
    ```{r, eval = FALSE}
    by_cyl <- mtcars %>%           # mtcars -> evaluated
      group_by(cyl) %>%            # cyl -> quoted
      summarise(mean = mean(mpg))  # mean = mean(mpg) -> quoted
    ```
   
   To find out what happens in `summarise()`, we inspect the source code. Tracing down the S3-dispatch of `summarise()`, we see that the `...` argument is quoted in the underlying `summarise.tbl_df()` method.
    
    ```{r}
    dplyr::summarise
    
    dplyr:::summarise.tbl_df
    ```
    
   In the following ggplot2 expression the `cyl`- and `mean`-objects are quoted.
    
    ```{r, eval = FALSE}
    ggplot(by_cyl,            # by_cyl -> evaluated
           aes(cyl, mean)) +  # aes() -> evaluated
                              # cyl, mean -> quoted (via aes)
      geom_point() 
    ```
    
   We can confirm this also by inspecting `aes()`'s source code.
   
    ```{r}
    ggplot2::aes
    ```
    
## Quoting

1. __[Q]{.Q}__: How is `expr()` implemented? Look at its source code.

   __[A]{.solved}__: `expr()` acts as a simple wrapper, which passes its argument to `enexpr()`.
   
    ```{r}
    expr
    ```

2. __[Q]{.Q}__: Compare and contrast the following two functions. Can you predict the output before running them?
   
    ```{r, results = FALSE}
    f1 <- function(x, y) {
      exprs(x = x, y = y)
    }
    f2 <- function(x, y) {
      enexprs(x = x, y = y)
    }
    f1(a + b, c + d)
    f2(a + b, c + d)
    ```
    
   __[A]{.solved}__: Both functions are able to capture multiple arguments and will return a named list of expressions. `f1()` will return the arguments defined within the body of `f1()`. This happens because `exprs()` captures the expressions as specified by the developer during the definition of `f1`.
   
    ```{r}
    f1(a + b, c + d)
    ```
    
   `f2()` will return the arguments supplied to `f2()` as specified by the user when the function is called. 
    
    ```{r}
    f2(a + b, c + d)
    ```
    
3. __[Q]{.Q}__: What happens if you try to use `enexpr()` with an expression (i.e. `enexpr(x + y)`)? What happens if `enexpr()` is passed a missing argument?
    
   __[A]{.solved}__: In the first case an error is thrown:
    
    ```{r, error = TRUE}
    on_expr <- function(x) {enexpr(expr(x))}
    on_expr(x + y)
    ```
    
   In the second case a missing argument is returned:
    
    ```{r}
    on_missing <- function(x) {enexpr(x)}
    on_missing()
    is_missing(on_missing())
    ```

4. __[Q]{.Q}__: How are `exprs(a)` and `exprs(a = )` different? Think about both the input and the output.
    
   __[A]{.solved}__: In `exprs(a)` the input `a` is interpreted as a symbol for an unnamed argument. Consequently the output shows an unnamed list with the first element containing the symbol `a`.
   
    ```{r}
    out1 <- exprs(a)
    str(out1)
    ```   
   
   In `exprs(a = )` the first argument is named `a`, but then no value is provided. This leads to the output of a named list with the first element named `a`, which contains the missing argument.
    
    ```{r}
    out2 <- exprs(a = )
    str(out2)
    is_missing(out2$a)
    ```

5. __[Q]{.Q}__: What are other differences between `exprs()` and `alist()`? Read the documentation for the named arguments of `exprs()` to find out.

   __[A]{.solved}__: `exprs()` provides the additional arguments `.named` (`= FALSE`), `.ignore_empty` (`c("trailing", "none", "all")`) and `.unquote_names` (`TRUE`). `.named` allows to ensure that all dots are named. `ignore_empty` allows to specify how empty arguments should be handled for dots (`"trailing"`) or all arguments (`"none"` and `"all"`). Further via `.unquote_names` one can specify if `:=` should be treated like `=`. `:=` can be useful as it supports unquoting (`!!`) on the left-hand-side.

6. __[Q]{.Q}__: The documentation for `substitute()` says:

   > Substitution takes place by examining each component of the parse tree 
   > as follows: 
   > 
   > * If it is not a bound symbol in `env`, it is unchanged. 
   > * If it is a promise object (i.e., a formal argument to a function) the expression slot of the promise replaces the symbol. 
   > * If it is an ordinary variable, its value is substituted, unless `env` is .GlobalEnv in which case the symbol is left unchanged.

   Create examples that illustrate each of the above cases.
    
   __[A]{.solved}__: Let's create a new environment `a`, which contains no objects. In this case `substitute()` will just return its first argument (`expr`):
   
    ```{r}
    my_env <- env()
    substitute(x, my_env)
    ```
    
   When we create a function containing one argument, which is directly returned after substitution. This function just returns the provided expression:
   
    ```{r}
    foo <- function(x) substitute(x)

    foo(x + y * sin(0))
    ```
    
   In case `substitute()` can find (parts of) the expression in `env`, it will literally substitute. However, unless `env` is `.GlobalEnv`.
   
    ```{r}
    my_env$x <- 7
    substitute(x, my_env)
    
    x <- 7
    substitute(x, .GlobalEnv)
    ```
    
## Unquoting

1. __[Q]{.Q}__: Given the following components:

    ```{r}
    xy <- expr(x + y)
    xz <- expr(x + z)
    yz <- expr(y + z)
    abc <- exprs(a, b, c)
    ```
    
   Use quasiquotation to construct the following calls:
    
    ```{r, eval = FALSE}
    (x + y) / (y + z)               # (1)
    -(x + z) ^ (y + z)              # (2)
    (x + y) + (y + z) - (x + y)     # (3)
    atan2(x + y, y + z)             # (4)
    sum(x + y, x + y, y + z)        # (5)
    sum(a, b, c)                    # (6)
    mean(c(a, b, c), na.rm = TRUE)  # (7)
    foo(a = x + y, b = y + z)       # (8)
    ```
    
   __[A]{.solved}__: We combine and unquote the given quoted expressions to construct the desired calls like this:
    
    ```{r}
    expr(!!xy / !!yz)                    # (1)
    
    expr(-(!!xz)^(!!yz))                 # (2)
    
    expr(!!xy + !!yz - !!xz)             # (3)
    
    expr(atan2(!!xy, !!yz))              # (4)
    
    expr(sum(!!xy, !!xy, !!yz))          # (5)
    
    expr(sum(!!!abc))                    # (6)
    
    expr(mean(c(!!!abc), na.rm = TRUE))  # (7)
    
    expr(foo(a = !!xy, b = !!yz))            # (8)
    ```

2. __[Q]{.Q}__: The following two calls print the same, but are actually different:

    ```{r}
    (a <- expr(mean(1:10)))
    (b <- expr(mean(!!(1:10))))
    identical(a, b)
    ```

   What's the difference? Which one is more natural?
    
   __[A]{.solved}__: It's easiest to see the difference with `lobstr::ast()`:
   
    ```{r}
    lobstr::ast(mean(1:10))
    lobstr::ast(mean(!!(1:10)))
    ```

   In the expression assigned to `b` the call `1:10` is evaluated to an integer vector, while still being a call object in `a`.
   
   The first version (`a`) seems more natural. It captures lazy evaluation, with a promise that is evaluated when the function is called. The second version (`b`) inlines a vector directly in to a call.

## Dot-dot-dot (`...`)

1. __[Q]{.Q}__: One way to implement `exec()` is shown below. Describe how it works. What are the key ideas?
    
    ```{r, eval = FALSE}
    exec <- function(f, ..., .env = caller_env()) {
      args <- list2(...)
      do.call(f, args, envir = .env)
    }
    ```
    
   __[A]{.solved}__: `exec()` takes a function and its arguments as input, as well as an environment. This allows to construct a call from the function and the arguments and evaluate it in the supplied environment. As the `...` argument is handled via `list2()`, `exec()` supports tidy dots (quasiquotation), which means that arguments and names (on the left-hand-side of `:=`) can be unquoted via `!!` and `!!!`.

2. __[Q]{.Q}__: Carefully read the source code for `interaction()`, `expand.grid()`, and `par()`. Compare and contrast the techniques they use for switching between dots and list behaviour.

   __[A]{.solved}__:  All three functions capture the dots via `args <- list(...)`. `interaction()` and `expand.grid()` return early in case of `length(args) == 0`. 
   
   As `interaction()` computes factors regarding combinations of `args` elements, `interaction()` iterates over `args` and doesn't differentiate further between list and dots behaviour. Only the case `length(args) == 1 && is.list(args[[1]])` are treated via `args <- args[[1]]`. Consequently lists deeper than 1 level raise errors in other parts of the code.
   
    ```{r, error = TRUE}
    # These work and return the same
    identical(
      interaction(     a = c("a", "b", "c", "d"), b = c("e", "f")),
      interaction(list(a = c("a", "b", "c", "d"), b = c("e", "f")))
    )
    
    # This doesn't work
    interaction(list(list(a = c("a", "b", "c", "d"), b = c("e", "f"))))
    ```
    
   `expand.grid()` switches in exactly the same way as `interaction()`. I.e. it also assigns `args <- args[[1]]` in case of `length(args) == 1 && is.list(args[[1]])` is `TRUE`.
   
   `par()` preprocesses `args` the most in order to ensure that it becomes a list (or `NULL`). First, in case no dots were supplied (`length(list(...)) == 0`) `par()` creates a list from an internal character vector (partly depending on `par()`'s `no.readonly` argument). Further, in case all elements of `args` are character vectors (`all(unlist(lapply(args, is.character)))`) `args` is turned into a list via `as.list(unlist(args))`. When `args` is of length one with its first element being a list or `NULL` `args` becomes `args <- args[1]`.

3. __[Q]{.Q}__: Explain the problem with this definition of `set_attr()`
    
    ```{r, error = TRUE}
    set_attr <- function(x, ...) {
      attr <- rlang::list2(...)
      attributes(x) <- attr
      x
    }
    set_attr(1:10, x = 10)
    ```
    
   __[A]{.solved}__: `set_attr()` expects an object to be passed as the `x` argument and its new attributes via the dots. Unfortunately, this prohibits us to provide attributes named `x` as these would collide with the argument name of our object. Even omitting the object's argument name doesn't help in this case - as can be seen in the example where the object is consequently treated as an unnamed attribute. 
   
   However, we may name the first argument `.x`, which seems clearer and less likely to invoke errors. In this case `1:10` will get the (named) attribute `x = 10` assigned:

    ```{r}
    set_attr <- function(.x, ...) {
      attr <- rlang::list2(...)
      
      attributes(.x) <- attr
      .x
    }
    
    set_attr(1:10, x = 10)
    ```

## Case studies {#expr-case-studies}
    
1. __[Q]{.Q}__: In the linear-model example, we could replace the `expr()` in `reduce(summands, ~ expr(!!.x + !!.y))` with `call2()`: `reduce(summands, call2, "+")`. Compare and contrast the two approaches. Which do you think is easier to read?

   __[A]{.solved}__: We would consider the first version to be more readable. There seems to be a little more boilerplate code at first, but the unquoting syntax is very readable. Overall the whole expression seems more explicit and less complex.

2. __[Q]{.Q}__: Re-implement the Box-Cox transform defined below using unquoting and `new_function()`:
   
    ```{r}
    bc <- function(lambda) {
      if (lambda == 0) {
        function(x) log(x)
      } else {
        function(x) (x ^ lambda - 1) / lambda
      }
    }
    ```
    
   __[A]{.solved}__: Here `new_function()` allows us to create a function factory using tidy evaluation.
    
    ```{r}
    bc2 <- function(lambda){
      lambda <- enexpr(lambda)
      
      if (!!lambda == 0) {
        new_function(exprs(x = ), expr(log(x)))
      } else {
        new_function(exprs(x = ), expr((x ^ (!!lambda) - 1) / !!lambda))
      }
    }
    
    bc2(0)
    bc2(2)
    bc2(2)(2)
    ```
   
3. __[Q]{.Q}__:Re-implement the simple `compose()` defined below using quasiquotation and `new_function()`:
     
    ```{r}
    compose <- function(f, g) {
      function(...) f(g(...))
    }
    ```
    
   __[A]{.solved}__: The implementation is fairly straightforward, even though a lot of parentheses are required:
    
    ```{r}
    compose2 <- function(f, g){
      f <- enexpr(f)
      g <- enexpr(g)
      
      new_function(exprs(... = ), expr((!!f)((!!g)(...))))
    }
    
    compose(sin, cos)
    compose(sin, cos)(pi)
    compose2(sin, cos)
    compose2(sin, cos)(pi)
    ```
