# Vectors

## Atomic vectors

1. __<span style="color:red">Q</span>__: How do you create scalars of type raw and complex? (See `?raw` and `?complex`)

   __<span style="color:green">A</span>__: In R scalars are represented as vectors of length one. For raw and complex types these can be created via `raw()` and `complex()`, i.e.:
  
    ```{r}
    raw(1)
    complex(1)
    ```
  
   For raw vectors it's easy to coerce numeric or character scalars to raw.
  
    ```{r}
    as.raw(42)
    charToRaw("A")
    ```
  
   For complex numbers real and imaginary parts can be provided directly.
  
    ```{r}
    complex(length.out = 1, real = 1, imaginary = 1)
    ```

2. __<span style="color:red">Q</span>__: Test your knowledge of vector coercion rules by predicting the output of
   the following uses of `c()`: 

    ```{r, eval=FALSE}
    c(1, FALSE)      # will be coerced to numeric   -> 1 0
    c("a", 1)        # will be coerced to character -> "a" "1"
    c(TRUE, 1L)      # will be coerced to integer   -> 1 1
    ```

3. __<span style="color:red">Q</span>__: Why is `1 == "1"` true? Why is `-1 < FALSE` true? Why is `"one" < 2` false?  

   __<span style="color:green">A</span>__: These comparisons are carried out by operator-functions, which coerce their arguments to a common type. In the examples above these cases will be character, double and character: `1` will be coerced to `"1"`, `FALSE` is represented as `0` and `2` becomes `"2"` (in ASCII numerals preceed letter).

4. __<span style="color:red">Q</span>__: Why is the default missing value, `NA`, a logical vector? What's special about logical vectors? (Hint: think about `c(FALSE, NA_character_)`.)  
   
   __<span style="color:green">A</span>__: It is a practical thought. When you combine `NA`s in `c()` with other atomic types they will be coerced like `TRUE` and `FALSE` to integer `(NA_integer_)`, double `(NA_real_)`, complex `(NA_complex_)` and character `(NA_character_)`. Recall that in R there is a hierarchy of recursion that goes logical >> integer >> double >> character.  If `NA` was a character and provided as part of a set of other values, all of these would be coerced into character as well. Making `NA` a logical means that involving an `NA` in a dataset (which happens often) will not affect coercion.

5. __<span style="color:red">Q</span>__: Precisely what do is.atomic(), is.numeric(), and is.vector() test for?
   
   __<span style="color:green">A</span>__: __`is.atomic()`__ tests if `typeof()` is one of `"logical"`, `"integer"`, `"double"`, `"complex"`, `"character"`, `"raw"` or `"NULL"`.
   
   __`is.numeric()`__ tests if `typeof()` is integer or double and class is not one of `"factor"`, `"Date"`, `"POSIXt"` or `"difftime"`.
   
   __`is.vector()`__ tests if an object has no attributes, except of names and if its `mode()` is atomic (`"logical"`, `"integer"`, `"double"`, `"complex"`, `"character"`, `"raw"`), `"list"` or `"expression"`.

## Attributes

1. __<span style="color:red">Q</span>__: How is `setNames()` implemented? How is `unname()` implemented? Read the source code.
   
   __<span style="color:green">A</span>__: `setNames()` is implemented as:
   
   ```{r}
   setNames
   ```
   
   As the data comes first `setNames()` also works well with the magrittr-pipe operator. When no first argument is given, the result is a named vector:
   
   ```{r}
   setNames( , c("a", "b", "c"))
   ```
   
   However, the implemention also means that `setNames()` only affects the names attribute and not any other more specific naming-attributes like the dimnames attribute for matrices and arrays.
   
   `unname()` is implemented in the following way:
   
   ```{r}
   unname
   ```
   
   If set, `unname()` removes the names and dimnames attributes from its input object. Note that the dimnames attribute (names and row names) won't be affected for data frames, even if the documentation currently (R 3.5.1) mentions this in the case where `force == TRUE` is supplied. Instead the line `dimnames(obj) <- NULL` setting `NULL` as the replacement value in the underlying `dimnames<-.data.frame` function, raises the first error condition in the underlying source code:
   
   ```{r}
   `dimnames<-.data.frame`
   ```
    
2. __<span style="color:red">Q</span>__: What does `dim()` return when applied to a 1d vector? When might you use `NROW()` or `NCOL()`?
    
   __<span style="color:green">A</span>__: From `?nrow`:
    
   > `dim()` will return `NULL` when applied to a 1d vector.
   
   One might want to use `NROW()` or `NCOL()` to handle atomic vectors, lists and NULL values (1 column, 0 rows) analog do a 1-column matrix / data frame. In these cases the alternatives `nrow()` and `ncol()` would return `NULL` (consistently to the behaviour of `dim()`). When subsetting data frames (interactively) this might be convenient as it is not affected by and hence more robust regarding the `drop = TRUE`-idiom:

    ```{r}
    NROW(iris[, 1, drop = TRUE])
    nrow(iris[, 1, drop = TRUE])
    NCOL(iris[, 1, drop = TRUE])
    ncol(iris[, 1, drop = TRUE])
    ```    
3. __<span style="color:red">Q</span>__: How would you describe the following three objects? What makes them different to `1:5`?

    ```{r}
    x1 <- array(1:5, c(1, 1, 5)) # 1 row, 1 column, 5 in third dimension
    x2 <- array(1:5, c(1, 5, 1)) # 1 row, 5 columns, 1 in third dimension
    x3 <- array(1:5, c(5, 1, 1)) # 5 rows, 1 column, 1 in third dimension
    ```
    
   __<span style="color:green">A</span>__: They are of class array and so they have a `dim` attribute.

4. __<span style="color:red">Q</span>__: An early draft used this code to illustrate `structure()`:

    ```{r}
    structure(1:5, comment = "my attribute")
    ```

   But when you print that object you don't see the comment attribute.
   Why? Is the attribute missing, or is there something else special about
   it? (Hint: try using help.) \index{attributes!comment}
    
   __<span style="color:green">A</span>__: From the help of comment `(?comment)`:  
    
   > Contrary to other attributes, the comment is not printed (by print or print.default).
    
    
   Also from the help of attributes `(?attributes)`:  
    
   > Note that some attributes (namely class, comment, dim, dimnames, names, row.names and tsp) are treated specially and have restrictions on the values which can be set.
   
   Apart of that, we can get it easily when we are more specific, i.e.:
   
   ```{r}
   bla <- structure(1:5, comment = "my attribute")
   
   attributes(bla)
   attr(bla, "comment")
   ```
    
## S3 atomic vectors

1. __<span style="color:red">Q</span>__: What sort of object does `table()` return? What is its type? What attributes does it have? How does the dimensionality change as you tabulate more variables?
    
    __<span style="color:green">A</span>__: `table()` returns a crosstabulation of its input. The result is an S3 table object, which is an array (implicit class) of integers (type) under the hood. Attributes are dim (dimension of the underlying array) and dimnames (one for each input column). The dimensionality equals to the number of unique values (accordingly factor levels) of the input arguments, i.e.:
  
    ```{r}
    dim(table(iris))
    sapply(iris, function(x) length(unique(levels(as.factor(x)))))
    ```
    
2. __<span style="color:red">Q</span>__: What happens to a factor when you modify its levels? 
    
    ```{r, results = "none"}
    f1 <- factor(letters)
    levels(f1) <- rev(levels(f1))
    ```
    
   __<span style="color:green">A</span>__: Both, the entries of the factor and also its levels are being reversed:
    
    ```{r}
    f1
    ```
    

3. __<span style="color:red">Q</span>__: What does this code do? How do `f2` and `f3` differ from `f1`?

    ```{r, results = "none"}
    f2 <- rev(factor(letters)) # changes only the entries of the factor
    f3 <- factor(letters, levels = rev(letters)) # changes only the levels of the factor
    ```
    
   __<span style="color:green">A</span>__: Unlike `f1` `f2` and `f3` change only one thing. They change the order of the factor or its levels, but not both at the same time.

## Lists

1. __<span style="color:red">Q</span>__: List all the ways that a list differs from an atomic vector.

   __<span style="color:green">A</span>__: 
   
   * Atomic vectors are homogeneous (all contents must be of the same type). Lists are heterogeneous (all contents can be of different types). 
   
   * Atomic vectors point to one value, while lists contain references which point to one value each:

      ```{r}
      lobstr::ref(1:3)
      lobstr::ref(list(1:3,2,3))
      ```
    
    * Subsetting with out of bound values and `NA`s leads to `NA`s for atomics and `NULL` values for lists:

      ```{r}
      (1:3)[4]
      as.list(1:3)[4]

      (1:3)[NA]
      as.list(1:3)[NA]
      ```
   
2. __<span style="color:red">Q</span>__: Why do you need to use `unlist()` to convert a list to an 
   atomic vector? Why doesn't `as.vector()` work?  
   __<span style="color:green">A</span>__: To get rid of (flatten) the nested structure `as.vector()` doesn't work, because a list is already a vector.

3. __<span style="color:red">Q</span>__: Compare and contrast `c()` and `unlist()` when combining a date and date-time into a single vector.

   __<span style="color:green">A</span>__: Date and date-time objects are build upon doubles. Dates are represented as days, while date-time-objects (POSIXct) represent seconds (counted in regard to the reference date 1970-01-01, also known as "The Epoch").

    ```{r}
    # define date(-time) objects
    date    <- as.Date("1970-01-02")
    dttm_ct <- as.POSIXct("1970-01-01 01:00", tz = "UTC")

    # method-dispatch leads to suprising output whe combining date and date-time
    c(date, dttm_ct)  # equal to c.Date(date, dttm_ct) 
    c(dttm_ct, date)  # equal to c.POSIXct(date, dttm_ct)
    ```

   The generic function dispatches based on the class of its first argument. When `c.Date()` is executed, `dttm_ct` is converted to a date, but the 3600 seconds are mistaken for 3600 days! (And when `c.POSIXct()` is called on `date`, one day counts as one second only.)

    ```{r}
    unclass(c(date, dttm_ct))  # internal representation
    date + 3599
    ```

   Some of these problems may be avoided by explicit conversion of the classes:

    ```{r}
    c(as.Date(dttm_ct, tz = "UTC"), date)
    ```

   Let's look at `unlist()`, which operates on list input.

    ```{r}    
    # attributes are stripped
    unlist(list(date, dttm_ct))  
    ```

   We see again, that internally dates(-times) are doubles. Unfortunately this is all we are left with, when unlist strips the attributes of the list. (This wouldn't happen for vector input, but then we would have to combine the different classes into one vector, which is tricky as seen above.)

   To summarise: `c()` coerces types and errors may occur because of inappropriate method dispatch. `unlist()` strips attributes.
   
<!-- ```{r} -->
<!-- # interestingly combining date and POSIXlt is working -->
<!-- dttm_lt <- as.POSIXlt(dttm_ct) -->
<!-- c.POSIXlt(date, dttm_lt)  # o.k. -->
<!-- ``` -->
<!-- ```{r} -->
<!-- c(date,    dttm_ct) -->
<!-- unlist(c(date,    dttm_ct)) # no difference -->
<!-- ``` -->

## Data frames and tibbles

1. __<span style="color:red">Q</span>__: Can you have a data frame with 0 rows? What about 0 columns?  
   __<span style="color:green">A</span>__: Yes, you can create them easily. Also both dimensions can be 0:

    ```{r}
    # here we use the recycling rules for logical subsetting, but you could
    # also subset with 0, a negative index or a zero length atomic (i.e.
    # logical(0), character(0), integer(0), double(0))
    
    iris[FALSE,]
    
    iris[ , FALSE] # or iris[FALSE]
    
    iris[FALSE, FALSE] # or just data.frame()
    ```

2. __<span style="color:red">Q</span>__: What happens if you attempt to set rownames that are not unique?

    __<span style="color:green">A</span>__
    It depends on the object and how you attempt this. For matrices it will work without any problems. For data frames and using the `row.names<-` replacement function, no further arguments can be set and the underlying `row.names<-.data.frame` method will throw an error (and an additional warning):
   
    ```{r, error = TRUE, warning = TRUE}
    row.names(mtcars) <- rep(1, nrow(mtcars))
    ```
    
    However, by calling the underlying `.rowNamesDF<-` directly you can set the `make.names` argument to `NA`, which sets `seq_len(nrow(x))` as (integer) row names or to `TRUE`, which triggers a call of `make.names(value, unique = TRUE)` for constructing valid (character) row names.

3. __<span style="color:red">Q</span>__: If `df` is a data frame, what can you say about `t(df)`, and `t(t(df))`? Perform some experiments, making sure to try different column types.

    __<span style="color:green">A</span>__ Both will return matrices with dimensions regaring the typical transposition rules. As `t()` uses `as.matrix.data.frame()` for the preprocessing infront of applying `t.default()` and elements of matrices need to be of the same type, all elements get coerced in the usual order (logical << integer << double << character), while factors, dates and datetimes are treated as characters during coercion.

4. __<span style="color:red">Q</span>__: What does `as.matrix()` do when applied to a data frame with 
   columns of different types? How does it differ from `data.matrix()`?
    
   __<span style="color:green">A</span>__: From `?as.matrix`:
    
   > The method for data frames will return a character matrix if there is only atomic columns and any non-(numeric/logical/complex) column, applying as.vector to factors and format to other non-character columns. Otherwise the usual coercion hierarchy (logical < integer < double < complex) will be used, e.g., all-logical data frames will be coerced to a logical matrix, mixed logical-integer will give a integer matrix, etc.
    
   To illustrate this, we create an easy example where the data frame gets coerced to a character matrix:

    ```{r}
    a <- c("a", "b", "c")
    b <- c(TRUE, FALSE, FALSE)
    c <- c("TRUE", "FALSE", "FALSE")
    d <- c(1L, 0L, 2L)
    e <- c(1.5, 2, 3)
    f <- c("one" = 1, "two" = 2, "three" = 3)
    g <- c("first" = 1, "second" = 2, "third" = 3)
    h <- factor(c("f1", "f2", "f3"))
    
    df_cols <- data.frame(a, b, c, d, e, f, g, h, stringsAsFactors = FALSE)
    
    # Note that format is applied to the characters, which can complicate
    # inverse conversion (back to the previous type).
    # For example TRUE in the b variable becomes " TRUE" (starting with a space)
    as.matrix(df_cols)
    ```
    
   From `?as.data.matrix`:
    
   > Return the matrix obtained by converting all the variables in a data frame to numeric mode and then binding them together as the columns of a matrix. Factors and ordered factors are replaced by their internal codes.
    
   So for `data.matrix` we'll get a numerix matrix containing `NA`s for original character columns:
    
    ```{r}
    data.matrix(df_cols)
    ```


## Old exercises

1. __<span style="color:red">Q</span>__: What does `dim()` return when applied to a vector?  
   __<span style="color:green">A</span>__: `NULL`

2. __<span style="color:red">Q</span>__: If `is.matrix(x)` is `TRUE`, what will `is.array(x)` return?  
   __<span style="color:green">A</span>__: `TRUE`, as also documented in `?array`:
    
   > A two-dimensional array is the same thing as a matrix.

3. __<span style="color:red">Q</span>__: What attributes does a data frame possess?  
   __<span style="color:green">A</span>__: names, row.names and class.

4. __<span style="color:red">Q</span>__: What are the six types of atomic vector? How does a list differ from an
   atomic vector?  
   __<span style="color:green">A</span>__: The six types are logical, integer, double, character, complex and raw. The elements of a list
don't have to be of the same type.

5. __<span style="color:red">Q</span>__: What makes `is.vector()` and `is.numeric()` fundamentally different to
   `is.list()` and `is.character()`?  
   __<span style="color:green">A</span>__: The first two tests don't check for a specific type.
   
