# Vectors

## Atomic vectors

1. __<span style="color:red">Q</span>__: How do you create scalars of type raw and complex? (See `?raw` and `?complex`)

   __<span style="color:green">A</span>__: Length 1 raw and complex vectors can be created via `raw()` and `complex()`, i.e.:
  
    ```{r}
    raw(1)
    complex(1)
    ```
  
    For raw vectors it's easiest to coerce numeric or character scalars to raw:
  
    ```{r}
    as.raw(42)
    charToRaw("A")
    ```
  
    For complex numbers real and imaninary parts can be provided directly, i.e.:
  
    ```{r}
    complex(length.out = 1, real = 1, imaginary = 1)
    ```

2. __<span style="color:red">Q</span>__: Test your knowledge of vector coercion rules by predicting the output of
   the following uses of `c()`: 

    ```{r, eval=FALSE}
    c(1, FALSE)      # will be coerced to numeric   -> 1 0
    c("a", 1)        # will be coerced to character -> "a" "1"
    c(TRUE, 1L)      # will be coerced to integer   -> 1 1
    ```

3. __<span style="color:red">Q</span>__: Why is `1 == "1"` true? Why is `-1 < FALSE` true? Why is `"one" < 2` false?  
   __<span style="color:green">A</span>__: These operators are all functions which coerce their arguments (in these cases) to character, double and character. To
   enlighten the latter case: "one" comes after "2" in ASCII.

4. __<span style="color:red">Q</span>__: Why is the default missing value, `NA`, a logical vector? What's special
about logical vectors? (Hint: think about `c(FALSE, NA_character_)`.)  
   
   __<span style="color:green">A</span>__: It is a practical thought. When you combine `NA`s in `c()` with other atomic types they will be coerced
like `TRUE` and `FALSE` to integer `(NA_integer_)`, double `(NA_real_)`, complex `(NA_complex_)` and character `(NA_character_)`. Recall that in R there is a hierarchy of recursion that
goes logical -> integer -> double -> character.  If `NA` were, for example, a character,
including `NA` in a set of integers or logicals would result in them getting coerced
to characters which would be undesirable. Making `NA` a logical means that involving an `NA`
in a dataset (which happens often) will not result in coercion.

5. __<span style="color:red">Q</span>__: Precisely what do is.atomic(), is.numeric(), and is.vector() test for?
   
   __<span style="color:green">A</span>__: 
   
   __is.atomic__: Tests if `typeof()` is one of `"logical"`, `"integer"`, `"double"`, `"complex"`, `"character"`, `"raw"` or `"NULL"`.
   
   __is.numeric__: Tests if `typeof()` is integer or double and class is not one of `"factor"`, `"Date"`, `"POSIXt"` or `"difftime"`.
   
   __is.vector__: Tests if and object has no attributes, except for names and its `mode()` is atomic (`"logical"`, `"integer"`, `"double"`, `"complex"`, `"character"`, `"raw"`), `"list"` or `"expression"`.

## Attributes

1. __<span style="color:red">Q</span>__: How is setNames() implemented? How is unname() implemented? Read the source code.
   
   __<span style="color:green">A</span>__: `setNames()` is implemented as:
   
   ```{r}
   setNames
   ```
   
   As the data comes first `setNames()` also works well with the magrittr-pipe operator. When no first argument is given, the result is a named vector:
   
   ```{r}
   setNames( , c("a", "b", "c"))
   ```
   
   However, the implemention also means that `setNames()` only affects the names attribute and not any ohter more specific naming-attributes like the dimnames attribute for matrices and arrays.
   
   `unname()` is implemented in the following way:
   
   ```{r}
   unname
   ```
   
   It always removes names and dimnames attributes when they are set, except for data frames. Note that the documentation is currently not correct (R 3.5.1). `force == TRUE` doesn't remove the dimnames (names and row names) as `NULL` is always catched within the first condition of the underlying `dimnames<-.data.frame`:
   
   ```{r}
   `dimnames<-.data.frame`
   ```
    
2. __<span style="color:red">Q</span>__: What does dim() return when applied to a 1d vector? When might you use NROW() or NCOL()?
    
   __<span style="color:green">A</span>__: From `?nrow`:
    
   `dim()` will return `NULL` when appleid to a 1d vector.
   
   One might want to use `NROW()` or `NCOL()` to handle atomic vectors, lists and NULL values (1 column, 0 rows) analog do a 1-column matrix / data frame. In these cases the alternatives `nrow()` and `ncol()` would return `NULL` (consistently to the behaviour of `dim()`). When subsetting data frames (interactively) this might be convenient as it is not affected by and hence more robust regarding the `drop = TRUE`-idiom:

    ```{r}
    NROW(iris[, 1, drop = TRUE])
    nrow(iris[, 1, drop = TRUE])
    NCOL(iris[, 1, drop = TRUE])
    ncol(iris[, 1, drop = TRUE])
    ```    
3. __<span style="color:red">Q</span>__: How would you describe the following three objects? What makes them different to `1:5`?

    ```{r}
    x1 <- array(1:5, c(1, 1, 5)) # 1 row, 1 column, 5 in third dimension
    x2 <- array(1:5, c(1, 5, 1)) # 1 row, 5 columns, 1 in third dimension
    x3 <- array(1:5, c(5, 1, 1)) # 5 rows, 1 column, 1 in third dimension
    ```
    
   __<span style="color:green">A</span>__: They are of class array and so they have a `dim` attribute.

4. __<span style="color:red">Q</span>__: An early draft used this code to illustrate `structure()`:

    ```{r}
    structure(1:5, comment = "my attribute")
    ```

   But when you print that object you don't see the comment attribute.
   Why? Is the attribute missing, or is there something else special about
   it? (Hint: try using help.) \index{attributes!comment}
    
   __<span style="color:green">A</span>__: From the help of comment `(?comment)`:  
    
   > Contrary to other attributes, the comment is not printed (by print or print.default).
    
    
   Also from the help of attributes `(?attributes)`:  
    
   > Note that some attributes (namely class, comment, dim, dimnames, names, row.names and tsp) are treated specially and have restrictions on the values which can be set.
    
## S3 atomic vectors

5. __<span style="color:red">Q</span>__: What sort of object does `table()` return? What is its type? What attributes does it have? How does the dimensionality change as you tabulate more variables?
    
  __<span style="color:green">A</span>__: 
   
    ```{r}
    x <- NULL
    # For NULL
    table(x) %>% sloop::otype()
    # Table kreiert ein S3-Object
    table(x) %>% class()
    # der Klasse table
    table(x) %>% typeof()
    # basierend auf dem type integer
    table(x) %>% attributes()
    # mit den attributen
    # - dim
    # - dimnames (x)
    # - class 

    x <- c("a", "b", "c")
    # For NULL
    table(x) %>% sloop::otype()
    # Table kreiert ein S3-Object
    table(x) %>% class()
    # der Klasse table
    table(x) %>% typeof()
    # basierend auf dem type integer
    table(x) %>% attributes()
    # mit den attributen
    # - dim
    # - dimnames (x)
    # - class 
 
    table(iris$Species) %>% unclass
    # Returns an array
    rm(x)

    # Need to read about table and unclass and tablate
    ```
    
6. __<span style="color:red">Q</span>__: What happens to a factor when you modify its levels? 
    
    ```{r, results = "none"}
    f1 <- factor(letters)
    levels(f1) <- rev(levels(f1))
    ```
    
   __<span style="color:green">A</span>__: Both, the entries of the factor and also its levels are being reversed:
    
    ```{r}
    f1
    ```
    

7. __<span style="color:red">Q</span>__: What does this code do? How do `f2` and `f3` differ from `f1`?

    ```{r, results = "none"}
    f2 <- rev(factor(letters)) # changes only the entries of the factor
    f3 <- factor(letters, levels = rev(letters)) # changes only the levels of the factor
    ```
    
   __<span style="color:green">A</span>__: Unlike `f1` `f2` and `f3` change only one thing. They change the order of the factor or its levels, but not both at the same time.

## Lists

1. __<span style="color:red">Q</span>__: List all the ways that a list differs from an atomic vector.
   __<span style="color:green">A</span>__: 
    
2. __<span style="color:red">Q</span>__: Why do you need to use `unlist()` to convert a list to an 
   atomic vector? Why doesn't `as.vector()` work?  
   __<span style="color:green">A</span>__: To get rid of (flatten) the nested structure. `as.vector()` doesn't work, because a list is already a vector.

3. __<span style="color:red">Q</span>__: Compare and contrast `c()` and `unlist()` when combining a date and date-time into a single vector.

## Data frames and tibbles

1. __<span style="color:red">Q</span>__: Can you have a data frame with 0 rows? What about 0 columns?  
   __<span style="color:green">A</span>__: Yes, you can create them easily. Also both dimensions can be 0:

    ```{r}
    # here we use the recycling rules for logical subsetting, but you could
    # also subset with 0, a negative index or a zero length atomic (i.e.
    # logical(0), character(0), integer(0), double(0))
    
    iris[FALSE,]
    
    iris[ , FALSE] # or iris[FALSE]
    
    iris[FALSE, FALSE] # or just data.frame()
    ```

2. __<span style="color:red">Q</span>__: What happens if you attempt to set rownames that are not unique?

3. __<span style="color:red">Q</span>__: If `df` is a data frame, what can you say about `t(df)`, and `t(t(df))`? Perform some experiments, making sure to try different column types.

4. __<span style="color:red">Q</span>__: What does `as.matrix()` do when applied to a data frame with 
   columns of different types? How does it differ from `data.matrix()`?
    
   __<span style="color:green">A</span>__: From `?as.matrix`:
    
   > The method for data frames will return a character matrix if there is only atomic columns and any non-(numeric/logical/complex) column, applying as.vector to factors and format to other non-character columns. Otherwise the usual coercion hierarchy (logical < integer < double < complex) will be used, e.g., all-logical data frames will be coerced to a logical matrix, mixed logical-integer will give a integer matrix, etc.
    
   To illustrate this, we create an easy example where the data frame gets coerced to a character matrix:

    ```{r}
    a <- c("a", "b", "c")
    b <- c(TRUE, FALSE, FALSE)
    c <- c("TRUE", "FALSE", "FALSE")
    d <- c(1L, 0L, 2L)
    e <- c(1.5, 2, 3)
    f <- c("one" = 1, "two" = 2, "three" = 3)
    g <- c("first" = 1, "second" = 2, "third" = 3)
    h <- factor(c("f1", "f2", "f3"))
    
    df_cols <- data.frame(a, b, c, d, e, f, g, h, stringsAsFactors = FALSE)
    
    # Note that format is applied to the characters, which can complicate
    # inverse conversion (back to the previous type).
    # For example TRUE in the b variable becomes " TRUE" (starting with a space)
    as.matrix(df_cols)
    ```
    
   From `?as.data.matrix`:
    
   > Return the matrix obtained by converting all the variables in a data frame to numeric mode and then binding them together as the columns of a matrix. Factors and ordered factors are replaced by their internal codes.
    
   So for `data.matrix` we'll get a numerix matrix containing `NA`s for original character columns:
    
    ```{r}
    data.matrix(df_cols)
    ```


## Old exercises

1. __<span style="color:red">Q</span>__: What does `dim()` return when applied to a vector?  
   __<span style="color:green">A</span>__: `NULL`

2. __<span style="color:red">Q</span>__: If `is.matrix(x)` is `TRUE`, what will `is.array(x)` return?  
   __<span style="color:green">A</span>__: `TRUE`, as also documented in `?array`:
    
   > A two-dimensional array is the same thing as a matrix.

3. __<span style="color:red">Q</span>__: What attributes does a data frame possess?  
   __<span style="color:green">A</span>__: names, row.names and class.

4. __<span style="color:red">Q</span>__: What are the six types of atomic vector? How does a list differ from an
   atomic vector?  
   __<span style="color:green">A</span>__: The six types are logical, integer, double, character, complex and raw. The elements of a list
don't have to be of the same type.

5. __<span style="color:red">Q</span>__: What makes `is.vector()` and `is.numeric()` fundamentally different to
   `is.list()` and `is.character()`?  
   __<span style="color:green">A</span>__: The first two tests don't check for a specific type.
   
