# Functions

## Function fundamentals

1. __<span style="color:red">Q</span>__: Given a name, like `"mean"`, `match.fun()` lets you find a function. Given a function, can you find its name? Why doesn't that make sense in R?  

   __<span style="color:green">A</span>__: If you know `body()`, `formals()` and `environment()` it can be possible to find the function. However, this approach won't be work for primitive functions, because they return `NULL` for these three properties. Also annonymous functions won't be found, because they are not bound to a name. It is also possible that different names in an environment contain bindings to one (or more functions) with the same `body()`, `formals()` and `environment()` properties. This means that the search wouldn't be unique <!-- eindeutig -->.

   More general: A (function) name is always bound to an object, but an object (i.e. a function) doesn't have a name (just a binding sometimes). <!-- @hb: look this up again -->
   
   Also: You "can only find the parents of an environment, not its children, so this is necessarily only going to be a partial search" (Hadley Wickham). This means, that some function names in child environments will not be found. <!-- check this. -->

2. __<span style="color:red">Q</span>__: It’s possible (although typically not useful) to call an anonymous function. Which of the two approaches below is correct? Why?
   
    ```{r}
    function(x) 3()
    (function(x) 3)()
    ```

   __<span style="color:green">A</span>__: The second approach is correct. In this approach the anonymous function `function(x) 3` is surrounded by a pair of brackets and followed by `()` which calls the former `(function(x) 3)`. The extra syntax is needed as otherwise the following `()` belong to the body of the anonymous function. Therefore, the expression in the first approach returns a function with the invalid body `3()` and we get an error when we call it:
   
    ```{r, error = TRUE}
    (function(x) 3())()
    ```
   
3. __<span style="color:red">Q</span>__: A good rule of thumb is that an anonymous function should fit on one line and shouldn't need to use `{}`. Review your code. Where could you have used an anonymous function instead of a named function? Where should you have used a named function instead of an anonymous function?  
    
   __<span style="color:green">A</span>__: The use of anonymous functions may lead to more concise and elegant code in certain situations. However, they miss a descriptive name and it can take a while to figure out what they do. That's why it's really helpful to give long and complex functions a descriptive name. You may take a look at your own projects (or some other online projects) to reflect your own style.

<!-- @hb: maybe describe advantages of anonymous functions more clearly. -->

4. __<span style="color:red">Q</span>__: What function allows you to tell if an object is a function? What function allows you to tell if a function is a primitive function?
       
   __<span style="color:green">A</span>__: To find out, if an object is a function `is.function()` can be used. To specifically check, if an object is a primitive function use `is.primitive()`.

5. __<span style="color:red">Q</span>__: This code makes a list of all functions in the base package. 
    
    ```{r}
    objs <- mget(ls("package:base"), inherits = TRUE)
    funs <- Filter(is.function, objs)
    ```
    
   Use it to answer the following questions:
    
   a. Which base function has the most arguments?
    
   a. How many base functions have no arguments? What's special about those functions?
   
   a. How could you adapt the code to find all primitive functions?  
    
   __<span style="color:green">A</span>__: Let's look at each subquestion seperately:
   
   a. To find the function with the most arguments, we pipe the list of functions into a chain of iterations. First, we create a list of each function's arguments. Then we count the arguments for each function. Finally, we subset the result with its maximum element, returning us the name of the function with the most arguments and the corresponding number of arguments:
   
      ```{r}
      library(magrittr)
      
      funs %>% 
        sapply(formals) %>%
        lengths() %>% 
        `[`(which.max(.))
      ```
      
   b. To find the number of 0-argument functions, we count the number of arguments again for each function. However, this time we check each count for beeing equal to `0` and sum up the result.
   
      ```{r}
      funs %>% sapply(formals) %>% 
        lengths() %>%
        `==`(0L) %>%
        sum()
      ```
     As we already know from the textbook that `formals()` returns `NULL` for primitive functions, all of them are included in our count of 0-argument functions. To check how this affects our count, let's create a table of 0-argument functions, dividing them into primitive and non-primitive functions. Therefore, we filter the function list for 0-argument functions first. Then we check each of these functions for beeing primitive or not and finally create our table:
     
      ```{r}
      funs %>%
        Filter(function(x) length(formals(x)) == 0, .) %>% 
        sapply(is.primitive) %>% 
        table()
      ```
     Indeed, most of the 0-argument functions in base R are primitive functions.
     
   c. To adapt the code in `Filter()` to find all primitive functions, we can simply change the predicate from `is.function()` to `is.primitive()`:
   
      ```{r, eval = FALSE}
      funs <- Filter(is.primitive, objs)
      ```
      
6. __<span style="color:red">Q</span>__: What are the three important components of a function?  
    
   __<span style="color:green">A</span>__: Thse components are the function's `body()`, `formals()` and `environment()`. However, as mentioned in the textbook:
    
   > There is one exception to the rule that functions have three components. Primitive functions, like `sum()`, call C code directly with `.Primitive()` and contain no R code. Therefore their `formals()`, `body()`, and `environment()` are all `NULL`.

7. __<span style="color:red">Q</span>__: When does printing a function not show what environment it was created in?  

   __<span style="color:green">A</span>__: When it is a primitive function or when it was created in the global environment.

## Lexical Scoping

1. __<span style="color:red">Q</span>__: What does the following code return? Why? Describe how each of the three `c`’s is interpreted.

    ```{r, eval = FALSE}
    c <- 10
    c(c = c)
    ```  
    
   __<span style="color:green">A</span>__: A named vector `c`, which first element has the value `10` and the name `"c"`. The first "c" is the `c()` function, the second is the name of the first element and the third is the value of the first element.
       
2. __<span style="color:red">Q</span>__: What are the four principles that govern how R looks for values?  
    
   __<span style="color:green">A</span>__: The four principles are those behind R's implementation of lexical scoping:
   
   * name masking
   * functions vs. variables
   * a fresh start
   * dynamic lookup

3. __<span style="color:red">Q</span>__: What does the following function return? Make a prediction before running the code yourself.

    ```{r, eval = FALSE}
    f <- function(x) {
      f <- function(x) {
        f <- function(x) {
          x ^ 2
        }
        f(x) + 1
      }
      f(x) * 2
    }
    f(10)
    ```
        
   __<span style="color:green">A</span>__: 202
      
## Lazy evaluation

1. __<span style="color:red">Q</span>__: What important property of `&&` makes `x_ok()` work?

    ```{r}
    x_ok <- function(x) {
      !is.null(x) && length(x) == 1 && x > 0
    }
    
    x_ok(NULL)
    x_ok(1)
    x_ok(1:3)
    ```

   What is different with this code? Why is this behaviour undesirable here?
       
    ```{r}
    x_ok <- function(x) {
      !is.null(x) & length(x) == 1 & x > 0
    }
    
    x_ok(NULL)
    x_ok(1)
    x_ok(1:3)
    ```
    
    __<span style="color:green">A</span>__: The intended usage of `x_ok()` is to check if an argument provided to a function is not `NULL`, has length `1` and is greater than `0`. To work with this function, we only want to know if this is `TRUE`, `FALSE` or `NA` (unknown). Therefore, the first version behaves as desired.
    
    To highlight the relevant consequences of lazy evaluation for this example, let's look at the provided test cases and compare both versions of `x_ok()` for each of them.
    
    - **`x = NULL`:** As `is.null(x)` returns `FALSE` also `&&()` returns `FALSE`. Because of lazy evaluation the other arguments of `&&()` are not even evaluated. This makes sense as when one part of a logical AND conjunction is `FALSE`, the whole logical conjunction must be `FALSE`. In contrast the `&()` function evaluates all of its arguments. Especially as the last part of `x_ok()`, `x > 0`, evaluates to `logical(0)`, it becomes relevant how the results are combined. As `&()` combines `FALSE` and `logical(0)` to `logical(0)` the `&`-version of `x_ok()` doesn't provide a desired result in this case.
    
    - **`x = 1`:** As all statements evaluate to `TRUE`. Both versions of `x_ok()` combine these into the desired output of `TRUE`.
    
    - **`x = 1:3`:** When it comes to elementwise evaluation `&&()` and `&()` differ fundamentally. While `&&()` compares only the first elements of its arguments, `&()` compares its arguments elementwise and even recycles shorter vectors, if necessary. As the `length(x) == 1` condition is `FALSE`, the `&&`-version of `x_ok()` doesn't even evaluate the last condition, `x > 0`, and returns `FALSE` immediately. In contrast, the `&`-version evaluates all of its arguments and since `x > 0` evaluates to `c(FALSE, FALSE, FALSE)`, this is also the result (due to recycling and elementwise comparison), which is - again - not as desired.

2. __<span style="color:red">Q</span>__: What does this function return? Why? Which principle does it illustrate?

    ```{r, eval = FALSE}
    f2 <- function(x = z) {
      z <- 100
      x
    }
    f2()
    ```  
    
   __<span style="color:green">A</span>__: 100, lazy evaluation.
    
3. __<span style="color:red">Q</span>__: What does this function return? Why? Which principle does it illustrate?
  
    ```{r}
    y <- 10
    f1 <- function(x = {y <- 1; 2}, y = 0) {
      c(x, y)
    }
    f1()
    y
    ```  
    
   __<span style="color:orange">A</span>__: The output of `f1()` shows, that the default argument for `y` is not accessed. Instead the value assigned to `y` in the expression set as the default argument for `x` is used. Also, calling `f1` doesn't change value of `y` in the global environment.

    ```{r}
    f1()
    y
    ```

   The principle here is "value caching of function arguments": R caches these arguments when their value is accessed for the first time. Here `x` is called before `y` (because of `c(x, y)`). When the default expression given to `x` is evaluated, a value for `y` is also created (and cached). Therefore the default argument for `y` will no longer be evaluated.

   If we change the functions body to `c(y, x)`, the default value given for `y` will be used.

    ```{r}
    f2 <- function(x = {y <- 1; 2}, y = 0) {
      c(y, x)
    }
    f2()
    ```

4. __<span style="color:red">Q</span>__: In `hist()`, the default value of `xlim` is `range(breaks)`, the default value for `breaks` is `"Sturges"`, and

    ```{r}
    range("Sturges")
    ```
    
   Explain how `hist()` works to get a correct `xlim` value.
    
   __<span style="color:green">A</span>__: Before the `hist()` function creates the final plot where `xlim` is provided and finally evaluated, the `hist()` function internally updates and checks the value of `breaks` several times to ensure that it is finally a numeric vector with at least two elements.
    
   The detailed behaviour is very specific to the input. According to `?hist` this must be one of:
   * a vector giving the breakpoints between histogram cells,
   * a function to compute the vector of breakpoints,
   * a single number giving the number of cells for the histogram,
   * a character string naming an algorithm to compute the number of cells (see ‘Details’),
   * a function to compute the number of cells.
    
   Further:
    
   > In the last three cases the number is a suggestion only; as the breakpoints will be set to pretty values, the number is limited to 1e6 (with a warning if it was larger). If breaks is a function, the x vector is supplied to it as the only argument (and the number of breaks is only limited by the amount of available memory).
    
   In case of `breaks = "Sturges"` this means that `breaks` is:
   * checked that it is provided (a corresponding flag is set; otherwise it is set to `nclass` regarding that is provided)
   * checked that its length is greater than 1 (a corresponding flag is set)
   * converted to lower case and matched to "sturges"
   * set to an integer value via `sturges = nclass.Sturges(x)` inside a `switch()` statement
   * checked that it is now numeric, finite and at least 1
   * set to 1000000 if it is greater than 1000000
   * turned into a numeric vector of length (possibly) greater one via `pretty(range(x), n = breaks, min.n = 1)`
   * checked that its length is now greater than 1 and not `NA`
   * checked that the differences between the breaks are strictly positive
   * during and after this process more variables are calculated and checks are made and finally `plot()` gets called, where `xlim` gets evaluated.
    
5. __<span style="color:red">Q</span>__: Explain why this function works. Why is it confusing?

    ```{r}
    show_time <- function(x = stop("Error!")) {
      stop <- function(...) Sys.time()
      print(x)
    }
    show_time()
    ```
    
   __<span style="color:green">A</span>__: It works because functions are objects, that can be modified and overwritten, and because of of lazy evaluation. Before the `x` argument is evaluated, the `stop()` function is overwritten by another function which gets called in the last line where `x` is finally evaluated.
    
   It's quite confusing, since there is no relation between the default value of `x` and its actual meaning in the context of the `show_time()` function. The user won't have any chance to guess the meaning of `x` or it's default value without looking up a possibly written documentation or analyzing of the source code.

6. __<span style="color:red">Q</span>__: How many arguments are required when calling `library()`?

   __<span style="color:green">A</span>__: Surprisingly no argument is required. When looking at `?library` we can see under usage, that library has nine arguments and two of them are without default arguments:
    
   library(package, help, pos = 2, lib.loc = NULL,
        character.only = FALSE, logical.return = FALSE,
        warn.conflicts = TRUE, quietly = FALSE,
        verbose = getOption("verbose"))
        
   However, when we call `library()` without any arguments, we get a list of all available libraries under the current library path (`.libPaths()`) as also document under the details section of the help file:
    
   > If library is called with no package or help argument, it lists all available packages in the libraries specified by lib.loc, and returns the corresponding information in an object of class "libraryIQR". (The structure of this class may change in future versions.) Use .packages(all = TRUE) to obtain just the names of all available packages, and installed.packages() for even more information.

## `...` (dot-dot-dot)

1. __<span style="color:red">Q</span>__: Explain the following results:
    
    ```{r}
    sum(1, 2, 3)
    mean(1, 2, 3)
    
    sum(1, 2, 3, na.omit = TRUE)
    mean(1, 2, 3, na.omit = TRUE)
    ```
    
   __<span style="color:green">A</span>__: The arguments of `sum()` are `...` and `na.rm`. For `...` `sum()` expects "numeric or complex or logical vectors" as documented in `?sum`. So any input not explicitly supplied named with `na.rm` is treated as part of the `...` argument and used for summation.
    
   In contrast `mean()` expects as first argument `x` typically a vector, as second argument `trim` a fraction of observations to be trimmed from each end of `x` and again `na.rm`. As both: `trim = 2` and `na.rm = 3` have not effect on the calculation of the mean of `1`, we get `1` as the result.
    
   In the next call, `na.omit` is supplied via the `...` argument to `sum()`, which treats it as logical vector and builds its sum with the other arguments.
    
   Finally in the last call to `mean` `na.omit = TRUE` is neither needed as part of the default method nor is it used for the mean calculation, since `mean()` calculates it's value only from its `x` argument.

2. __<span style="color:red">Q</span>__: In the following call, explain how to find the documentation for the named arguments in the following function call:
       
    ```{r, fig.asp = 1}
    plot(1:10, col = "red", pch = 20, xlab = "x", col.lab = "blue")
    ```
    
   __<span style="color:green">A</span>__: First we type `?plot` in the console and scan the usage section:
    
    ```
    plot(x, y, ...)
    ```
    
   Obviously we have to look under the `...` bullet in the arguments section.
    
   There we can find a bullet for `xlab` (check), and follow the recommendation to visit `?par` for further arguments. 
    
   From there we type "col" into the search bar, which leads us to a recommentation to search further under `Color Specification` (check). Again using the search we find a bullet for the `pch` argument. From there we get also the recommendation to look under `?points` for more specific settings (check). Finally we use the search functionality to find `col.lab` also as a bullet inside `?par`.
    
3. __<span style="color:red">Q</span>__: Why does `plot(1:10, col = "red")` only colour the points, not the axes 
    or labels? Read the source code of `plot.default()` to find out.
    
   __<span style="color:green">A</span>__: It is easiest to start by adding `browser()` to the first line of `plot.default()` and interactively run `plot(1:10, col = "red")`. In this way we can see how the plot is build during the last lines and especially find out where the axis are added. This leads us to the function call:

    ```{r, eval = FALSE}
    localTitle(main = main, sub = sub, xlab = xlab, ylab = ylab, ...)
    ```
    
   The `localTitle()` function was defined in the first lines of `plot.default()` as:

    ```{r, eval = FALSE}
    localTitle <- function(..., col, bg, pch, cex, lty, lwd) title(...)
    ```
    
   So the call to `localTitle()` clearly gets the `col` parameter as part of `...` argument. To find out if it is used we try following the source code of `title()`, which leads us to a line of C code. Instead of following further, can we stay in R and look at `?title`, which brings some clarity on the fact that the `title()` function specifies four parts of the plot: Main (title of the plot), sub (sub-title of the plot) and both axis labels. Therefore it would introduce ambiguity inside `title()` to use `col` directly. Instead on has the option to supply `col` via the `...` argument as `col.labs` or as part of `xlab` (similar for `ylab`) in the form `xlab = list(c("index"), col = "red")`.

## Exiting a function

1. __<span style="color:red">Q</span>__: What does `load()` return? Why don’t you normally see these values?

   __<span style="color:green">A</span>__: `load()` reloads datasets written with the function `save()`. It returns a character vector of the names of objects created, invisibly. To see the names of the objects, one can set the `verbose` argument to `TRUE`, which triggers a regarding `if` statement in the function's body. However, to print the value of the names it is also feasible to use brackets around the `load()` call to autoprint the returned value.
   
2. __<span style="color:red">Q</span>__: What does `write.table()` return? What would be more useful?

   __<span style="color:green">A</span>__: It invisibly returns `NULL`. It would be more useful to invisibly return the (data frame) object to be written as for example the **readr** package does. In this way it would be possible to save intermediate results from a sequence of processing steps directly, i.e. within a magrittr pipeline.
    
3. __<span style="color:red">Q</span>__: How does the `chdir` parameter of `source()` compare to `in_dir()`? Why might you prefer one approach to the other?

The `in_dir()` approach was given in the book as
       
    ```{r, eval = FALSE}
    in_dir <- function(dir, code) {
      old <- setwd(dir)
      on.exit(setwd(old))
      
      force(code)
      }
    ```
    
   __<span style="color:green">A</span>__: `in_dir()` takes a path to a working directory as an argument. At the beginning of the function the working directory is changed to this specification and with a call to `on.exit` it is guranteed, that when the function finishes the working directory also equals to this specification.
    
   In `source()` you need the `chdir` argument to specify, if the working directory should be changed during the evaluation to the `file` argument, if this is a pathname. The difference in `source()` is, that the actual working directory as output of `getwd()` is saved to set it in `on.exit` before changing the directory to the pathname (given to the `file` argument) for the rest of the execution of the `source()` function.

4. __<span style="color:red">Q</span>__: Write a function that opens a graphics device, runs the supplied code, and closes the graphics device (always, regardless of whether or not the plotting code worked).  
   
   __<span style="color:green">A</span>__: 
    
    ```{r, eval = FALSE}
    plot_pdf <- function(code){
      pdf("test.pdf")
      on.exit(dev.off())
      code
      }
    ```

5. __<span style="color:red">Q</span>__: We can use `on.exit()` to implement a simple version of `capture.output()`.

    ```{r}
    capture.output2 <- function(code) {
      temp <- tempfile()
      on.exit(file.remove(temp), add = TRUE)

      sink(temp)
      on.exit(sink(), add = TRUE)

      force(code)
      readLines(temp)
    }
    capture.output2(cat("a", "b", "c", sep = "\n"))
    ```
    
   Compare `capture.output()` to `capture.output2()`. How do the functions differ? What features have I removed to make the key ideas easier to see? How have I rewritten the key ideas to be easier to understand?
    
   __<span style="color:green">A</span>__: Using `body(capture.output)`, we can see the source code for the original `capture.output()` function. `capture.output()` is a good clip longer (39 lines vs. 7 lines). The reason for this is that `capture.output2()` is more modular, since `capture.output()` writes out entire methods like `readLines()` instead of invoking them. This makes `capture.output2` easier to understand if you understand the underlying methods.
    
   However, `capture.output2()` does remove potentially important functionality, as `capture.output()` appears to handle important exceptions not handled in `capture.output2()`, and `capture.output()` offers the ability to chose between overwriting or appending to a file.

## Function forms

1. __<span style="color:red">Q</span>__: Rewrite the following code snippets into prefix form:

    ```{r, eval = FALSE}
    1 + 2 + 3
    
    1 + (2 + 3)
    
    if (length(x) <= 5) x[[5]] else x[[n]]
    ```
    
   __<span style="color:green">A</span>__: 
    
    ```{r, eval = FALSE}
    `+`(`+`(1, 2), 3)
    
    `+`(1, `(`(`+`(2, 3)))
    
    `if`(`<=`(length(x), 5), `[[`(x, 5), `[[`(x, n))
    ```

2. __<span style="color:red">Q</span>__: Clarify the following list of odd function calls:

    ```{r, eval = FALSE}
    x <- sample(replace = TRUE, 20, x = c(1:10, NA))
    # -> sample(x = c(1:10, NA), size = 20, replace = TRUE)
    y <- runif(min = 0, max = 1, 20)
    # -> runif(n = 20, min = 0, max = 1)
    cor(m = "k", y = y, u = "p", x = x)
    # -> cor(x = x, y = y, use = "pairwise.complete.obs", method = "pearson")
    ```  
    
3. __<span style="color:red">Q</span>__: Explain why the following code fails:

    ```{r, eval = FALSE}
    modify(get("x"), 1) <- 10
    #> Error: target of assignment expands to non-language object
    ```
    
   __<span style="color:green">A</span>__: First let define `x` and recall the definition of `modify()` from the textbook:
    
    ```{r}
    x <- 1:3
    
    `modify<-` <- function(x, position, value) {
      x[position] <- value
      x
    }
    ```
    
   As described in the textbook R turns the code behind the scenes into
    
    ```{r, eval = FALSE}
    get("x") <- `modify<-`(get("x"), 1, 10)
    #> Error in get("x") <- `modify<-`(get("x"), 1, 10) : target of assignment expands to non-language object
    ```
    
   which can not work, because `get()` has no equivalent replacement function. To confirm this claim, we can reproduce the error via an easier example
    
    ```{r, eval = FALSE}
    get("x") <- 2
    #> Error in get("x") <- 2 : target of assignment expands to non-language object
    ```
    
   and modify the example to use a function with an available replacement function:
    
    ```{r}
    modify(names(x), 1) <- 10
    names(x)
    ```
    
4. __<span style="color:red">Q</span>__: Create a replacement function that modifies a random location in a vector.
    
   __<span style="color:green">A</span>__: 

    ```{r, eval = FALSE}
    `random<-` <- function(x, value){
      x[sample(length(x), 1)] <- value
      x
      }
    ```

5. __<span style="color:red">Q</span>__: Write your own version of `+` that will paste its inputs together if they are character vectors but behaves as usual otherwise. In other words, make this code work:
   
    ```{r, eval = FALSE}
    1 + 2
    #> [1] 3
    
    "a" + "b"
    #> [1] "ab"
    ```

   __<span style="color:green">A</span>__: We can simply override the `+` operator. In this case we need to take a bit of care to not use the `+` operator itself inside of the function definition, since otherwise we would end in an infinite recursion (a special case of an infinite loop). We also add `b = 0L` as a default value keep the behaviour of `+` as a unary operator, i.e. to keep `+ 1` working and not throwing an error:
    
    ```{r}
    `+` <- function(a, b = 0L){
      if (is.character(a) && is.character(b)) {return(paste0(a, b))}
      a -- b
    }
    
    # tests
    + 1
    
    1 + 2
    
    "a" + "b"
    
    # return back to the original `+` operator behaviour
    rm(`+`)
    ```

6. __<span style="color:red">Q</span>__: Create a list of all the replacement functions found in the base package.
   Which ones are primitive functions? (Hint use `apropros()`)
   
   __<span style="color:green">A</span>__: We can find replacementfunctions by searching for functions that end on "<-":
   
    ```{r, eval = FALSE}
    repls <- funs[grepl("<-$", names(funs))]
    Filter(is.primitive, repls)
    ```

7. __<span style="color:red">Q</span>__: What are valid names for user-created infix functions?  
  
   __<span style="color:green">A</span>__: As the section on "Function Forms" in Advanced R tells us, the "names of infix functions are more flexible than regular R functions: they can contain any sequence of characters except “%”."

8. __<span style="color:red">Q</span>__: Create an infix `xor()` operator.
    
   __<span style="color:green">A</span>__: 

    ```{r}
    `%xor_%` <- function(a, b){
      (a | b) & !(a & b)
    }
    ```

9. __<span style="color:red">Q</span>__: Create infix versions of the set functions `intersect()`, `union()`, and
   `setdiff()`. You might call them `%n%`, `%u%`, and `%/%` to match 
   conventions from mathematics.
   
   __<span style="color:green">A</span>__: 
   
    ```{r}
    `%union_%` <- function(a, b){
      unique(c(a, b))
      }

    `%intersect_%` <- function(a, b){
      unique(c(a[a %in% b], b[b %in% a]))
      }

    `%setdiff_%` <- function(a, b){
      a[!a %in% b]
      }
    ```
      
## Old exercises

1. __<span style="color:red">Q</span>__: What does this function return? Why? Which principle does it illustrate?
  
    ```{r include=FALSE, echo=TRUE}
    f1 <- function(x = {y <- 1; 2}, y = 0) {
      x + y
    }
    f1()
    ```  
    
   __<span style="color:green">A</span>__: It returns 3 and illustrates lazy evaluation. As you can see, y becomes 1, but only when `x` is evaluated (before `y`) inside the function (otherwise it is 0):
    
    ```{r}    
    f2 <- function(x = {y <- 1; 2}, y = 0) {
      y
    }
    f2()
    ```  
    
   Note that funny things can happen if we switch the evaluation order (even within one line)
    
    ```{r}    
    f3 <- function(x = {y <- 1; 2}, y = 0) {
      y + x
    }
    f3()
    ```

   or we evaluate `y` once before and once after the evaluation of `x`
    
    ```{r}    
    f4 <- function(x = {y <- 1; 2}, y = 0) {
      y_before_x <- y
      x
      y_after_x <- y
      c(y_before_x, y_after_x)
    }
    f4()
    ```
