# Control flow

## Choices

1. __<span style="color:red">Q</span>__: What type of vector does each of the following calls to `ifelse()`
    return?

    ```{r, eval = FALSE}
    ifelse(TRUE, 1, "no")
    ifelse(FALSE, 1, "no")
    ifelse(NA, 1, "no")
    ```

    Read the documentation and write down the rules in your own words.

   __<span style="color:green">A</span>__: The arguments of `ifelse()` are `test`, `yes` and `no`. When `yes` or `no` are shorter than `test`, they are getting recycled. When they are longer than `test`, their additional elements get ignored. `ifelse()` operates elementwise and returns the entry from `yes` (when `test` is `TRUE`), `no` (when `test` is `FALSE`) or `NA` (when `test` is `NA`). The resulting object gets coerced according to usual coercion rules and attributes (including the class) get stripped.
   
   Therefore, the expressions above return vectors of type double, character and logical.

2. __<span style="color:red">Q</span>__: Why does the following code work?

    ```{r}
    x <- 1:10
    if (length(x)) "not empty" else "empty"
      
    x <- numeric()
    if (length(x)) "not empty" else "empty"
    ```

   __<span style="color:green">A</span>__: `if()` expects a logical condition. Objects of other types are coercecd to logical if possible. Therefore, `0` gets coerced to `FALSE` and other numeric input gets coerced to `TRUE`. Numeric input which is not logically interpretable in the sense of `if()`, like `NaN`, `NA_integer` and `NA_real`, leads to an error.

## Loops

1. __<span style="color:red">Q</span>__: Why does this code succeed without errors or warnings? 
    
    ```{r, results = FALSE}
    x <- numeric()
    out <- vector("list", length(x))
    for (i in 1:length(x)) {
      out[i] <- x[i] ^ 2
    }
    out
    ```

   __<span style="color:green">A</span>__: The elements we are iterating over are of length 0. Our loop goes from `i = 1` to `i = 0`. As we use `[<-` and `[` for indexing, we need to be aware of its subsetting behaviour for out of bounds and zero indices. During the first iteration `x[1]` will generate an `NA` (out of bound indexing for atomics). The resulting `NA` from squaring will be coerced to a list and assigned to the empty length-1 list `out[1]` (out of bounds indexing for lists). Further `x[0]` will return `numeric(0)`. Again squaring doesn't change the value and `numeric(0)` gets coerced to `list()` when assigned to `out[0]`. Therefore, `out` doesn't change during the last iteration. This simply works, because it is valid R code.

2. __<span style="color:red">Q</span>__: What does the following code tell you about when the vector being 
    iterated over is evaluated?

    ```{r}
    xs <- c(1, 2, 3)
    for (x in xs) {
      xs <- c(xs, x * 2)
    }
    xs
    ```

   __<span style="color:green">A</span>__: `x` takes on the values of the initial `xs` (`1`, `2` and `3`). So, `xs` must be evaluated just once and before `x`.

3. __<span style="color:red">Q</span>__: What does the following code tell you about when the index is updated?

    ```{r}
    for (i in 1:3) {
      i <- i * 2
      print(i) 
    }
    ```

   __<span style="color:green">A</span>__: The index must get updated before each iteration. Therefore, reassigning the index symbol during the loop doesn't affect the index.
