# Control flow

## Choices

1. __<span style="color:red">Q</span>__: What type of vector does each of the following calls to `ifelse()`
    return?

    ```{r, eval = FALSE}
    ifelse(TRUE, 1, "no")
    ifelse(FALSE, 1, "no")
    ifelse(NA, 1, "no")
    ```

    Read the documentation and write down the rules in your own words.

   __<span style="color:green">A</span>__:

2. __<span style="color:red">Q</span>__: Why does the following code work?

    ```{r}
    x <- 1:10
    if (length(x)) "not empty" else "empty"
      
    x <- numeric()
    if (length(x)) "not empty" else "empty"
    ```

   __<span style="color:green">A</span>__:

## Loops

1. __<span style="color:red">Q</span>__: Why does this code succeed without errors or warnings? 
    
    ```{r, results = FALSE}
    x <- numeric()
    out <- vector("list", length(x))
    for (i in 1:length(x)) {
      out[i] <- x[i] ^ 2
    }
    out
    ```

   __<span style="color:green">A</span>__:

2. __<span style="color:red">Q</span>__: What does the following code tell you about when the vector being 
    iterated over is evaluated?

    ```{r}
    xs <- c(1, 2, 3)
    for (x in xs) {
      xs <- c(xs, x * 2)
    }
    xs
    ```

   __<span style="color:green">A</span>__:

3. __<span style="color:red">Q</span>__: What does the following code tell you about whe the index is updated?

    ```{r}
    for (i in 1:3) {
      i <- i * 2
      print(i) 
    }
    ```

   __<span style="color:green">A</span>__:
   
   
