# Control flow

## Choices

1. __<span style="color:red">Q</span>__: What type of vector does each of the following calls to `ifelse()` return?

    ```{r, eval = FALSE}
    ifelse(TRUE, 1, "no")
    ifelse(FALSE, 1, "no")
    ifelse(NA, 1, "no")
    ```

   Read the documentation and write down the rules in your own words.

   __<span style="color:green">A</span>__: The arguments of `ifelse()` are `test`, `yes` and `no`. `ifelse()` operates elementwise, so when `yes` or `no` are shorter than `test`, they will be recycled. (When they are longer than `test`, their additional elements will be ignored.)
   
   The function returns the entry for `yes` (when `test` is `TRUE`), `no` (when `test` is `FALSE`) or `NA` (when `test` is `NA`). The final object is coerced according to usual coercion rules with attributes (including the class) being stripped. Therefore, the expressions above return vectors of type `double`, `character` and `logical`.

2. __<span style="color:red">Q</span>__: Why does the following code work?

    ```{r}
    x <- 1:10
    if (length(x)) "not empty" else "empty"
      
    x <- numeric()
    if (length(x)) "not empty" else "empty"
    ```

   __<span style="color:green">A</span>__: `if()` expects a logical condition. Objects of other types are coerced to logical if possible. `0` is coerced to `FALSE` while other numeric input is coerced to `TRUE`. Numeric input which is not logically interpretable by `if()`, like `NaN`, `NA_integer` and `NA_real`, will lead to an error.

## Loops

1. __<span style="color:red">Q</span>__: Why does this code succeed without errors or warnings? 
    
    ```{r, results = FALSE}
    x <- numeric()
    out <- vector("list", length(x))
    for (i in 1:length(x)) {
      out[i] <- x[i] ^ 2
    }
    out
    ```

   __<span style="color:green">A</span>__: The vector we are iterating over are of length 0. Our loop goes from `i = 1` to `i = 0`. As we use `[<-` and `[` for indexing, we need to be aware of their subsetting behaviour for out-of-bounds and zero indices.
   
   During the first iteration `x[1]` will generate an `NA` (out-of-bounds indexing for atomics). The resulting `NA` (from squaring) will be assigned to the empty length-1 list `out[1]` (out-of-bounds indexing for lists).
   
   In the next iteration, `x[0]` will return `numeric(0)` (zero indexing for atomics). Again squaring doesn't change the value and `numeric(0)` is assigned to `out[0]` (zero indexing for lists). Please be aware, that `out` doesn't change during the last iteration.
   
   Overall the code works, because each step includes valid R operations. (Though it may have been very helpful to warn the user about the unusual inputs to this loop.)

2. __<span style="color:red">Q</span>__: What does the following code tell you about when the vector being iterated over is evaluated?

    ```{r}
    xs <- c(1, 2, 3)
    for (x in xs) {
      xs <- c(xs, x * 2)
    }
    xs
    ```

   __<span style="color:green">A</span>__: In this loop `x` takes on the values of the initial `xs` (`1`, `2` and `3`), which evaluated just once in the beginning of the loop.

3. __<span style="color:red">Q</span>__: What does the following code tell you about when the index is updated?

    ```{r}
    for (i in 1:3) {
      i <- i * 2
      print(i) 
    }
    ```

   __<span style="color:green">A</span>__: In a `for`-loop the index is updated in the beginning of each iteration. Therefore, reassigning the index symbol during one iteration doesn't affect the following iterations.
