# Functions

## Function fundamentals

1. __<span style="color:red">Q</span>__: Given a name, like `"mean"`, `match.fun()` lets you find a function. Given a function, can you find its name? Why doesn't that make sense in R?  

   __<span style="color:orange">A</span>__: If you know `body()`, `formals()` and `environment()` it is generally possible to find the corresponding function. (The search doesn't work for primitive functions, because they return `NULL` for these three properties and they are not bound to a name.) The search may also be ambiguous, because different names in an environment may contain bindings to more than one function with the same `body()`, `formals()` and `environment()` properties.

   Keep in mind: A (function) name has to be bound to an object, but an object (i.e. a function) doesn't always have a name. <!-- @hb: look this up again -->
   Also: You "can only find the parents of an environment, not its children, so this is necessarily only going to be a partial search" (Hadley Wickham). This means, that some function names in child environments will not be found. <!-- check this. -->

2. __<span style="color:red">Q</span>__: It’s possible (although typically not useful) to call an anonymous function. Which of the two approaches below is correct? Why?
   
    ```{r}
    function(x) 3()
    (function(x) 3)()
    ```

   __<span style="color:green">A</span>__: The second approach is correct.
   
   The anonymous function `function(x) 3` is surrounded by a pair of parentheses before it is called by `()`. These extra parentheses separate the function call from the anonymous functions body. Without these a function with the invalid body `3()` is returned, which throws an error when we call it:
   
    ```{r, error = TRUE}
    (function(x) 3())()
    ```


3. __<span style="color:red">Q</span>__: A good rule of thumb is that an anonymous function should fit on one line and shouldn't need to use `{}`. Review your code. Where could you have used an anonymous function instead of a named function? Where should you have used a named function instead of an anonymous function?  
    
   __<span style="color:green">A</span>__: The use of anonymous functions allows concise and elegant code in certain situations. However, they miss a descriptive name and when re-reading the code it can take a while to figure out what they do (even it it's future you reading). That's why it's helpful to give long and complex functions a descriptive name. It may be worthwhile to take a look at your own projects or other peoples code to reflect on this part of your coding style.

<!-- @hb: maybe describe advantages of anonymous functions more clearly. -->
<!-- comment: maybe write about use of anonymous functions
for functional programming (lapply, purrr) -->

4. __<span style="color:red">Q</span>__: What function allows you to tell if an object is a function? What function allows you to tell if a function is a primitive function?
       
   __<span style="color:green">A</span>__: Use `is.function()` to test, if an object is a function. You may also consider `is.primitive()` to test specifically for primitive functions.

5. __<span style="color:red">Q</span>__: This code makes a list of all functions in the base package. 
    
    ```{r}
    objs <- mget(ls("package:base", all = TRUE), inherits = TRUE)
    funs <- Filter(is.function, objs)
    ```
    
   Use it to answer the following questions:
    
   a. Which base function has the most arguments?
    
   a. How many base functions have no arguments? What's special about those functions?
   
   a. How could you adapt the code to find all primitive functions?  
    
   __<span style="color:green">A</span>__: Let's look at each sub-question separately:
   
   a. To find the function with the most arguments, access the arguments of each function in the list. We count each functions arguments and subset to the maximum value. This returns the name of the function with the most arguments and also the corresponding number of arguments:
   
      ```{r}
      library(magrittr)
      
      funs %>% 
        sapply(formals) %>%
        lengths() %>% 
        `[`(which.max(.))
      ```
      
   b. To find the number functions with no argument, we compute the number of arguments for each function and check, if they are equal to `0`. The sum of this result gives us the answer we want.
   
      ```{r}
      funs %>%
        sapply(formals) %>% 
        lengths() %>%
        `==`(0L) %>%
        sum()
      ```
        Because `formals()` returns `NULL` for primitive functions, these are included in our count of functions with no argument. We could also divide the functions into primitive and non-primitive functions. We filter the function list for functions with zero arguments first. We then test if these functions are primitive and tabulate the result:
     
      ```{r}
      funs %>%
        Filter(function(x) length(formals(x)) == 0, .) %>% 
        sapply(is.primitive) %>% 
        table()
      ```
        Indeed, most of the functions with no argrument in base R are primitive.
     
   c. To find all primitive functions, we can change the predicate in `Filter()` from `is.function()` to `is.primitive()`:
   
      ```{r, eval = FALSE}
      funs <- Filter(is.primitive, objs)
      ```
      
6. __<span style="color:red">Q</span>__: What are the three important components of a function?  
    
   __<span style="color:green">A</span>__: These components are the function's `body()`, `formals()` and `environment()`. However, as mentioned in the textbook:
    
   > There is one exception to the rule that functions have three components. Primitive functions, like `sum()`, call C code directly with `.Primitive()` and contain no R code. Therefore their `formals()`, `body()`, and `environment()` are all `NULL`.

7. __<span style="color:red">Q</span>__: When does printing a function not show what environment it was created in?  

   __<span style="color:green">A</span>__: Only primitive functions or functions created in the global environment do not show the environment they were created in when they are printed.

## Lexical Scoping

1. __<span style="color:red">Q</span>__: What does the following code return? Why? Describe how each of the three `c`’s is interpreted.

    ```{r, eval = FALSE}
    c <- 10
    c(c = c)
    ```  
    
   __<span style="color:green">A</span>__: This code returns a named numeric vector of length one - with one element of the value `10` and the name `"c"`. The first `c` represents the `c()` function, the second `c` is interpreted as a (quoted) name and the third `c` as a value.
       
2. __<span style="color:red">Q</span>__: What are the four principles that govern how R looks for values?  
    
   __<span style="color:green">A</span>__: R's lexical scoping rules are based on these four principles:
   
   * Name masking
   * Functions vs. variables
   * A fresh start
   * Dynamic lookup

3. __<span style="color:red">Q</span>__: What does the following function return? Make a prediction before running the code yourself.

    ```{r, eval = FALSE}
    f <- function(x) {
      f <- function(x) {
        f <- function(x) {
          x ^ 2
        }
        f(x) + 1
      }
      f(x) * 2
    }
    f(10)
    ```
        
   __<span style="color:green">A</span>__: Within this function two more functions also named `f()` are defined and called. Because the functions are each executed in their own environment R will look up and use the functions defined in these environments. The innermost `f()` is called last, though it is the first function to return a value. Because of this the order of the calculation passes "from the inside to the outside" and the function returns `202`.


## Lazy evaluation

1. __<span style="color:red">Q</span>__: What important property of `&&` makes `x_ok()` work?

    ```{r}
    x_ok <- function(x) {
      !is.null(x) && length(x) == 1 && x > 0
    }
    
    x_ok(NULL)
    x_ok(1)
    x_ok(1:3)
    ```

   What is different with this code? Why is this behaviour undesirable here?
       
    ```{r}
    x_ok <- function(x) {
      !is.null(x) & length(x) == 1 & x > 0
    }
    
    x_ok(NULL)
    x_ok(1)
    x_ok(1:3)
    ```
    
   __<span style="color:green">A</span>__: We expect `x_ok()` to validate its input vie certain criteria: it must not be `NULL`, but have length `1` and a value greater than `0`. Meaningful outcomes for this assertion will be `TRUE`, `FALSE` or `NA`.
   
   The desired behaviour is reached by combining the assertions through `&&` instead of `&`. `&&` does not perform elementwise comparisons, instead it uses the first element of each value only. It also uses lazy evaluation, in the sense that evaluation "proceeds only until the result is determined" (from `?Logic`).
   
   For some situations (`x = 1`) both operators will lead to the same result. But this is not always the case. For `x = NULL`, the `&&`-operator will stop after the `!is.null`-statement and return the result. The following conditions won't even be evaluated! (If the other conditions are also evaluated (by the use of `&`), the outcome would change. `NULL > 0` returns `logical(0)`, which is not a helpful in this case.)
   
   We can also see the difference in behaviour, when we set `x = 1:3`. The `&&`-operator returns the result from `length(x) == 1`, which is `FALSE`. Using `&` as the logical operator leads to the (vectorised) `x > 0` condition being be evaluated and also returned.

2. __<span style="color:red">Q</span>__: What does this function return? Why? Which principle does it illustrate?

    ```{r, eval = FALSE}
    f2 <- function(x = z) {
      z <- 100
      x
    }
    f2()
    ```  
    
   __<span style="color:green">A</span>__: The function returns `100`. The default arguments are evaluated in the function environment. Because of *lazy evaluation* these arguments are not evaluated before they are accessed. At the time `x` is accessed `z` has already been bound to the value `100`.
    
3. __<span style="color:red">Q</span>__: What does this function return? Why? Which principle does it illustrate?
  
    ```{r, eval = FALSE}
    y <- 10
    f1 <- function(x = {y <- 1; 2}, y = 0) {
      c(x, y)
    }
    f1()
    y
    ```  
    
   __<span style="color:green">A</span>__: The function returns `c(2, 1)`. This is due to *name masking*. When `x` is accessed within `c()`, the promise `x = {y <- 1; 2}` is evaluated inside `f1()`'s environment. `y` is bound to the value `1` and the return value of `{()` (`2`) is assigned to `x`. When `y` is accessed within `c()`, it has already the value `1` and R doesn't need to look it up any further. Therefore, the promise `y = 0` won't be evaluated. Also, because `y` is assigned within `f1()`'s environment, the value of the global variable `y` is left untouched.

4. __<span style="color:red">Q</span>__: In `hist()`, the default value of `xlim` is `range(breaks)`, the default value for `breaks` is `"Sturges"`, and

    ```{r}
    range("Sturges")
    ```
    
   Explain how `hist()` works to get a correct `xlim` value.
    
   __<span style="color:green">A</span>__: The `xlim` argument of `hist()` defines the range of the histogram's x-axis. In order to provide a valid axis `xlim` must contain a numeric vector of exactly two unique values. Consequently for the default `xlim = range(breaks)`), `breaks` must evaluate to a vector with at least two unique values.
   
   During execution `hist()` overwrites the `breaks` argument. The `breaks` argument is quite flexible and allows the users to provide the breakpoints directly or compute them in several ways. Therefore the specific behaviour depends highly on the input. But `hist` ensures that `breaks` evaluates to a numeric vector containing at least two unique elements before `xlim` is accessed.
   
5. __<span style="color:red">Q</span>__: Explain why this function works. Why is it confusing?

    ```{r}
    show_time <- function(x = stop("Error!")) {
      stop <- function(...) Sys.time()
      print(x)
    }
    show_time()
    ```
    
   __<span style="color:green">A</span>__: Before `show_time()` accesses `x` (default `stop("Error")`), the `stop()` function is overwritten as `function(...) Sys.time()`. Because default arguments are evaluated in the function environment, `print(x)` will be evaluated as `print(Sys.time())`.
   
   This function is confusing, because its behaviour changes when `x`'s value is supplied directly. Now the value from the calling environment will be used and the overwriting of `stop` won't affect the outcome any more.
  
    ```{r, error = TRUE}
    show_time(x = stop("Error!"))
    ```

6. __<span style="color:red">Q</span>__: How many arguments are required when calling `library()`?

   __<span style="color:green">A</span>__: `library()` doesn't require any arguments. When called without arguments `library()` (invisibly) returns a list of class "libraryIQR", which contains a results matrix with one row and three columns per installed package. These columns contain entries for the name of the package ("Package"), the path to the package ("LibPath") and the title of the package ("Title"). `library()` also has its own print method (`print.libraryIQR`), which displays this information conveniently in its own window.
   
   This behaviour is also documented under the details section of the help page for `?library`:
   
   > If library is called with no package or help argument, it lists all available packages in the libraries specified by lib.loc, and returns the corresponding information in an object of class “libraryIQR”. (The structure of this class may change in future versions.) Use .packages(all = TRUE) to obtain just the names of all available packages, and installed.packages() for even more information.
   
   Because the `package` and `help` argument from `library()` do not show a default value, it's easy to overlook the possibility to call `library()` without these arguments. (Instead of providing `NULL`s as default values `library()` uses `missing()` to check if these arguments were provided.)

    ```{r}
    str(formals(library))
    ```

## `...` (dot-dot-dot)

1. __<span style="color:red">Q</span>__: Explain the following results:
    
    ```{r}
    sum(1, 2, 3)
    mean(1, 2, 3)
    
    sum(1, 2, 3, na.omit = TRUE)
    mean(1, 2, 3, na.omit = TRUE)
    ```
    
   __<span style="color:green">A</span>__: Let's inspect the arguments and their order for both functions. For `sum()` these are `...` and `na.rm`:
   
    ```{r}
    str(sum)
    ```
    
   For the `...` argument `sum()` expects numeric, complex or logical vector input (see `?sum`). Unfortunately, when `...` is used, misspelled arguments (!) like `na.omit` won't raise an error (When no further input checks are implemented). So instead, `na.omit` is treated as a logical and becomes part of the `...` argument. It will be coerced to `1` and be part of the sum. All other arguments are left unchanged. Therefore `sum(1, 2, 3)` returns `6` and `sum(1, 2, 3, na.omit = TRUE)` returns `7`.
   
   In contrast, the generic function `mean()` expects `x`, `trim`, `na.rm` and `...` for its default method.
   
    ```{r}
    str(mean.default)
    ```
   
  Because `na.omit` is not one of `mean()`'s named arguments (and also not a candidate for partial matching), `na.omit` again becomes part of the `...` argument. The other supplied objects are matched by their order, i.e.: `x = 1`, `trim = 2` and `na.rm = 3`. Because `x` is of length 1 and not `NA`, the settings of `trim` and `na.rm` do not affect the calculation of the mean. Both calls (`mean(1, 2, 3)` and `mean(1, 2, 3, na.omit = TRUE)`) return `1`.

2. __<span style="color:red">Q</span>__: In the following call, explain how to find the documentation for the named arguments in the following function call:
       
    ```{r, fig.asp = 1}
    plot(1:10, col = "red", pch = 20, xlab = "x", col.lab = "blue")
    ```
    
   __<span style="color:green">A</span>__: First we type `?plot` in the console and check the "Usage" section:
    
    ```
    plot(x, y, ...)
    ```
    
   The arguments we want to learn more about are part of the `...` argument. We can find information for `xlab` and follow the recommendation to visit `?par` for the other arguments. Here we type "col" into the search bar, which leads us the section "Color Specification". We also search for the `pch` argument, which leads to the recommendation to check `?points`. Finally `col.lab` is also directly documented within `?par`.
    
3. __<span style="color:red">Q</span>__: Why does `plot(1:10, col = "red")` only colour the points, not the axes or labels? Read the source code of `plot.default()` to find out.
    
   __<span style="color:green">A</span>__: To learn about the internals of `plot.default()` we add `browser()` to the first line of the code and interactively run `plot(1:10, col = "red")`. This way we can see how the plot is build and learn where the axis are added.
   
   This leads us to the function call

    ```{r, eval = FALSE}
    localTitle(main = main, sub = sub, xlab = xlab, ylab = ylab, ...)
    ```
    
   The `localTitle()` function was defined in the first lines of `plot.default()` as:

    ```{r, eval = FALSE}
    localTitle <- function(..., col, bg, pch, cex, lty, lwd) title(...)
    ```
    
   The call to `localTitle()` will be passed the `col` parameter as part of `...` argument. `?title` tells us that the `title()` function specifies four parts of the plot: Main (title of the plot), sub (sub-title of the plot) and both axis labels. Because of this it would introduce ambiguity inside `title()` to use `col` directly. Instead one has the option to supply `col` via the `...` argument as `col.labs` or as part of `xlab` (similar for `ylab`) in the form `xlab = list(c("index"), col = "red")`.

## Exiting a function

1. __<span style="color:red">Q</span>__: What does `load()` return? Why don’t you normally see these values?

   __<span style="color:green">A</span>__: `load()` reloads datasets which were previously written to disk via `save()`. When run successfully, `load()` invisibly returns a character vector, containing the names of the newly loaded objects. To print these names to the console, one can set the argument `verbose` to `TRUE` or surround the call in parentheses to trigger R's auto-printing mechanism.
   
2. __<span style="color:red">Q</span>__: What does `write.table()` return? What would be more useful?

   __<span style="color:green">A</span>__: `write.table()` writes an object, usually a data frame or a matrix, to disk. The function invisibly returns `NULL`. It would be more useful, if `write.table()` would (invisibly) return its `x` object. This would allow to save intermediate results and directly take on further processing steps without breaking the flow of the code (i.e. breaking it into different lines). One package which uses this pattern is the readr package, which is part of the "tidyverse"-ecosystem.
    
3. __<span style="color:red">Q</span>__: How does the `chdir` parameter of `source()` compare to `in_dir()`? Why might you prefer one approach to the other?

   The `in_dir()` approach was given in the book as
       
    ```{r, eval = FALSE}
    in_dir <- function(dir, code) {
      old <- setwd(dir)
      on.exit(setwd(old))
      
      force(code)
    }
    ```
    
   __<span style="color:orange">A</span>__: `in_dir()` takes a path to a working directory as an argument. First the working directory is changed accordingly. `on.exit()` ensures, that the modification to the working directory are reset to the initial value, when the function exits.
    
   In `source()` the `chdir` argument specifies, if the working directory should be changed during the evaluation of the `file` argument (which in this case has to be a pathname). But in `source()` the current working directory is stored and passed to `on.exit()`, before the working directory is changed for the subsequent evaluation.
   
<!-- Why might you prefer one approach to the other? Explain what happens in `source` more clearly. -->

4. __<span style="color:red">Q</span>__: Write a function that opens a graphics device, runs the supplied code, and closes the graphics device (always, regardless of whether or not the plotting code worked).  
   
   __<span style="color:green">A</span>__: To control the graphics device we use `pdf()` and `dev.off()`. To ensure a clean termination `on.exit()` is used.
    
    ```{r, eval = FALSE}
    plot_pdf <- function(code){
      pdf("test.pdf")
      on.exit(dev.off())
      code
      }
    ```

5. __<span style="color:red">Q</span>__: We can use `on.exit()` to implement a simple version of `capture.output()`.

    ```{r}
    capture.output2 <- function(code) {
      temp <- tempfile()
      on.exit(file.remove(temp), add = TRUE)

      sink(temp)
      on.exit(sink(), add = TRUE)

      force(code)
      readLines(temp)
    }
    capture.output2(cat("a", "b", "c", sep = "\n"))
    ```
    
   Compare `capture.output()` to `capture.output2()`. How do the functions differ? What features have I removed to make the key ideas easier to see? How have I rewritten the key ideas to be easier to understand?
    
   __<span style="color:green">A</span>__: Using `body(capture.output)` we inspect the source code of the original `capture.output()` function. `capture.output()` is a quite a bit longer (39 lines vs. 7 lines). `capture.output()` writes out entire methods, such as `readLines()`. Instead `capture.output2()` calls these methods directly. This brevity and modularity makes `capture.output2` easier to understand (given you know the underlying methods).

   However `capture.output2()` does miss a couple of features: `capture.output()` appears to handle important exceptions and it also offers a choice between overwriting or appending to a file.

## Function forms

1. __<span style="color:red">Q</span>__: Rewrite the following code snippets into prefix form:

    ```{r, eval = FALSE}
    1 + 2 + 3
    
    1 + (2 + 3)
    
    if (length(x) <= 5) x[[5]] else x[[n]]
    ```
    
   __<span style="color:green">A</span>__: Let's rewrite the expressions to match the exact syntax from the code above. Because prefix functions already define the execution order, we may omit the parentheses in the second expression.
    
    ```{r, eval = FALSE}
    `+`(`+`(1, 2), 3)
    
    `+`(1, `(`(`+`(2, 3)))  # `+`(1, `+`(2, 3))
    
    `if`(`<=`(length(x), 5), `[[`(x, 5), `[[`(x, n))
    ```

2. __<span style="color:red">Q</span>__: Clarify the following list of odd function calls:

    ```{r, eval = FALSE}
    x <- sample(replace = TRUE, 20, x = c(1:10, NA))
    y <- runif(min = 0, max = 1, 20)
    cor(m = "k", y = y, u = "p", x = x)
    ```  
    
   __<span style="color:green">A</span>__: None of these functions provides a `...` argument. Therefore the function arguments are first matched exactly, then via partial matching and finally by position. This leads us to the following explicit function calls:
   
    ```{r, eval = FALSE}
    x <- sample(x = c(1:10, NA), size = 20, replace = TRUE)
    y <- runif(n = 20, min = 0, max = 1)
    cor(x = x, y = y, use = "pairwise.complete.obs", method = "kendall")
    ```
    
3. __<span style="color:red">Q</span>__: Explain why the following code fails:

    ```{r, eval = FALSE}
    modify(get("x"), 1) <- 10
    #> Error: target of assignment expands to non-language object
    ```
    
   __<span style="color:orange">A</span>__: First, let's define `x` and recall the definition of `modify()` from the textbook:
    
    ```{r}
    x <- 1:3
    
    `modify<-` <- function(x, position, value) {
      x[position] <- value
      x
    }
    ```
    
   R internally transforms the code and the transformed code reproduces the error above.
    
    ```{r, eval = FALSE}
    get("x") <- `modify<-`(get("x"), 1, 10)
    #> Error in get("x") <- `modify<-`(get("x"), 1, 10) : 
    #>   target of assignment expands to non-language object
    ```
    
   The error occurs during the assignment, because no corresponding replacement function, i.e. `get<-` exists for `get()`. To confirm this we can reproduce the error via the following simple example.
    
    ```{r, eval = FALSE}
    get("x") <- 2
    #> Error in get("x") <- 2 : target of assignment expands to non-language object
    ```
    
4. __<span style="color:red">Q</span>__: Create a replacement function that modifies a random location in a vector.
    
   __<span style="color:green">A</span>__: Lets define `%random%` like this:

    ```{r, eval = FALSE}
    `random<-` <- function(x, value){
      x[sample(length(x), 1)] <- value
      x
      }
    ```

5. __<span style="color:red">Q</span>__: Write your own version of `+` that will paste its inputs together if they are character vectors but behaves as usual otherwise. In other words, make this code work:
   
    ```{r, eval = FALSE}
    1 + 2
    #> [1] 3
    
    "a" + "b"
    #> [1] "ab"
    ```

   __<span style="color:green">A</span>__: To achieve this behaviour, we need to override the `+` operator. We need to take care to not use the `+` operator itself inside of the function definition, because this would lead to an undesired infinite recursion. We also add `b = 0L` as a default value to keep the behaviour of `+` as a unary operator, i.e. to keep `+ 1` working and not throwing an error
    
    ```{r}
    `+` <- function(a, b = 0L){
      if (is.character(a) && is.character(b)) {return(paste0(a, b))}
      sum(a, b)
    }
    
    # test functionality
    + 1
    1 + 2
    "a" + "b"
    
    # return back to the original `+` operator
    rm(`+`)
    ```

6. __<span style="color:red">Q</span>__: Create a list of all the replacement functions found in the base package. Which ones are primitive functions? (Hint use `apropos()`)
   
   __<span style="color:green">A</span>__: The hint suggests to look for functions with a specific naming pattern: Replacement functions conventionally end on `<-`. We can search these objects with a regular expression (`<-$`). 
   
   However, instead of `apropros()`we will use `ls()` and adopt a bit of the code from a previous exercise. (This makes it easier to work with environments explicitly.) From all the objects in the base package which end on `<-`, we subset the functions. Additionally, we also filter for primitive functions.
   
    ```{r}
    repl_objects <- mget(ls("package:base", all.names = TRUE, pattern = "<-$"),
                      inherits = TRUE)
    repl_functions <- Filter(is.function, repl_objects)
    length(repl_functions)
    ```
   
   Overall base R contains `r length(repl_functions)` replacement functions. The following `r length(Filter(is.primitive, repl_functions))` of them are also primitive functions:
   
    ```{r}
    names(Filter(is.primitive, repl_functions))
    ```

7. __<span style="color:red">Q</span>__: What are valid names for user-created infix functions?  
  
   __<span style="color:green">A</span>__: Let's cite **Advanced R** here (section on "Function Forms"):
   
   > ... names of infix functions are more flexible than regular R functions: they can contain any sequence of characters except “%”.

8. __<span style="color:red">Q</span>__: Create an infix `xor()` operator.
    
   __<span style="color:green">A</span>__: We could create an infix `%xor_%` like this:

    ```{r}
    `%xor_%` <- function(a, b){
      (a | b) & !(a & b)
    }
    ```

9. __<span style="color:red">Q</span>__: Create infix versions of the set functions `intersect()`, `union()`, and`setdiff()`. You might call them `%n%`, `%u%`, and `%/%` to match conventions from mathematics.
   
   __<span style="color:green">A</span>__: These infix operators could be defined in the following way. (`%/%` is chosen instead of `%\%`, because `\` serves as an escape character.)
   
    ```{r}
    `%n%` <- function(a, b){
      unique(a[a %in% b])
    }

    `%u%` <- function(a, b){
      unique(c(a, b))
    }

    `%/%` <- function(a, b){
      a[!a %in% b]
    }
    ```
