# Conditions

## Signalling conditions

1. __<span style="color:red">Q</span>__: Write a wrapper around `file.remove()` that throws an error if the file to be deleted does not exist.

   __<span style="color:green">A</span>__: We have several options here. However, we prefer the following solution for its clarity and simplicity:
    
    ```{r, error = TRUE}
    file_remove_strict <- function(...) {
      if(!file.exists(...)) {
        stop("Can't delete ", `...`, ", because the file doesn't exist.", call. = FALSE)
      }
      file.remove(...)
    }
    
    # test
    saveRDS(iris, "iris.RDS")
    file_remove_strict("iris.RDS")
    file_remove_strict("iris.RDS")
    ```

2. __<span style="color:red">Q</span>__: What does the `appendLF` argument to `message()` do? How is it related to `cat()`?

   __<span style="color:green">A</span>__: The `appendLF` argument allows to append a new line to the message. Let's illustrate this behaviour with a small example function:
   
    ```{r, eval=FALSE}
    multiline_msg <- function(appendLF = TRUE) {
      message("first line", appendLF = appendLF)
      cat("second line")
    }
    
    multiline_msg(appendLF = TRUE)
    #> first line
    #> second line
    multiline_msg(appendLF = FALSE)
    #> first linesecond line
    ```

   This is very similar to the `fill` argument in `cat()`:
    
    ```{r, eval=FALSE}
    multiline_cat <- function(fill = TRUE) {
      cat("first line", fill = fill)
      message("second line")
    }
    
    multiline_cat(fill = TRUE) 
    #> first line
    #> second line
    multiline_cat(fill = FALSE)
    #> first linesecond line
    ```
    
   Alternatively, you can also create a new line by adding `\n` to the end of the message string.

## Handling conditions

1. __<span style="color:red">Q</span>__: What extra information does the condition generated by `abort()` contain compared to the condition generated by `stop()`? i.e. what's the difference between these two objects? Read the help for `?abort` to learn more.

    ```{r, include=FALSE}
    library(rlang)
    ```

    ```{r, eval = FALSE}
    catch_cnd(stop("An error"))
    catch_cnd(abort("An error"))
    ```
    
   __<span style="color:green">A</span>__: In contrast to `stop()`, which contains the call, `abort()` stores the whole backtrace generated by `rlang::trace_back()`: 
   
    ```{r, eval = FALSE}
    str(catch_cnd(stop("An error")))
    #> List of 2
    #> $ message: chr "An error"
    #> $ call   : language force(expr)
    #> - attr(*, "class")= chr [1:3] "simpleError" "error" "condition"
    
    str(catch_cnd(abort("An error")))
    #> List of 3
    #> $ message: chr "An error"
    #> $ trace  :List of 3
    #>  ..$ calls  :List of 7
    #>  .. ..$ : language str(catch_cnd(abort("An error")))
    #>  .. ..$ : language catch_cnd(abort("An error"))
    #>  .. ..$ : language tryCatch(condition = identity, {     force(expr) ...
    #>  .. ..$ : language tryCatchList(expr, classes, parentenv, handlers)
    #>  .. ..$ : language tryCatchOne(expr, names, parentenv, handlers[[1L]])
    #>  .. ..$ : language doTryCatch(return(expr), name, parentenv, handler)
    #>  .. ..$ : language force(expr)
    #>  ..$ parents: int [1:7] 0 0 2 3 4 5 2
    #>  ..$ envs   :List of 7  # (abbreviated)
    #>  ..- attr(*, "class")= chr "rlang_trace"
    #> $ parent : NULL
    #> - attr(*, "class")= chr [1:3] "rlang_error" "error" "condition"
    ```

2. __<span style="color:red">Q</span>__: Predict the results of evaluating the following code

    ```{r, include = FALSE}
    show_condition <- function(code) {
      tryCatch(
        error = function(cnd) "error",
        warning = function(cnd) "warning",
        message = function(cnd) "message",
        {
          code
          NULL
        }
      )
    }
    ```
    
    ```{r}
    show_condition <- function(code) {
      tryCatch(
        error = function(cnd) "error",
        warning = function(cnd) "warning",
        message = function(cnd) "message",
        {
          code
          NULL
        }
      )
    }
    
    show_condition(stop("!"))
    show_condition(10)
    show_condition(warning("?!"))
    show_condition({
      10
      message("?")
      warning("?!")
    })
    ```

__<span style="color:green">A</span>__: The first three examples are straightforward. The last example is the most interesting and makes us aware of the exiting qualities of `tryCatch()`, it will terminate the evaluation of the code as soon as it is called.
   
    ```{r}
    show_condition(stop("!"))      # stop raises an error
    show_condition(10)             # no condition is signalled
    show_condition(warning("?!"))  # warning raises a warning
    show_condition({               # tryCatch() is an exiting handler
      10
      message("?")
      warning("?!")
    })
    ```

3. __<span style="color:red">Q</span>__: Explain the results of running this code:

    ```{r}
    withCallingHandlers(
      message = function(cnd) message("b"),
      withCallingHandlers(
        message = function(cnd) message("a"),
        message("c")
      )
    )
    ```
    
   __<span style="color:green">A</span>__: Lets look at the inner `withCallingHandlers()` first:
   The first message printed by this statement would be `a`, so the outer `withCallingHandlers` prints `b` (1). 
   <!-- so what exactly is happening here? -->
   Afterwards the message of the inner handler gets printed: `a` (2).
   Next the inner `withCallingHandlers()` would print `c`, so the outer `withCallingHandlers()` prints again `b` (3).
   Finally `c` (4) is printed.

4. __<span style="color:red">Q</span>__: Read the source code for `catch_cnd()` and explain how it works.

   __<span style="color:green">A</span>__: `catch_cnd` basically wraps `tryCatch` and returns the default result. The expression is evaluated and it is ensured that `NULL` is returned, when there is no signalling condition (instead of the expression's return value). `catch_cnd` returns a list with the the first condition caught (containing message and call) or `NULL`.
    
    ```{r}
    rlang::catch_cnd
    ```

5. __<span style="color:red">Q</span>__: How could you rewrite `show_condition()` to use a single handler?

   __<span style="color:green">A</span>__: Let's use the `condition` argument of `tryCatch` as shown in `rlang::catch_cond()` above:
    
    ```{r}
    show_condition2 <- function(code) {
      tryCatch(
        condition = function(cnd) {
          if (inherits(cnd, "error"))   {return("error")}
          if (inherits(cnd, "warning")) {return("warning")}
          if (inherits(cnd, "message")) {return("message")}
        },
        {
          code
          NULL
        }
      )
    }
    
    # test
    show_condition2(stop("!"))
    show_condition2(10)
    show_condition2(warning("?!"))
    show_condition2({
      10
      message("?")
      warning("?!")
    })
    ```

<!-- ideally explain a bit more what is happening here -->

## Custom conditions

1. __<span style="color:red">Q</span>__: Inside a package, it’s occasionally useful to check that a package is installed before using it. Write a function that checks if a package is installed (with `requireNamespace("pkg", quietly = FALSE))` and if not, throws a custom condition that includes the package name in the metadata.

   __<span style="color:green">A</span>__: It's quite tedious to build up this behaviour on top of `requireNamespace()`. However, if all arguments are supplied correctly, `requiredNamespace()`'s return behaviour can be summarised in the following way:
   
   - when the namespace has already been loaded, `TRUE` is returned
   - when the namespace is not yet loaded: 
     - a message is thrown
     - depending on further success or failure `TRUE` or `FALSE` is returned.

   Therefore, we change the behaviour of `requireNamespace` for missing packages and we return the error condition:
   
    ```{r}
    library(rlang)

    requireNamespace2 <- function(package)  {
      # First let us include a basic check for the input:
      if (!is.character(package) | !length(package) == 1L) {
        abort("error_bad_argument",
              message = "`package` must be a length-1 character vector.")
      }
      
      withCallingHandlers(        
        requireNamespace(package, quietly = FALSE),
        # Because the original error messages are hard to differentiate in a robust way,
        # we catch all messages first:
        message = function(cnd)  {
          # In case of failure, we throw the error. 
          # Otherwise the original message will be returned
          if (!requireNamespace(package, quietly = TRUE)) {
            abort(
              "error_pkg_not_found",
              message = paste0("`package` ", package, " not found."),
              package = package
            )}
        }
      )
  
      invisible(TRUE)
    }
    ```
      
   Now we can test our new function for all cases:
      
    ```{r, message = TRUE, error = TRUE}
    (requireNamespace2("base"))    # loaded already
    (requireNamespace2("lobstr"))  # initially loads package
    (requireNamespace2("lobstr")) 
    catch_cnd(requireNamespace2("pkg"))  # throws an error
    catch_cnd(requireNamespace2("pkg"))$package  # returns package-name
    ```

2. __<span style="color:red">Q</span>__: Inside a package you often need to stop with an error when something is not right. Other packages that depend on your package might be tempted to check these errors in their unit tests. How could you help these packages to avoid relying on the error message which is part of the user interface rather than the API and might change without notice?

   __<span style="color:green">A</span>__: Instead returning an error it might be preferable to throw a customized condition and place a standardized error message inside the metadata. This message would remain stable and also won't be affected by automatic translations.

## Applications

1. __<span style="color:red">Q</span>__: Create `suppressConditions()` that works like `suppressMessages()` and `supressWarnings()` but suppresses everything. Think carefully about how you should handle errors.

   __<span style="color:green">A</span>__: In general we would like to catch errors, since they contain important information for debugging. In order to suppress the error message and hide the returned error object from the console, we handle errors within a `tryCatch()` and return the error object invisibly:

    ```{r}
    suppressErrors <- function(expr) {
      tryCatch(
        error = function(cnd) invisible(cnd),
        interrupt = function(cnd) stop("Terminated by the user", call. = FALSE),
        expr
      )
    }
    ```
    
   After we defined the error handling, we can just combine it with the other handlers to create `suppressConditions()`:

    ```{r}
    suppressConditions <- function(expr) {
      suppressErrors(suppressWarnings(suppressMessages(expr)))
    }
    ```
    
   To test the new function we apply it to a set of conditions and inspect the returned error object.
    
    ```{r}
    error_obj <- suppressConditions({
        message("message")
        warning("warning")
        abort("error")
      })  # the messages/ wrarnings/ conditions are suppressed successfully
    
    error_obj
    ```

2. __<span style="color:red">Q</span>__: Compare the following two implementations of `message2error()`. What is the main advantage of `withCallingHandlers()` in this scenario? (Hint: look carefully at the traceback.)
   
    ```{r}
    message2error <- function(code) {
      withCallingHandlers(code, message = function(e) stop(e))
    }
    message2error <- function(code) {
      tryCatch(code, message = function(e) stop(e))
    }
    ```
    
   __<span style="color:green">A</span>__: Calling handlers are called in the context of the call that signalled the condition. Exiting handlers are called in the context of the call to `tryCatch()`.
   
    ```{r, error = TRUE, eval=FALSE}
    message2error1 <- function(code) {
      withCallingHandlers(code, message = function(e) stop("error"))
    }

    message2error1({1;  message("hidden error"); NULL})
    #> Error in (function (e) : error
    traceback()
    #> 9: stop("error") at #2
    #> 8: (function (e) 
    #>    stop("error"))(list(message = "hidden error\n", call = message("hidden error")))
    #> 7: signalCondition(cond)
    #> 6: doWithOneRestart(return(expr), restart)
    #> 5: withOneRestart(expr, restarts[[1L]])
    #> 4: withRestarts({
    #>        signalCondition(cond)
    #>        defaultHandler(cond)
    #>    }, muffleMessage = function() NULL)
    #> 3: message("hidden error") at #1
    #> 2: withCallingHandlers(code, message = function(e) stop("error")) at #2
    #> 1: message2error1({
    #>        1
    #>        message("hidden error")
    #>        NULL
    #>    })
    ```

   As seen above, the use of `withCallingHandlers()` returns more information and points us to the exact call in our code:

    ```{r, error = TRUE, eval=FALSE}
    message2error2 <- function(code) {
      tryCatch(code, message = function(e) (stop("error")))
    }
    message2error2({1; message("hidden error"); NULL})
    #> Error in value[[3L]](cond) : error
    traceback()
    #> 6: stop("error") at #2
    #> 5: value[[3L]](cond)
    #> 4: tryCatchOne(expr, names, parentenv, handlers[[1L]])
    #> 3: tryCatchList(expr, classes, parentenv, handlers)
    #> 2: tryCatch(code, message = function(e) (stop("error"))) at #2
    #> 1: message2error2({
    #>        1
    #>        message("hidden error")
    #>        NULL
    #>    })
    ```
    
3. __<span style="color:red">Q</span>__:  How would you modify the `catch_cnds()` defined if you wanted to recreate the original intermingling of warnings and messages?

   __<span style="color:green">A</span>__: The output of `catch_cnds()` should be returned in the order: messages, errors, warnings. To mimic the orignal behaviour we can sort the output accordingly:

    ```{r}
    catch_cnds <- function(expr) {
      conds <- list()
      add_cond <- function(cnd) {
        conds <<- append(conds, list(cnd))
        cnd_muffle(cnd)
      }
  
      tryCatch(
        error = function(cnd) {
          conds <<- append(conds, list(cnd))
        },
        withCallingHandlers(
          message = add_cond,
          warning = add_cond,
          expr
        )
      )
  
      filter_cndts <- function(conds, condition) {
        Filter(function(x) inherits(x, condition), conds)
      }
  
      c(filter_cndts(conds, "message"),
        filter_cndts(conds, "error"),
        filter_cndts(conds, "warning"))
    }
    
    # test
    catch_cnds({
      inform("message a")
      warn("warning b")
      inform("message c")
    })
    ```
   
4. __<span style="color:red">Q</span>__:  Why is catching interrupts dangerous? Run this code to find out.

    ```{r, eval = FALSE}
    bottles_of_beer <- function(i = 99) {
      message("There are ", i, " bottles of beer on the wall, ", i, " bottles of beer.")
      while(i > 0) {
        tryCatch(
          Sys.sleep(1),
          interrupt = function(err) {
            i <<- i - 1
            if (i > 0) {
              message(
                "Take one down, pass it around, ", i, 
                " bottle", if (i > 1) "s", " of beer on the wall."
              )
            }
          }
        )
      }
      message("No more bottles of beer on the wall, no more bottles of beer.")
    }
    ```
    
   __<span style="color:green">A</span>__: Basically there seems to be an issue that interrupts are not handled in a general way, but seem to depend on the R interface and the implementation of the R code which is interrupted. Not sure if I understand this correct, but there needs to be a way from the frontend to supply the r code with the info that an interrupt has happened and the r code itself needs to check for this information regularly.
   
   This issue is tracked i.e. here https://github.com/HenrikBengtsson/Wishlist-for-R/issues/47 and reported also here https://www.r-bloggers.com/interrupting-r-processes-in-ubuntu/ and more official documentation is for example under `?setTimeLimit`:
   
   > Time limits are checked whenever a user interrupt could occur. This will happen frequently in R code and during Sys.sleep, but only at points in compiled C and Fortran code identified by the code author.
   
   A minimal example pointed out in the former source would be `system.time(try(Sys.sleep(3)))` which will be hard to trigger under a certain time. The behaviour of `Sys.sleep()` is also documented under in the "Details" section of `Sys.sleep()`:
   
   > Using this function allows R to temporarily be given very low priority and hence not to interfere with more important foreground tasks. A typical use is to allow a process launched from R to set itself up and read its input files before R execution is resumed. The intention is that this function suspends execution of R expressions but wakes the process up often enough to respond to GUI events, typically every half second. It can be interrupted (e.g. by Ctrl-C or Esc at the R console). There is no guarantee that the process will sleep for the whole of the specified interval (sleep might be interrupted), and it may well take slightly longer in real time to resume execution. 
   
   However, in the exercise it seems to be more the question why it is hard to get i from `99` to `0` and the answer might be, that interruptions are insecure since it might be hard to anticipate which part of the program is running and instead of triggering a specific interruption-handler it's easy to interrupt the process as a whole. The specific behaviour of `Sys.sleep()` as mentioned above makes it even harder to guess which part of a program is currently executed.
   
   We can see this behaviour explicitly when we modify the code above a bit to wrap the `tryCatch()`-part into another `tryCatch()` part where we handle the regarding further interruptions and play with this code interactively in the console again:
   
    ```{r, eval = FALSE}
    bottles_of_beer <- function(i = 99) {
    
      message("There are ", i, " bottles of beer on the wall, ", i, " bottles of beer.")
      
      tryCatch(
        while(i > 0) {
          tryCatch(
            Sys.sleep(1),
            interrupt = function(err) {
              i <<- i - 1
              if (i > 0) {
                message(
                  "Take one down, pass it around, ", i, 
                  " bottle", if (i > 1) "s", " of beer on the wall."
                )
              }
            }
          )
        },
      interrupt = function(cnd) message("outer part"))
      
      message("No more bottles of beer on the wall, no more bottles of beer.")
    }
    ```
