# (PART) Metaprogramming {-} 

# Expressions

## Abstract syntax trees

1. __<span style="color:red">Q</span>__: Use `ast()` and experimentation to figure out the three arguments to an `if()` call. What would you call them? Which arguments are required and which are optional?
    
   __<span style="color:green">A</span>__: You can write an `if()` statement in several ways: with or without `else`, formatted or in one line and also in prefix notation. Here are several versions focussing on the possibility of leaving out curly brackets.
    
    ```{r}
    lobstr::ast(if (TRUE) {} else {})
    lobstr::ast(if (TRUE) 1 else 2)
    lobstr::ast(`if`(TRUE, 1, 2))
    ```
    
   One possible way of naming the arguments would be: condition (1), conclusion (2), alternative (3).
    
   The *condition* is always required. If the *condition* is `TRUE`, also the *conclusion* is required. If the *condition* is `FALSE` and `if()` is called in combination with `else()`, then also the *alternative* is required.
    
1. __<span style="color:red">Q</span>__: What does the call tree of an `if` statement with multiple `else if` conditions look like? Why?

   __<span style="color:green">A</span>__: The ast of nested `else if` statements might look a bit confusing because it contains multiple brackets. However, we can see that in the `else` part of the *ast* just another expression is being evaluated, which happens to be an `if` statement and so forth.
    
    ```{r}
    lobstr::ast(
    if (FALSE) {
      1
    } else if (FALSE) {
      2
    } else if (TRUE) {
      3
    }
    )
    ```
    
   We can see the structure more clearly when we avoid the curly brackets through prefix notation.
    
    ```{r}
    lobstr::ast(`if`(FALSE, 1, `if`(FALSE, 2, `if`(TRUE, 3))))
    ```

1. __<span style="color:red">Q</span>__: What are the arguments to the `for()` and `while()` calls? 

   __<span style="color:green">A</span>__: `for()` requires an *index* (called `var` in the docs), a *sequence* and an *expression*, for example
    
    ```{r}
    `for`(i, 1:3, {print(i)})
    ```
    
   `while()` requires a *condition* and an *expression*. Again, an example in prefix notation:
    
    ```{r}
    set.seed(123)
    `while`((i <- rnorm(1)) < 1, {print(i)})
    i
    ```
    
   Note that a minimal expression can consist of `{` only.
    
1. __<span style="color:red">Q</span>__: Two arithmetic operators can be used in both prefix and infix style.
   What are they?
    
   __<span style="color:green">A</span>__: I am not sure how this is meant to be. Theoretically every arithmetic operator can be written in prefix notation via backticks. On the other hand, `+` and `-` seem to be the only ones, which can be written in infix notation without backticks.
    
    ```{r}
    x <- 1

    +(x)
    -(x)
    ```
    
   However, when we look more closely, the call tree is not what we would expect from a prefix function
    
    ```{r}
    lobstr::ast(+(x))  
    lobstr::ast(-(x))  
    ```
    
   So maybe it is meant to look like this...
    
    ```{r}
    lobstr::ast(+x)
    lobstr::ast(-x)
    ```
    
   Of course also this doesn't make too much sense, since in `?Syntax` one can read, that R clearly differentiates between unary and binary `+` and `-` operators and a unary operator is not really what we mean, when we speak about infix operators.
    
   However, if we don't differentiate in this way, this is probably the solution, since it's obviously also an infix function:
    
    ```{r}
    lobstr::ast(x + y)
    lobstr::ast(x - y)
    ```

## R's grammar

1. __<span style="color:red">Q</span>__: R uses parentheses in two slightly different ways as illustrated by these two calls:

    ```{r, eval = FALSE}
    f((1))
    `(`(1 + 1)
    ```
    
   Compare and contrast the two uses by referencing the AST.
    
   __<span style="color:green">A</span>__: The trick with these examples lies in the fact, that `(` can represent a primitive function but also be a part of R's general prefix function syntax.
    
   So in the AST of the first example, we will not see the outer `(`, which belongs to `f()` and is therefore not shown in the syntax, while the inner `(` is treated as a function (symbol):
    
    ```{r}
    lobstr::ast(f((1)))
    ```
    
   In the second example, we can see that the outer `(` is treated as a function and the inner `(` belongs to its syntax:
    
    ```{r}
    lobstr::ast(`(`(1 + 1))
    ```
    
   For the sake of clarity, let's also create a third example, where none of the `(` is part of another functions syntax:
    
    ```{r}
    lobstr::ast(((1 + 1)))
    ```
    
1. __<span style="color:red">Q</span>__: `=` can also be used in two ways. Construct a simple example that shows both uses.
    
   __<span style="color:green">A</span>__: I was not exactly aware of a similar case with multiple syntactical meanings for the `=` symbol, but one can get there somehow. `=` is used as an operator for assignment. It is also part of the logical operators `==`, `>=`, `<=`, `!=` and is also used within functions to assign parameters or the definition of default settings.
    
   The question probably aims at the difference of global assignment and parameter definition within functions.
    
   So when we play with `ast()`, we can directly see that the following is not possible
    
    ```{r, error = TRUE}
    lobstr::ast(a = 1)
    ```
    
   We get an error, because `a = ` makes R looking for an argument called `a`. Since `x` is the only argument of `lobstr::ast()`, we get an error.
    
   When we build our workaround for the problem, the solution to the question becomes obvious.
    
   Instead `a = 1`, we pass the expression via brackets to `ast()`. Once via matching by position and once via matching by name
    
    ```{r}
    lobstr::ast((a = 1))
    lobstr::ast(x = (a = 1))
    ```
    
   The second way is more explicit, but both return the same syntax tree. When wee ignore the `brackets` and compare the trees, we can finally see from the second tree, that the first `=` is just part of the syntax and the second one is for the usage of assignment.

1. __<span style="color:red">Q</span>__: What does `!1 + !1` return? Why?

   __<span style="color:green">A</span>__: The first answer is quite simple
    
    ```{r}
    !1 + !1
    ```
    
   To answer the "Why", we have a look at the syntax tree first
    
    ```{r}
    lobstr::ast(!1 + !1)
    ```
    
   So first, the second `!1` becomes evaluated, which results in `FALSE`, because in R every non 0 numeric, becomes coerced to `TRUE`, when a logical operator is applied on it.
    
   Next `1 + FALSE` is evaluated to `1`, since `FALSE` is coerced to `0`.
    
   Finally `!1` is evaluated to `FALSE`, because it is the opposite of `TRUE`, which is what `1` becomes coerced to.
    
   However, note that if `!` had a higher precedence, the result would get `FALSE + FALSE` as intermediate result, which would be evalutated (again involving coercion) to `0`.

1. __<span style="color:red">Q</span>__: Why does `x1 <- x2 <- x3 <- 0` work? There are two reasons.

   __<span style="color:orange">A</span>__: One reason is that `<-` is right-associative.

1. __<span style="color:red">Q</span>__: Compare the ASTs `x + y %+% z` to `x ^ y %+% z`. What does that tell you about the precedence of custom infix functions?

   __<span style="color:green">A</span>__: Comparison of the syntax trees:
    
    ```{r}
    # for ast(x + y %+% z)
    # y %+% z will be calculated first and the result will be added to x
    lobstr::ast(x + y %+% z)
    
    # for ast(x ^ y %+% z) 
    # x ^ y will be calculated first, and the result will be used as 
    # first argument of %+%()
    lobstr::ast(x ^ y %+% z)
    ```
    
   So we can conclude that custom infix functions must have a precedence between addition and exponentiation. The general precedence rules can be found for example [here](https://cran.r-project.org/doc/manuals/r-release/R-lang.html).

## Data structures

1. __<span style="color:red">Q</span>__: Which two of the six types of atomic vector can't appear in an expression? 
    Why? Why can't you create an expression that contains an atomic vector of 
    length greater than one? 
    
   __<span style="color:green">A</span>__: It is not possible to create an expression that evaluates to an atomic of length greater than one without using a function (i.e. the `c()` function). But expressions that include a function would be calls.
    
   Let us illustrate this observation via the following example:
    
    ```{r}
    is.atomic(quote(1))       # atomic
    is.atomic(quote(c(1,1)))  # not an atomic (it would just evaluate to an atomic).
    is.call(quote(c(1,1)))    # still a call! (so at least a valid expression).
    ```
    <!-- HB: a little more explanation would be nice, I think. -->
    
   Two of the six atomic vector types of R do not work with expressions, the first one being raws. We assume, that raws may only be constructed through using `as.raw()`, but this function would then creating another call in the AST.
    
   For similar reasons complex numbers also won't work:
    
    ```{r}
    (function(x){is.atomic(x) & length(x) == 1})(quote(1 + 1.5i))
    
    # however, imaginary parts of complex numbers work:
    lobstr::ast(1i)
    ```
    <!-- HB: don't fully understand the first line of code in the above code-chunk -->
    
1. __<span style="color:red">Q</span>__: How is `rlang::maybe_missing()` implemented? Why
does it work?

   __<span style="color:green">A</span>__: Let us take a look at the functions source code to see what's going on
    
    ```{r, eval = FALSE}
    lang::maybe_missing
    function (x) 
      {
      # is_missing checks if one of the following is TRUE
      # 1. check via substitute if typeof(x) is symbol and missing(x) is TRUE
      # 2. check if x identical to missing_arg()
      if (is_missing(x)) {
        missing_arg()  # returns missing argument
                       # implemented in lower level code -> .Call())
      }
      else {
        x  # when it's not missing, x is simply returned
      }
    }
    <bytecode: 0x00000000195ed740>
    <environment: namespace:rlang>
    ```
    <!-- HB: I think, I would try to split the explanation into two parts. One Overfiew with comments in the functions source code and then some prose summarising what's going on at a little higher level. Why does it work?-->    
    
   First it is checked if the argument is missing. If so, the missing arg is returned, otherwise the argument (`x`) itsself is returned.
   
1. __<span style="color:red">Q</span>__: `rlang::call_standardise()` doesn't work so well for the following calls.
    Why? What makes `mean()` special?

    ```{r}
    library(rlang)
    
    call_standardise(quote(mean(1:10, na.rm = TRUE)))
    call_standardise(quote(mean(n = T, 1:10)))
    call_standardise(quote(mean(x = 1:10, , TRUE)))
    ```
    
   __<span style="color:green">A</span>__: The reason for this unexpected behaviour lies in the fact that `mean()` uses S3 dispatch (i.e., `UseMethod`) and therefore does not store its formals on `mean()`, but rather on `mean.default()`. `rlang::call_standardise()` can do much better when the S3 dispatch is explicit.
    
    ```{r}
    call_standardise(quote(mean.default(1:10, na.rm = TRUE)))
    call_standardise(quote(mean.default(n = T, 1:10)))
    call_standardise(quote(mean.default(x = 1:10, , TRUE)))
    ```

1. __<span style="color:red">Q</span>__: Why does this code not make sense?

    ```{r, eval = FALSE}
    x <- expr(foo(x = 1))
    names(x) <- c("x", "")
    ```
    
   __<span style="color:green">A</span>__: As stated in the book
    
   > The first element of a call is always the function that gets called.
    
   We can just look what will happen
    
    ```{r}
    x <- rlang::expr(foo(x = 1))
    x
    
    names(x) <- c("x", "")
    x
    
    names(x) <- c("", "x")
    x
    ```
    
   So giving the first element a name just adds useless metadata.

1. __<span style="color:red">Q</span>__: Construct the expression `if(x > 1) "a" else "b"` using multiple calls to `lang()`. How does the structure code reflect the structure of the AST?

   __<span style="color:green">A</span>__: Similar to the prefix version we get
    
    ```{r}
    rlang::lang("if", rlang::expr(x > 1), "a", "b")
    ```
    
   When we reed the AST from left to right, we get the same structure:
   Function to evaluate, expression, which is another function and becomes evaluated first and two constants which will be evaluated next
    
    ```{r}
    lobstr::ast(`if`(x > 1, "a", "b"))
    ```

## Parsing and deparsing

1. __<span style="color:red">Q</span>__: What happens if you attempt to parse an invalid expression? e.g. `"a +"` or `"f())"`.

   __<span style="color:green">A</span>__: We get an error from the underlying `parse` function
    
    ```{r, error = TRUE}
    rlang::parse_expr("a +")
    rlang::parse_expr("f())")
    
    parse(text = "a +")
    parse(text = "f())")
    ```

1. __<span style="color:red">Q</span>__: `deparse()` produces vectors when the input is long. For example, the following call produces a vector of length two:

    ```{r, eval = TRUE}
    expr <- rlang::expr(g(a + b + c + d + e + f + g + h + i + j + k + l + m +
      n + o + p + q + r + s + t + u + v + w + x + y + z))
    
    deparse(expr)
    ```

   What do `expr_text()`, `expr_name()`, and `expr_label()` do with this input?
    
   __<span style="color:green">A</span>__: 
    
   * `expr_text()` pastes the output string into one and inserts `\n` (new line identifiers) as separators
    
    ```{r}
    cat(rlang::expr_text(expr)) # cat is used for printing with linebreak
    ```
   
   * `expr_name()` recreates the call into the form f(...) and deparses this expression into a string
    
    ```{r}
    rlang::expr_name(expr)
    ```
    
   * `expr_label()` does the same as `expr_name()`, but surrounds the output also with backticks
    
    ```{r}
    rlang::expr_label(expr)
    ```

1. __<span style="color:red">Q</span>__: Why does as.Date.default() use substitute() and deparse()? Why does pairwise.t.test() use them? Read the source code.

   __<span style="color:green">A</span>__:

1. __<span style="color:red">Q</span>__: pairwise.t.test() assumes that deparse() always returns a length one character vector. Can you construct an input that violates this expectation? What happens?

   __<span style="color:green">A</span>__:

## Case study: Walking the AST with recursive functions {#ast-funs}

1. __<span style="color:red">Q</span>__: `logical_abbr()` returns `TRUE` for `T(1, 2, 3)`. How could you modify `logical_abbr_rec()` so that it ignores function calls that use `T` or `F`?

   __<span style="color:green">A</span>__: We can apply a similar logic as in the multiple assignment example from the textbook and just treat this case as a special case handled within a sub function called `find_T_call()` which finds `T()` calls and "bounces them out":
    
    ```{r, include = FALSE}
    library(lobstr)
    library(rlang)
    
    expr_type <- function(x) {
      if (rlang::is_syntactic_literal(x)) {
      "constant"
      } else if (is.symbol(x)) {
      "symbol"
      } else if (is.call(x)) {
      "call"
      } else if (is.pairlist(x)) {
      "pairlist"
      } else {
      typeof(x)
      }
    }
  
    switch_expr <- function(x, ...) {
      switch(expr_type(x),
      ...,
      stop("Don't know how to handle type ", typeof(x), call. = FALSE))
    }
    ```
    
    ```{r}
    find_T_call <- function(x) {
      if (is_call(x, "T")) {
      x <- as.list(x)[-1]
      purrr::some(x, logical_abbr_rec)
      } else {
      purrr::some(x, logical_abbr_rec)
      }
    }
    
    logical_abbr_rec <- function(x) {
      switch_expr(
      x,
      # Base cases
      constant = FALSE,
      symbol = as_string(x) %in% c("F", "T"),
      
      # Recursive cases
      pairlist = purrr::some(x, logical_abbr_rec),
      call = find_T_call(x)
      )
    }
    
    logical_abbr <- function(x) {
      logical_abbr_rec(enexpr(x))
    }
    ```
    
    Now lets test our new `logical_abbr()` function:
    
    ```{r}
    logical_abbr(T(1, 2, 3))
    logical_abbr(T(T, T(3, 4)))
    logical_abbr(T(T))
    logical_abbr(T())
    logical_abbr()
    logical_abbr(c(T, T, T))
    ```

1. __<span style="color:red">Q</span>__: `logical_abbr()` works with expressions. It currently fails when you give it a function. Why not? How could you modify `logical_abbr()` to make it work? What components of a function will you need to recurse over?

    ```{r, eval = FALSE}
    f <- function(x = TRUE) {
      g(x + T)
    }
    logical_abbr(!!f)
    ```
    
   __<span style="color:green">A</span>__: It currently fails, because `"closure"` is not handled within `switch_expr()` within `logical_abbr_rec()`. If we wanted to make it work, we must open a case there and write a function to inspect the formals and the body of the input function.

1. __<span style="color:red">Q</span>__: Modify find assignment to also detect assignment using replacement functions, i.e. `names(x) <- y`.

   __<span style="color:green">A</span>__: Let`s see what the AST of such an assignment looks like:
    
    ```{r}
    ast(names(x) <- x)
    ```

   So we need to catch the case where the first two elements are both calls. Further the first call is identical to `<-` and we must return only the second call to see which objects got new values assigned.
    
   This is why we add the following block Within another `else` statement in `find_assign_call()`:
    
    ```{r, eval = FALSE}
    if (is_call(x, "<-") && is_call(x[[2]])) {
      lhs <- expr_text(x[[2]])
      children <- as.list(x)[-1]
    }
    ```
      
   Let us finish with the whole code including some tests for our new function:
    
    ```{r}
    flat_map_chr <- function(.x, .f, ...) {
      purrr::flatten_chr(purrr::map(.x, .f, ...))
    }
    
    find_assign <- function(x) unique(find_assign_rec(enexpr(x)))
    
    find_assign_call <- function(x) {
      if (is_call(x, "<-") && is_symbol(x[[2]])) {
        lhs <- as_string(x[[2]])
        children <- as.list(x)[-1]
      } else {
      if (is_call(x, "<-") && is_call(x[[2]])) {
        lhs <- expr_text(x[[2]])
        children <- as.list(x)[-1]
      } else {
        lhs <- character()
        children <- as.list(x)
      }}
      
      c(lhs, flat_map_chr(children, find_assign_rec))
    }
    
    find_assign_rec <- function(x) {
      switch_expr(
        x,
        # Base cases
        constant = ,symbol = character(),
        # Recursive cases
        pairlist = flat_map_chr(x, find_assign_rec),
        call = find_assign_call(x)
      )
    }
    
    find_assign(x <- y)
    find_assign(names(x))
    find_assign(names(x) <- y)
    find_assign(names(x(y)) <- y)
    find_assign(names(x(y)) <- y <- z)
    ```
  
1. __<span style="color:red">Q</span>__: Write a function that extracts all calls to a specified function.

   __<span style="color:green">A</span>__: We just need to delete the former added else statement and check for a call (not necessarily `<-`) within the first `if()` in `find_assign_call()`. We save a call when we found one and return it later as part of our character output. Everything else stays the same:

    ```{r}
    find_assign_call <- function(x) {
      if (is_call(x)) {
        lhs <- expr_text(x)
        children <- as.list(x)[-1]
        } else {
          lhs <- character()
          children <- as.list(x)
        }
      
      c(lhs, flat_map_chr(children, find_assign_rec))
    }
    
    find_assign_rec <- function(x) {
      switch_expr(x,
                  # Base cases
                  constant = ,
                  symbol = character(),
                  
                  # Recursive cases
                  pairlist = flat_map_chr(x, find_assign_rec),
                  call = find_assign_call(x)
      )
    }
    
    find_assign(x <- y)
    find_assign(names(x(y)) <- y <- z)
    find_assign(mean(sum(1:3)))
    ```
