```{r, include=FALSE}
source("common.R")
```

# S4

## Prerequisites

We load the methods package as it contains the S4 object-orientation system.

```{r, message = FALSE}
library(methods)
```

## Basics

1. __[Q]{.Q}__: `lubridate::period()` returns an S4 class. What slots does it have? What class is each slot? What accessors does it provide?
    
   __[A]{.solved}__: Objects of the S4 `Period` class have six slots named `year`, `month`, `day`, `hour`, `minute`, and `.Data` (which cotains the number of seconds). All slots are of type double. Most fields can be retrieved by identically named accessor (e.g. `lubridate::year()` will return the field), using `second()` to get the the `.Data` slot.

   As a short example, we';; create a period of 1 second, 2 minutes, 3 hours, 4 days and 5 weeks.
   
    ```{r}
    example_12345 <- lubridate::period(
      c(1, 2, 3, 4, 5), 
      c("second", "minute", "hour", "day", "week")
    )
    ```
    
   This should add up to a period of 39 days, 3 hours, 2 minutes and 1 second.
   
    ```{r}
    example_12345
    ```
    
   When we inspect `example_12345`, we see the fields and infer that the seconds are stored in the `.Data` field.
   
    ```{r}
    str(example_12345)
    ```

2. __[Q]{.Q}__: What other ways can you find help for a method? Read `?"?"` and summarise the details.
    
   __[A]{.solved}__: Besides adding `?` in front of a function call (`?method()`), we may find:
    
   * general documentation for a generic via `?genericName`
   * general documentation for the methods of a generic via `methods?genericName`
   * documentation for a specific method via `ClassName?methodName`.
    
## Classes

1. __[Q]{.Q}__: Extend the Person class with fields to match `utils::person()`. Think about what slots you will need, what class each slot should have, and what you’ll need to check in your validity method.
    
   __[A]{.solved}__: The Person class from the textbook contains the slots `name` and `age`. The person class from the utils package contains the slots `given`, `family`, `role`, `email` and `comment`. All these slots must be of type character.
   
   The entries in the `role` slot must match one of the following abbreviations "aut", "com", "cph", "cre", "ctb", "ctr", "dtc", "fnd", "rev", "ths", "trl". Therefore we include all these slots in our new definition of the `Person` class. `role` might be of different length than the other slots. Because of this we add a constraint to the validator, that all slots must be of length one.
   
    ```{r}
    # Definition of the Person class
    setClass("Person", 
         slots = c(
           age = "numeric",
           given = "character",
           family = "character",
           role = "character",
           email = "character",
           comment = "character"
         ),
         prototype = list(
           age = NA_real_,
           given = NA_character_,
           family = NA_character_,
           role = NA_character_,
           email = NA_character_,
           comment = NA_character_
         )
    )
    
    # Helper to create instances of the Person class
    Person <- function(given, family, 
                       age = NA_real_, 
                       role = NA_character_,
                       email = NA_character_,
                       comment = NA_character_) {
      age <- as.double(age)
      
      new("Person", 
        age = age, 
        given = given, 
        family = family, 
        role = role, 
        email = email,
        comment = comment
      )
    }

    # Validator to ensure that each slot is of length one
    setValidity("Person", function(object) {
      if (length(object@age)     != 1 ||
          length(object@given)   != 1 ||
          length(object@family)  != 1 ||
          length(object@email)   != 1 ||
          length(object@comment) != 1) {
        "@name, @age, @given, @family, @email, @comment must be of length 1"
      } 
  
      known_roles <- c(
        NA_character_, "aut", "com", "cph", "cre", "ctb",
        "ctr", "dtc", "fnd", "rev", "ths", "trl"
      )
      
      if (!all(object@role %in% known_roles)) {
        paste(
          "@role(s) must be one of", 
          paste(known_roles, collapse = ", ")
        )
      }
  
      TRUE
    })
    ```

2. __[Q]{.Q}__: What happens if you define a new S4 class that doesn't have any slots?  (Hint: read about virtual classes in `?setClass`.)
    
   __[A]{.solved}__: It depends on the other arguments. If we inherit
   from another class, we get the same slots. But something interesting 
   happens if we don't inherit from an existing class. We get a virtual
   class. A virtual class can't be instantiated:
    
    ```{r, eval = TRUE, error = TRUE}
    setClass("Human")
    new("Human")
    ```
    
    But can be inherited from:
    
    ```{r}
    setClass("Programmer", contains = "Human")
    ```

3. __[Q]{.Q}__: Imagine you were going to reimplement factors, dates, and data frames in S4. Sketch out the `setClass()` calls that you would use to define the classes. Think about appropriate `slots` and `prototype`.

   __[A]{.solved}__: The basic idea is to use a slot for the data and one slot per attribute. Inheritance matters for ordered factors and dates. Special checks like equal lengths of list elements for columns of a data frame should be done within a validator.
   
   For simplicity we don't introduce an own subclass for ordered factors. Instead, we introduce `ordered` as a slot.
   
    ```{r}
    setClass("Factor", 
      slots = c(
        data = "integer",
        levels = "character",
        ordered = "logical"
      ),
      prototype = list(
        data = integer(),
        levels = character(),
        ordered = FALSE
      )
    )
 
    abc <- new("Factor", data = c(1L, 2L), levels = letters[1:3])
    abc
    ```
    
   The `Date2` class stores its dates as integers, similarly to base R. Dates don't have any other attributes.
   
    ```{r}
    setClass("Date2",
      slots = list(
        data = "integer"
      ),
      prototype = list(
        data = integer()
      )
    )
         
    new("Date2")
    ```
    
   Our `DataFrame` class consists of a list and a slot for `row.names`. Most of the logic (e.g. checking that all elements of the list are a vector, and that they all have the same length) would need to be part of a validator.
   
    ```{r}
    setClass("DataFrame",
      slots = c(
        data = "list", 
        row.names = "character"
      ),
      prototype = list(
        data = list(),
        row.names = character(0)
      )
    )
         
    df <- new("DataFrame", data = list(a = 1, b = 2))         
    df
    ```

## Generics and methods

1. __[Q]{.Q}__: Add `age()` accessors for the `Person` class.

   __[A]{.solved}__: Similar as shown for `name()` in the chapter, we define an `age()` generic, with a method for the `Person` class and a replacement function `age<-()`:
   
    ```{r}
    setGeneric("age", function(x) standardGeneric("age"))
    setMethod("age", "Person", function(x) x@age)
    
    setGeneric("age<-", function(x, value) standardGeneric("age<-"))
    setMethod("age<-", "Person", function(x, value) {
      x@age <- value
      validObject(x)
      x
    })
    ```

2. __[Q]{.Q}__: In the definition of the generic, why is it necessary to repeat the name of the generic twice?

   __[A]{.solved}__: First the name is needed as the name of the generic. The name also explicitly incorporates method dispatch via `standardGeneric()` within the generic's body (`def` parameter). This behaviour is similar to `UseMethod()` in S3.

3. __[Q]{.Q}__: Why does the `show()` method defined in Section 15.4.3 use `is(object)[[1]]`? (Hint: try printing the employee subclass.)

   __[A]{.solved}__: `is(object)` returns the class of the object. `is(object)` also contains the superclass, for subclasses like `Employee`. In order to return always the most specific class (the subclass), `show()` returns the first element of `is(object)`.

4. __[Q]{.Q}__: What happens if you define a method with different argument names to the generic?

   __[A]{.solved}__: It depends. Let's first create the object `hadley` of class "Person":
    
    ```{r}
    .Person <- setClass(
      "Person", 
      slots = c(name = "character", age = "numeric")
    )
    
    hadley <- .Person(name = "Hadley")
    hadley
    ```
    
   Now let's see, which arguments can be supplied to the `show()` generic.
    
    ```{r}
    formals("show")
    ```
    
   Usually we would use this argument when defining a new method
    
    ```{r}
    setMethod("show", "Person", function(object) {
      cat(object@name, "creates hard exercises")
    })
    hadley
    ```
    
   When we supply another name as a first element of our method (e.g. `x` instead of `object`), this will be matched to the correct `object` argument and we receive a warning. Our method will work, though:
    
    ```{r, eval = TRUE}
    setMethod("show", "Person", function(x) {
      cat(x@name, "creates hard exercises")
    })
    hadley
    ```
    
   If we add more arguments to our method than our generic can handle, we will get an error.
    
    ```{r, eval = TRUE, error = TRUE}
    setMethod("show", "Person", function(x, y) {
      cat(x@name, "is", x@age, "years old")
    })
    ```
    
   If we do this with arguments added to the correctly written `object` argument, we will get an informative error message. It states, that we could add other argument names for generics, which can take the `...` argument.
    
    ```{r, eval = TRUE, error = TRUE}
    setMethod("show", "Person", function(object, y) {
      cat(object@name, "is", object@age, "years old")
    })
    ```

## Method dispatch 

1. __[Q]{.Q}__: Draw the method graph for `f(😅, 😽)`.

   __<span style="color:yellow">A</span>__: TODO: This should be straight forward. We just need the class graph from above and insert the graph including the superclasses for both emojis into `f(,)`. Everything then should follow the logic from the chapter...

2. __[Q]{.Q}__: Draw the method graph for `f(😃, 😉, 😙)`.

   __[A]{.solved}__: TODO: The same as in the last exercise. The challenge here might be to visualize the dispatch efficiently for the combination of three arguments at once.

3. __[Q]{.Q}__: Take the last example which shows multiple dispatch over two classes that use multiple inheritance. What happens if you define a method for all terminal classes? Why does method dispatch not save us much work here?

   __[A]{.solved}__: We will introduce ambiguity, since one class has distance 2 to all terminal nodes and the other four have distance 1 to two terminal nodes each. To resolve this ambiguity we have to define five more methods, one per class combination.


## S4 and S3

1. __[Q]{.Q}__: What would a full `setOldClass()` definition look like for an ordered factor (i.e. add `slots` and `prototype` the definition above)?

   __[A]__[A]{.open
   
   <!-- HW: I think this could just inherit from the Factor class defined in the text -->
   
2. __[Q]{.Q}__: Define a `length` method for the `Person` class.

   __[A]{.solved}__: We can define it as either as an S3 or S4 method:
   
    ```{r}
    length.Person <- function(x) "a"
    setMethod("length", "Person", function(x) "a")
    ```
