```{r, include=FALSE}
source("common.R")
```

# S4

## Prerequisites

We load the methods package for anything specifically related to S4 and the lubridate package for the first exercise.

```{r, message = FALSE}
library(methods)
library(lubridate)
```

## Basics

1. __[Q]{.Q}__: `lubridate::period()` returns an S4 class. What slots does it have? What class is each slot? What accessors does it provide?
    
   __[A]{.solved}__: Objects of the S4 `Period` class have the six slots `.Data`, `year`, `month`, `day`, `hour` and `minute`, which are each of type double. Except for `.Data` all fields have similar named accessors. `.Data` can be accessed via `methods::gedDataPart()`. As a short example, we create a period of 1 second, 2 minutes, 3 hours, 4 days and 5 weeks.
   
    ```{r}
    example12345 <- period(c(1, 2, 3, 4, 5), 
              c("second", "minute", "hour", "day", "week"))
    ```
    
   This should add up to a period of 39 days, 3 hours, 2 minutes and 1 second.
   
    ```{r}
    example12345
    ```
    
   When we inspect our `example12345`, we can see the fields and conclude that the seconds are stored in the `.Data` field.
   
    ```{r}
    str(example12345)
    ```

2. __[Q]{.Q}__: What other ways can you find help for a method? Read `?"?"` and summarise the details.
    
   __[A]{.solved}__: Besides the already mentioned example of adding `?` in front of a function call (`?method(john)`), we can find:
    
   * general documentation for a generic via `?genericName`
   * general documentation for the methods of a generic via `methods?genericName`
   * documentation for a specific method via `ClassName?methodName`.
    
## Classes

1. __[Q]{.Q}__: Extend the Person class with fields to match `utils::person()`. Think about what slots you will need, what class each slot should have, and what you’ll need to check in your validity method.
    
   __[A]{.solved}__: The Person class from the textbook contains the slots `name` and `age`. The person class from the utils package contains the slots `given`, `family`, `role`, `email` and `comment`. All these slots must be of type character. Further, the entries in the `role` slot must match one of the following abbreviations "aut", "com", "cph", "cre", "ctb", "ctr", "dtc", "fnd", "rev", "ths", "trl". Therefore, we include all these slots in our new definition of the `Person` class. As `role` might be of different length than the other slots, we include the constraint that all slots must be of length one to the validator.
   
    ```{r}
    # Definition of the Person class
    setClass("Person", 
         slots = c(
           name = "character", 
           age = "numeric",
           given = "character",
           family = "character",
           role = "character",
           email = "character",
           comment = "character"
         ),
         prototype = list(
           name = NA_character_,
           age = NA_real_,
           given = NA_character_,
           family = NA_character_,
           role = NA_character_,
           email = NA_character_,
           comment = NA_character_
         )
    )
    
    # Helper to create instances of the Person class
    Person <- function(name, age = NA, 
                   given = NA_character_,
                   family = NA_character_,
                   role = NA_character_,
                   email = NA_character_,
                   comment = NA_character_) {
      age <- as.double(age)
      
      new("Person", name = name, age = age, 
      given = given, family = family, 
      role = role, email = email,
      comment = comment)
    }

    # Validator to ensure that each slot is of length one
    setValidity("Person", function(object) {
      if (length(object@name)    != 1 |
          length(object@age)     != 1 |
          length(object@given)   != 1 |
          length(object@family)  != 1 |
          length(object@email)   != 1 |
          length(object@comment) != 1) {
        "@name, @age, @given, @family, @email, @comment must be of length 1"
      } 
  
      if (!all(object@role %in% c(NA_character_, 
            "aut", "com", "cph", "cre", "ctb",
            "ctr", "dtc", "fnd", "rev", "ths", "trl"))) {
        paste("@role (s) must be one of", 
              paste (c(NA_character_, 
                                     "aut", "com", "cph", "cre", "ctb",
                                     "ctr", "dtc", "fnd", "rev", "ths", "trl"),
                     collapse = ", "), ".")
      }
  
      TRUE
    })
    ```

2. __[Q]{.Q}__: What happens if you define a new S4 class that doesn’t have any slots?  (Hint: read about virtual classes in `?setClass`.)
    
   __[A]{.solved}__: It depends on the other arguments.
   
   If we supply a class that doesn't exist, we'll get an error
    
    ```{r, error = TRUE}
    setClass("Programmer",
             slots = c(skill = "ANY"),
             contains = "Human")
    ```
    
   To can get around that, we register the new class before we define the new class.
   
    ```{r}
    setOldClass("Human")
    .Programmer <- setClass("Programmer",
                            slots = c(Skill = "ANY"),
                            contains = "Human")
    ```
    
   Supplying neither `slots` nor `contains` results in a constructor for virtual classes 
    
    ```{r}
    .VirtualProgrammer <- setClass("VirtualProgrammer")
    # equal to contains = "VIRTUAL" (here you could also supply slots)
    .VirtualProgrammer <- setClass("VirtualProgrammer",
                                   contains = "VIRTUAL")
    ```
    
   Just leaving out `contains`, but supplying slots results in a constructor without superclass
    
    ```{r}
    .DataScientist <- setClass("RProgrammer",
                               slots = c(stats = "ANY",
                                         math = "ANY",
                                         programming = "ANY"))
    ```

3. __[Q]{.Q}__: Imagine you were going to reimplement factors, dates, and data frames in S4. Sketch out the `setClass()` calls that you would use to define the classes. Think about appropriate `slots` and `prototype`.

   __[A]{.solved}__: The basic idea is to use a slot for the base type and one slot per attribute. Inheritance matters for ordered factors and dates. Special checks like equal lengths of list elements for columns of a data frame should be done within a validator.
   
   For simplicity we don't introduce an own subclass for ordered factors. Instead, we introduce `ordered` as a slot.
   
    ```{r}
    setClass("Factor", 
          slots = c(
            x = "character",
            levels = "character",
            ordered = "logical"
          ),
          prototype = list(
            x = character(0),
            levels = character(0),
            ordered = FALSE
          )
    )
 
    abc <- new("Factor", x = c("a", "b", "c"))
    abc
    ```
    
   The `Date2` class stores it's Dates as integers, similarly as the base R variant. It's pretty obvious, that the next step would be to provide a print method for our new class.
   
    ```{r}
    setClass("Date2",
         slots = c(
           Date = "integer", 
           format = "character",
           origin = "integer"
           tz = "character"
         ),
         prototype = list(
           Date = integer(),
           format = "%Y-%m-%d",
           origin = 0L,
           tz = "UTC"
         )
    )
         
    new("Date2")
    ```
    
   Our `DataFrame` class consists of a list and a slot for `row.names`. More serious checks should be part of a validator.
   
    ```{r}
    setClass("DataFrame",
         slots = c(
           x = "list", 
           row.names = "character"
         ),
         prototype = list(
           x = list(),
           row.names = character(0)
           )
         )
         
    df <- new("DataFrame", 
              x = list(a = 1, b = 2))         
    df
    ```

## Generics and methods

1. __[Q]{.Q}__: Add `age()` accessors for the `Person` class.

   __[A]{.solved}__: Similar as shown for `name()` in the chapter, we define an `age()` generic, with a method for the `Person` class and a replacement function `age<-()`:
   
    ```{r}
    setGeneric("age", function(x) standardGeneric("age"))
    setMethod("age", "Person", function(x) x@age)
    
    setGeneric("age<-", function(x, value) standardGeneric("age<-"))
    setMethod("age<-", "Person", function(x, value) {
      x@age <- value
      validObject(x)
      x
    })
    ```

2. __[Q]{.Q}__: In the definition of the generic, why is it necessary to repeat the name of the generic twice?

   __[A]{.solved}__: The name is needed as the name of the generic as well as to explicitly incorporate method dispatch via `standardGeneric()` within the generic's body (`def` parameter). This is similar to `UseMethod()` in S3.

3. __[Q]{.Q}__: Why does the `show()` method defined in Section 15.4.3 use `is(object)[[1]]`? (Hint: try printing the employee subclass.)

   __[A]{.solved}__: `is(object)` returns the class of the object. In cases of subclasses like `Employee`, `is(object)` contains also the superclass. In order to return always the most specific class (the subclass), `show()` returns the first element of `is(object)`.

4. __[Q]{.Q}__: What happens if you define a method with different argument names to the generic?

   __[A]{.solved}__: It depends. Lets first create the object `hadley` of class "Person":
    
    ```{r}
    .Person <- setClass("Person", 
                        slots = c(name = "character", 
                                  age = "numeric"))
    
    hadley <- .Person(name = "Hadley")
    hadley
    ```
    
   Now let us see, which arguments can be supplied to the `show()` generic
    
    ```{r}
    formals("show")
    ```
    
   Usually we would use this argument when defining a new method
    
    ```{r}
    setMethod("show", "Person", 
              function(object){
                cat(object@name, "creates hard exercises")
              })
    hadley
    ```
    
   When we supply another name as a first element of our method (e.g. `x` instead of `object`), this will be matched to the correct `object` argument and we receive a warning. Our method will work, though
    
    ```{r, eval = TRUE}
    setMethod("show", "Person", 
              function(x){
                cat(x@name, "creates hard exercises")
              })
    hadley
    ```
    
   If we add more arguments to our method than our generic can handle, we will get an error
    
    ```{r, eval = TRUE, error = TRUE}
    setMethod("show", "Person", 
              function(x, y){
                cat(x@name, "is", x@age, "years old")
              })
    ```
    
   If we do this with arguments added to the correctly written `object` argument, we will get the informative error message, that we could add other argument names for generics, which can take the `...` argument
    
    ```{r, eval = TRUE, error = TRUE}
    setMethod("show", "Person", 
              function(object, y){
                cat(object@name, "is", object@age, "years old")
              })
    ```

## Method dispatch 

1. __[Q]{.Q}__: Draw the method graph for `f(😅, 😽)`.

   __<span style="color:yellow">A</span>__: TODO: This should be straight forward. We just need the class graph from above and insert the graph including the superclasses for both emojis into `f(,)`. Everything then should follow the logic from the chapter...

2. __[Q]{.Q}__: Draw the method graph for `f(😃, 😉, 😙)`.

   __[A]{.solved}__: TODO: The same as in the last exercise. The challenge here might be to visualize the dispatch efficiently for the combination of three arguments at once.

3. __[Q]{.Q}__: Take the last example which shows multiple dispatch over two classes that use multiple inheritance. What happens if you define a method for all terminal classes? Why does method dispatch not save us much work here?

   __[A]{.solved}__: We will introduce ambiguity, since one class has distance 2 to all terminal nodes and the other four have distance 1 to two terminal nodes each. To resolve this ambiguity we have to define five more methods, one per class combination.


## S4 and S3

1. __[Q]{.Q}__: What would a full `setOldClass()` definition look like for an ordered factor (i.e. add `slots` and `prototype` the definition above)?

   __[A]{.solved}__:

2. __[Q]{.Q}__: Define a `length` method for the `Person` class.

   __[A]{.solved}__: We can define this method as an S3 method and register it afterwards:
   
    ```{r}
    length.Person <- function(x) "a"
    setMethod("length", "Person", length.Person)
    ```
