```{r, include=FALSE}
source("before_chapter.R")
```

# S4

## Basics

1. __<span style="color:red">Q</span>__: `lubridate::period()` returns an S4 class. What slots does it have? What class is each slot? What accessors does it provide?
    
   __<span style="color:green">A</span>__: 

2. __<span style="color:red">Q</span>__: What other ways can you find help for a method? Read `?"?"` and summarise the details.
    
   __<span style="color:green">A</span>__: We can find
    
   * general documentation for a generic via `?genericName`
   * general documentation for the methods of a generic via `methods?genericName`
   * documentation for a specific method via `ClassName?methodName`
    
   We can also get help for a specific method by adding `?` in front of a function call, e.g. `?show(hadley)`.

## Classes
1. __<span style="color:red">Q</span>__: Extend the Person class with fields to match `utils::person()`. Think about what slots you will need, what class each slot should have, and what you’ll need to check in your validity method.
    
   __<span style="color:green">A</span>__: 

2. __<span style="color:red">Q</span>__: What happens if you define a new S4 class that doesn’t have any slots?  (Hint: read about virtual classes in `?setClass`.)
    
   __<span style="color:green">A</span>__: It depends on the other arguments.
   
   If we supply a class that doesn't exist, we'll get an error
    
    ```{r, error = TRUE}
    setClass("Programmer",
             slots = c(skill = "ANY"),
             contains = "Human")
    ```
    
   To can get around that, we register the new class before we define the new class.
   
    ```{r}
    setOldClass("Human")
    .Programmer <- setClass("Programmer",
                            slots = c(Skill = "ANY"),
                            contains = "Human")
    ```
    
   Supplying neither `slots` nor `contains` results in a contructor for virtual classes 
    
    ```{r}
    .VirtualProgrammer <- setClass("VirtualProgrammer")
    # equal to contains = "VIRTUAL" (here you could also supply slots)
    .VirtualProgrammer <- setClass("VirtualProgrammer",
                                   contains = "VIRTUAL")
    ```
    
   Just leaving out `contains`, but supplying slots results in a constructor without superclass
    
    ```{r}
    .DataScientist <- setClass("RProgrammer",
                               slots = c(stats = "ANY",
                                         math = "ANY",
                                         programming = "ANY"))
    ```

3. __<span style="color:red">Q</span>__: Imagine you were going to reimplement factors, dates, and data frames in S4. Sketch out the `setClass()` calls that you would use to define the classes. Think about appropriate `slots` and `prototype`.

   __<span style="color:orange">A</span>__: The basic idea is to use a slot for the base type and one slot per attribute. Inheritance matters for ordered factors and dates. Special checks like equal lengths of list elements for columns of a data frame should be done within a validator.

## Generics and methods

1. __<span style="color:red">Q</span>__: Add `age()` accessors for the Person class.

   __<span style="color:green">A</span>__:

2. __<span style="color:red">Q</span>__: In the defintion of the generic, why is it necessary to repeat the name of the generic twice?

   __<span style="color:green">A</span>__: The name is needed as the name of the generic as well as to explicitly incorporate method dispatch via `standardGeneric()` within the generic's body (`def` parameter). This is similar to `UseMethod()` in S3.

3. __<span style="color:red">Q</span>__: Why does the `show()` method defined in Section 15.4.3 use i`s(object)[[1]]`? (Hint: try printing the employee subclass.)

   __<span style="color:green">A</span>__:

4. __<span style="color:red">Q</span>__: What happens if you define a method with different argument names to the generic?

   __<span style="color:green">A</span>__: It depends. Lets first create the object `hadley` of class "Person":
    
    ```{r}
    .Person <- setClass("Person", 
                        slots = c(name = "character", 
                                  age = "numeric"))
    
    hadley <- .Person(name = "Hadley")
    hadley
    ```
    
   Now let us see, which arguments can be supplied to the `show()` generic
    
    ```{r}
    formals("show")
    ```
    
   Usually we would use this argument when defining a new method
    
    ```{r}
    setMethod("show", "Person", 
              function(object){
                cat(object@name, "creates hard exercises")
              })
    hadley
    ```
    
   When we supply another name as a first element of our method (e.g. `x` instead of `object`), this will be matched to the correct `object` argument and we recieve a warning. Our method will work, though
    
    ```{r, eval = TRUE}
    setMethod("show", "Person", 
              function(x){
                cat(x@name, "creates hard exercises")
              })
    hadley
    ```
    
   If we add more arguments to our method than our generic can handle, we will get an error
    
    ```{r, eval = TRUE, error = TRUE}
    setMethod("show", "Person", 
              function(x, y){
                cat(x@name, "is", x@age, "years old")
              })
    ```
    
   If we do this with arguments added to the correctly written `object` argument, we will get the informative error message, that we could add other argument names for generics, which can take the `...` argument
    
    ```{r, eval = TRUE, error = TRUE}
    setMethod("show", "Person", 
              function(object, y){
                cat(object@name, "is", object@age, "years old")
              })
    ```

## Method dispatch 
1. __<span style="color:red">Q</span>__: Draw the method graph for `f(😅, 😽)`.

   __<span style="color:green">A</span>__:

2. __<span style="color:red">Q</span>__: Draw the method graph for `f(😃, 😉, 😙)`.

   __<span style="color:green">A</span>__:

3. __<span style="color:red">Q</span>__: Take the last example which shows multiple dispatch over two classes that use multiple inheritance. What happens if you define a method for all terminal classes? Why does method dispatch not save us much work here?

   __<span style="color:green">A</span>__: We will introduce ambiguity, since one class has distance 2 to all terminal nodes and the other four have distance 1 to two terminal nodes each. To resolve this ambiguity we have to define five more methods, one per class combination.


## S4 and S3
1. __<span style="color:red">Q</span>__: What would a full `setOldClass()` definition look like for an ordered factor (i.e. add `slots` and `prototype` the definition above)?

   __<span style="color:green">A</span>__:

2. __<span style="color:red">Q</span>__: Define a `length` method for the `Person` class.

   __<span style="color:green">A</span>__:
