```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE
)
```
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE
)
```
# Translating R Code

### HTML

1.  __<span style="color:red">Q</span>__: The escaping rules for `<script>` and `<style>` tags are different: you
    don't want to escape angle brackets or ampersands, but you do want to
    escape `</script>` or `</style>`.  Adapt the code above to follow these
    rules.
    
    __<span style="color:green">A</span>__: 

1.  __<span style="color:red">Q</span>__: The use of `...` for all functions has some big downsides. There's no
    input validation and there will be little information in the
    documentation or autocomplete about how they are used in the function. 
    Create a new function that, when given a named list of tags and their   
    attribute names (like below), creates functions which address this problem.

    ```{r, eval = FALSE}
    list(
      a = c("href"),
      img = c("src", "width", "height")
    )
    ```

    All tags should get `class` and `id` attributes.
    
    __<span style="color:green">A</span>__: 

1.  __<span style="color:red">Q</span>__: Currently the HTML doesn't look terribly pretty, and it's hard to see the
   structure. How could you adapt `tag()` to do indenting and formatting?
   
    __<span style="color:green">A</span>__: 

1.  __<span style="color:red">Q</span>__: Reason about the following code that calls `with_html()` referening objects
    from the environment. Will it work or fail? Why? Run the code to 
    verify your predictions.
    
    ```{r, eval = FALSE}
    greeting <- "Hello!"
    with_html(p(greeting))
    
    address <- "123 anywhere street"
    with_html(p(address))
    ```
    
    __<span style="color:green">A</span>__: 
    
## LaTeX

1.  __<span style="color:red">Q</span>__: Add escaping. The special symbols that should be escaped by adding a backslash
    in front of them are `\`, `$`, and `%`. Just as with HTML, you'll need to 
    make sure you don't end up double-escaping. So you'll need to create a small 
    S3 class and then use that in function operators. That will also allow you 
    to embed arbitrary LaTeX if needed.
    
    __<span style="color:green">A</span>__: 

1.  __<span style="color:red">Q</span>__: Complete the DSL to support all the functions that `plotmath` supports.

    __<span style="color:green">A</span>__: 
