# Translating R Code

### HTML

1.  __<span style="color:red">Q</span>__: The escaping rules for `<script>` and `<style>` tags are different: you don't want to escape angle brackets or ampersands, but you do want to escape `</script>` or `</style>`.  Adapt the code above to follow these rules.

    ```{r required code from advr, include=FALSE}
    library(rlang)
    library(purrr)

    html <- function(x) structure(x, class = "advr_html")
    cat_line <- function(...) cat(..., "\n", sep = "")

    print.advr_html <- function(x, ...) {
      out <- paste0("<HTML> ", x)
      cat_line(paste(strwrap(out), collapse = "\n"))
    }
    
    escape <- function(x) UseMethod("escape")
    
    escape.character <- function(x) {
      x <- gsub("&", "&amp;", x)
      x <- gsub("<", "&lt;", x)
      x <- gsub(">", "&gt;", x)
      html(x)
    }
    
    escape.advr_html <- function(x) x
    
    dots_partition <- function(...) {
      dots <- dots_list(...)
      is_named <- names(dots) != ""
      
      list(
        named = dots[is_named],
        unnamed = dots[!is_named]
      )
    }

    # source("dsl-html-attributes.r", local = TRUE)
    html_attributes <- function(list) {
      if (length(list) == 0) return("")
      attr <- map2_chr(names(list), list, html_attribute)
      paste0(" ", unlist(attr), collapse = "")
    }

    html_attribute <- function(name, value = NULL) {
      if (length(value) == 0) return(name) # for attributes with no value
      if (length(value) != 1) stop("`value` must be NULL or length 1")
      if (is.logical(value)) {
        # Convert T and F to true and false
        value <- tolower(value)
      } else {
        value <- escape_attr(value)
      }
      paste0(name, "='", value, "'")
    }

    escape_attr <- function(x) {
      x <- escape.character(x)
      x <- gsub("\'", '&#39;', x)
      x <- gsub("\"", '&quot;', x)
      x <- gsub("\r", '&#13;', x)
      x <- gsub("\n", '&#10;', x)
      x
    }
    ```

    __<span style="color:green">A</span>__: In regular html we need to escape angle brackets and ampersands, so they will be rendered correctly in the browser. Without escaping they would insteadt indicate tag information (`<>`) or a reference (`&`).

```{r, include = FALSE}
#TODO: - nochmal Hintergrund klären, wieso \/script-escaping notwendig ist
```

    Sourcecode wrapped in `<script>` or `<style>` tags is different:
    
    * The `<script>` tag is used to insert (mainly) JavaScript into the html source. Here the angle brackets and ampersands should not be escaped so the code remains intact. Because JavaScript may contain multiple nested `<script>` tags, we need to escape the inner tags as `<\/script>`.
    
    * `style` tags encapsulate css-styling guidelines and the escaping follows the same rules as for the javascript above...
    
    So for tags, we have two competing requirements for escaping. For most of the tag-functions the content needs to be escaped (-> brackets and ampersands), but for script and style the content should NOT be escaped, but closing `script`- and `style`-tags need to be escaped.
    
    To distinguish between these options, we extend `escape.character()` to include an argument (`escape_choice`) to choose the escape-style, that we want.
    
    When we create the tag functions we can then specify the escape style we want:
 
    ```{r}
    escape <- function(x, ...) UseMethod("escape")
    # add argument to the escape-function
    escape.character <- function(x, escape_choice = c("content", "script_or_style")) {
      escape_choice <- match.arg(escape_choice)
      
      if(escape_choice == "content") {
        x <- gsub("&", "&amp;", x)
        x <- gsub("<", "&lt;", x)
        x <- gsub(">", "&gt;", x)
      }

      if(escape_choice == "script_or_style") {
        x <- gsub("</script>", "<\\/script>", x, fixed = TRUE)
        x <- gsub("</style>",  "<\\/style>",  x, fixed = TRUE)
      }

      html(x)
    }

    escape.advr_html <- function(x, ...) x

    # create tag with specified escape-style
    tag <- function(tag,
                    escape_choice = c("content", "script_or_style")) {
      
      escape_choice <- match.arg(escape_choice)
      
      new_function(
        exprs(... = ),
        expr({
          dots <- dots_partition(...)
          attribs <- html_attributes(dots$named)
          children <- map_chr(dots$unnamed,
                              # choose the escaping
                              ~ escape(., escape_choice = !!escape_choice))
          html(paste0(
            !!paste0("<", tag), attribs, ">",
            paste(children, collapse = ""),
            !!paste0("</", tag, ">")
            ))
          }),
      caller_env()
      )
    }
    ```
    
    Let's test our new `tag()` function:

    ```{r}
    p <- tag("p")
    b <- tag("b")

    identical(p("This &","this <content>", b("& this will be escaped")) %>%
                as.character(),
              "<p>This &amp;this &lt;content&gt;<b>&amp; this will be escaped</b></p>")

    script <- tag("script", escape_choice = "script_or_style")

    script("These signs will not be escaped: &, <, >, ",
           "but these ones will: </script> or </style>")
    ```

1.  __<span style="color:red">Q</span>__: The use of `...` for all functions has some big downsides. There's no
    input validation and there will be little information in the
    documentation or autocomplete about how they are used in the function. 
    Create a new function that, when given a named list of tags and their   
    attribute names (like below), creates functions which address this problem.

    ```{r, eval = FALSE}
    list(
      a = c("href"),
      img = c("src", "width", "height")
    )
    ```

    All tags should get `class` and `id` attributes.
    
    __<span style="color:green">A</span>__: 

1.  __<span style="color:red">Q</span>__: Currently the HTML doesn't look terribly pretty, and it's hard to see the
   structure. How could you adapt `tag()` to do indenting and formatting?
   
    __<span style="color:green">A</span>__: 

1.  __<span style="color:red">Q</span>__: Reason about the following code that calls `with_html()` referening objects
    from the environment. Will it work or fail? Why? Run the code to 
    verify your predictions.
    
    ```{r, eval = FALSE}
    greeting <- "Hello!"
    with_html(p(greeting))
    
    address <- "123 anywhere street"
    with_html(p(address))
    ```
    
    __<span style="color:green">A</span>__: 
    
## LaTeX

1.  __<span style="color:red">Q</span>__: Add escaping. The special symbols that should be escaped by adding a backslash
    in front of them are `\`, `$`, and `%`. Just as with HTML, you'll need to 
    make sure you don't end up double-escaping. So you'll need to create a small 
    S3 class and then use that in function operators. That will also allow you 
    to embed arbitrary LaTeX if needed.
    
    __<span style="color:green">A</span>__: 

1.  __<span style="color:red">Q</span>__: Complete the DSL to support all the functions that `plotmath` supports.

    __<span style="color:green">A</span>__: 
