# (PART) Miscellaneous {-} 

# Deprecated exercises

## Condidions

1. __<span style="color:red">Q</span>__: What does `options(error = recover)` do? Why might you use it?

   __<span style="color:green">A</span>__: In case of `options(error = recover)` `utils::recover()` will be called (without arguments) in case of an error. This will print out a list of calls which precede the error and lets the user choose to incorporate `browser()` directly in any of the regarding environments allowing a practical mode for debugging.

2. __<span style="color:red">Q</span>__: What does `options(error = quote(dump.frames(to.file = TRUE)))` do? Why might you use it?

   __<span style="color:green">A</span>__: This option writes a dump of the evaluation environment where an error occurs into a file ending on `.rda`. When this option is set, R will continue to run after the first error. To stop R at the first error use `quote({dump.frames(to.file=TRUE); q()})`. These options are especially useful for debugging non-interactive R scripts afterwards ("post mortem debugging").

## Expressions (new)

1.  __<span style="color:red">Q</span>__: `base::alist()` is useful for creating pairlists to be used for function arguments:
    
    ```{r}
    foo <- function() {}
    formals(foo) <- alist(x = , y = 1)
    foo
    ```
    
    What makes `alist()` special compared to `list()`?
    
    __<span style="color:green">A</span>__: From `?alist`:
    
    > alist handles its arguments as if they described function arguments. So the values are not evaluated, and tagged arguments with no value are allowed whereas list simply ignores them. alist is most often used in conjunction with formals.

## Quasiquotation (new)

1.  __<span style="color:red">Q</span>__: Why does `as.Date.default()` use `substitute()` and `deparse()`? Why does `pairwise.t.test()` use them? Read the source code.

    __<span style="color:green">A</span>__: `as.Date.default()` uses them to convert unexpected input expressions (neither dates, nor `NAs`) into a character string and return it within an error message.
    
    `pairwise.t.test()` uses them to convert the names of its datainputs (response vector `x` and grouping factor `g`) into character strings to format these further into a part of the desired output.

1.  __<span style="color:red">Q</span>__: `pairwise.t.test()` assumes that `deparse()` always returns a length one character vector. Can you construct an input that violates this expectation? What happens? 

    __<span style="color:green">A</span>__: We can pass an expression to one of `pairwise.t.test()`'s data input arguments, which exceeds the default cutoff width in `deparse()`. The expression will be split into a character vector of length greater 1. The deparsed data inputs are directly pasted (read the source code!) with "and" as separator and the result is just used to be displayed in the output. Just the data.name output will change (it will include more than one "and").
    
    ```{r}
    d=1
    pairwise.t.test(2, d+d+d+d+d+d+d+d+d+d+d+d+d+d+d+d+d)
    ```
