```{r, include=FALSE}
source("common.R")
```

# Rewriting R code in C++

## Getting started with C++

__[Q1]{.Q}__: With the basics of C++ in hand, it's now a great time to practice by reading and writing some simple C++ functions. For each of the following functions, read the code and figure out what the corresponding base R function is. You might not understand every part of the code yet, but you should be able to figure out the basics of what the function does.

```{cpp}
double f1(NumericVector x) {
  int n = x.size();
  double y = 0;
  
  for(int i = 0; i < n; ++i) {
    y += x[i] / n;
  }
  return y;
}

NumericVector f2(NumericVector x) {
  int n = x.size();
  NumericVector out(n);
  
  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
    out[i] = out[i - 1] + x[i];
  }
  return out;
}

bool f3(LogicalVector x) {
  int n = x.size();
  
  for(int i = 0; i < n; ++i) {
    if (x[i]) return true;
  }
  return false;
}

int f4(Function pred, List x) {
  int n = x.size();
  
  for(int i = 0; i < n; ++i) {
    LogicalVector res = pred(x[i]);
    if (res[0]) return i + 1;
  }
  return 0;
}

NumericVector f5(NumericVector x, NumericVector y) {
  int n = std::max(x.size(), y.size());
  NumericVector x1 = rep_len(x, n);
  NumericVector y1 = rep_len(y, n);
  
  NumericVector out(n);
  
  for (int i = 0; i < n; ++i) {
    out[i] = std::min(x1[i], y1[i]);
  }
  
  return out;
}
```

__[A]{.solved}__:  The code above corresponds to the following R functions:

*   f1: `mean()`  
*   f2: `cumsum()`  
*   f3: `any()`
*   f4: `Position()`
*   f5: `pmin()`

__[Q1]{.Q}__:  To practice your function writing skills, convert the following functions into C++. For now, assume the inputs have no missing values.

1. `all()`.

2. `cumprod()`, `cummin()`, `cummax()`.

3. `diff()`. Start by assuming lag 1, and then generalise for lag `n`.

4. `range()`.

5. `var()`. Read about the approaches you can take on [Wikipedia](http://en.wikipedia.org/wiki/Algorithms_for_calculating_variance). Whenever implementing a numerical algorithm, it's always good to check what is already known about the problem.

__[A]{.solved}__: Let's port these functions to C++.

1. `all()`

```{cpp}
bool allC(LogicalVector x) {
  int n = x.size();
  
  for(int i = 0; i < n; ++i) {
    if (!x[i]) return false;
  }
  return true;
}
```

2. `cumprod()`, `cummin()`, `cummax()`.

```{cpp}
NumericVector cumprodC(NumericVector x) {
  
  int n = x.size();
  NumericVector out(n);
  
  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
    out[i]  = out[i - 1] * x[i];
  }
  return out;
}

NumericVector cumminC(NumericVector x) {
  int n = x.size();
  NumericVector out(n);
  
  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
    out[i]  = std::min(out[i - 1], x[i]);
  }
  return out;
}

NumericVector cummaxC(NumericVector x) {
  int n = x.size();
  NumericVector out(n);
  
  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
    out[i]  = std::max(out[i - 1], x[i]);
  }
  return out;
}
```

3. `diff()` (Start by assuming lag 1, and then generalise for lag `n`.)

```{cpp}
NumericVector diffC(NumericVector x){
  int n = x.size();
  NumericVector out(n - 1);
  
  for(int i = 1; i < n; i++){
    out[i - 1] = x[i] - x[i - 1];
  }
  return out ;
}

NumericVector difflagC(NumericVector x, int lag = 1){
  int n = x.size();
  NumericVector out(n - lag);
  
  for(int i = lag; i < n; i++){
    out[i - lag] = x[i] - x[i - lag];
  }
  return out;
}
```

4. `range()`

```{cpp}
NumericVector rangeC(NumericVector x){
  double omin, omax;  
  int n = x.size();
  NumericVector out(2);
  
  omin = x[0];
  omax = x[0];
  
  for(int i = 1; i < n; i++){
    omin = std::min(x[i], omin);
    omax = std::max(x[i], omax);
  }
  
  out[0] = omin;
  out[1] = omax;
  return out;
```
  
5. `var()`
  
```{cpp}
double varC(NumericVector x) {
  int n = x.size();
  double mx = 0;
  double out = 0;
  
  if(n < 2) {
    return NA_REAL;
  }
  
  for(int i = 0; i < n; ++i) {
    mx += x[i] / n;
  }
  
  for(int i = 0; i < n; ++i) {
    out += pow(x[i] - mx, 2);
  }
  
  out = out / (n-1);
  return out;
}
```

## Missing values

__[Q1]{.Q}__:  Rewrite any of the functions from the first exercise to deal with missing values. If `na.rm` is true, ignore the missing values. If `na.rm` is false, return a missing value if the input contains any missing values. Some good functions to practice with are `min()`, `max()`, `range()`, `mean()`, and `var()`.

__[A]{.solved}__: We will refactor `cumsum()`, `any()`, `Position()` and `pmin()` so they can deal with missing values, but we practice and rewrite `min()`, `max()`, `range()`, `mean()` and `var()` first. We try to keep the overall function behaviour close to the original function, whenever `na_rm = false`. We mostly stick with vector data types as return values to avoid irregular type conversions.

We introduce an `na_rm` argument to make `minC()` aware of `NA`s. In case `x` contains exclusively `NA` values `minC()` should return `Inf` for `na_rm == TRUE`.

```{cpp}
NumericVector minC(NumericVector x, bool na_rm = false){
  int n = x.size();
  NumericVector out = NumericVector::create(R_PosInf);
  
  if (na_rm == false) {
    for(int i = 0; i < n; ++i) {
      if (NumericVector::is_na(x[i])) {
        out[0] = NA_REAL;
        return out;
      }
      if (x[i] < out[0]) {
        out[0] = x[i];
      }
    }
  }
  
  if (na_rm) {
    for(int i = 0; i < n; ++i) {
      if (x[i] == NA_REAL) {
        continue;
      }
      if (x[i] < out[0]) {
        out[0] = x[i];
      }
    }
  }
  
  return out;
}
```

To implement `maxC()` we reuse `minC()` and take advantage of a connection between the minimum and the maximum: $\max(x) = -\min(-x)$.

```{cpp}
NumericVector maxC(NumericVector x, bool na_rm = false){
  return -minC(-x, na_rm);
}
```

`minC()` and `maxC()` enable us to write a compact and `NA`-aware `rangeC()` function.

```{cpp}
NumericVector rangeC(NumericVector x, bool na_rm = false){
  NumericVector out(2);
  
  out[0] = minC(x, na_rm)[0];
  out[1] = maxC(x, na_rm)[0];
  
  return out;
}
```

Our `NA`-aware `meanC()` function should return `NaN`, if `na_rm = TRUE` and `all(is.na(x))`.

```{cpp}
NumericVector meanC(NumericVector x, bool na_rm = false){
  int n = x.size();
  int n_count = 0;
  NumericVector out = NumericVector::create(0);
  
  if (na_rm == false) {
    for(int i = 0; i < n; i++){
      if (NumericVector::is_na(x[i])) {
        out[0] = NA_REAL;
        return out;
      }
      out[0] += x[i];
      n_count++;
    }
    out[0] /= n_count;
  }
  
  if (na_rm) {
    for(int i = 0; i < n; i++){
      if (NumericVector::is_na(x[i])) {
        continue;
      }
      out[0] += x[i];
      n_count++;
    }
    if (n_count == 0) {
      out[0] = NAN;
      return out;
    }
    out[0] /= n_count;
  }
  
  return out;
}
```

For `varC()`, we handle both cases of `na_rm` inside the first for loop, as this reduces code duplication.

```{cpp}
NumericVector varC(NumericVector x, bool na_rm = false) {
  int n = x.size();
  int n_count = 0;
  double m_x = 0;
  NumericVector out = NumericVector::create(0);
  
  for(int i = 0; i < n; i++) {
    if (NumericVector::is_na(x[i])) {
      if (na_rm == FALSE) {
        out[0] = NA_REAL;
        return out;
      } else {
        continue;
      }
    }
    m_x += x[i];
    n_count++;
  }
  
  if (n_count < 2) {
    out[0] = NA_REAL;
    return out;
  }
  
  m_x /= n_count;
  
  for(int i = 0; i < n; ++i) {
    if (NumericVector::is_na(x[i])) {
      continue;
    }
    out[0] += pow(x[i] - m_x, 2);
  }
  
  out[0] /= n_count - 1;
  
  return out;
}
```

Now, let's extend the functions `cumsum()`, `any()`, `Position()` and `pmin()` from the (first exercise).

For `na_rm = true`, we keep the `NA`'s in the output but ignore them in the cumulative sums.

```{cpp}
NumericVector cumsumC(NumericVector x, bool na_rm = false) {
  int n = x.size();
  NumericVector out(n);
  double sum_i = 0;
  
  for(int i = 0; i < n; ++i) {
    if (NumericVector::is_na(x[i])) {
      if (na_rm == false) {
        return NumericVector::create(NA_REAL);
      }
      out[i] = NA_REAL;
      continue;
    }
    sum_i += x[i];
    out[i] = sum_i;
  }
  
  return out;
}
```

In our new implementation of `anyC()` we use `LogicalVetor` as return type. If we would use `bool` instead, the C++ `NA_LOGICAL` would be converted into R's logical `TRUE`.

```{cpp}
LogicalVector anyC(LogicalVector x, bool na_rm = false) {
  int n = x.size();
  LogicalVector out = LogicalVector::create(false);
  
  if (na_rm == false) {
    for(int i = 0; i < n; ++i) {
      if (LogicalVector::is_na(x[i])) {
        out[0] = NA_LOGICAL;
        return out;
      } else {
        if (x[i]) {
          out[0] = true;
        }
      }
    }
  }
  
  if (na_rm) {
    for(int i = 0; i < n; ++i) {
      if (LogicalVector::is_na(x[i])) {
        continue;
      }
      if (x[i]) {
        out[0] = true;
        return out;
      }
    }
  }
  return out;
}
```

For `PositionC()` we check the results of the predicate function for `NA`s. In some cases it may also make sense check the elements of the list input for `NA`s, and provide an `NA`-handling for the predicate function.

```{cpp}
int PositionC(Function pred, List x, bool na_rm = false) {
  int n = x.size();
  
  for(int i = 0; i < n; ++i) {
    LogicalVector res = pred(x[i]);
    
    if (LogicalVector::is_na(res[0])) {
      if (na_rm) {continue;}
      return NA_INTEGER;
    }
    if (res[0]) return i + 1;
  }
  return 0;
}
```

When we set `na_rm = TRUE` in our `pminC()` function it only returns `NA`s at indices where both, `x` and `y`, contain an `NA`.

```{cpp}
NumericVector pminC(NumericVector x, NumericVector y,
                    bool na_rm = false) {
  int n = std::max(x.size(), y.size());
  NumericVector x1 = rep_len(x, n);
  NumericVector y1 = rep_len(y, n);
  
  NumericVector out(n);
  
  for (int i = 0; i < n; ++i) {
    if (na_rm == false) {
      if (NumericVector::is_na(x[i])) {
        return NumericVector::create(NA_REAL);
      }
      if (NumericVector::is_na(y[i])) {
        return NumericVector::create(NA_REAL);
      }
    }
    if (na_rm == true) {
      if (NumericVector::is_na(x[i])) {
        out[i] = y[i];
        continue;
      }
      if (NumericVector::is_na(y[i])) {
        out[i] = x[i];
        continue;
      }
    }
    out[i] = std::min(x1[i], y1[i]);
  }
  
  return out;
}
```

__[Q2]{.Q}__: Rewrite `cumsum()` and `diff()` so they can handle missing values. Note that these functions have slightly more complicated behaviour.

__[A]{.solved}__: As we already wrote an `NA`-aware `cumsumC()` function under the assumption to return a single `NA` in case of `na_rm = false`, we modify it here slightly to always return a vector with the same length as it's `x` argument. Our new `cumsumC2()` function treats `NA` values always like zeros. In case of `na_rm = false` the `NA` values are kept in the output and in case of `na_rm = true` the `NA` values are replaced by the last occurring non-`NA`-value or zero (if it's entirely `NA` values).

```{cpp}
NumericVector cumsumC(NumericVector x) {
  int n = x.size();
  NumericVector out(n);
  LogicalVector index = is_na(x);
  
  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
    if (index[i - 1]) {
      out[i] = NA_REAL;
    } else{
      out[i] = out[i - 1] + x[i];
    }
  }
  
  return out;
}
```

For `diffC()`'s implementation, we again return just an `NA` whenever an `NA` value occurs. In case of `na_rm = true`, we ensure that calculations which are affected by `NA`'s will also return an `NA`. (We could have also chosen to exclude `NA`s completely from the input or - equivalently - the output).

```{cpp}
NumericVector difflagC(NumericVector x, int lag = 1,
                       bool na_rm = false){
  int n = x.size();
  NumericVector out(n - lag);
  for(int i = lag; i < n; i++){
    if (NumericVector::is_na(x[i]) ||
        NumericVector::is_na(x[i - lag])) {
      if (na_rm == false) {
        return NumericVector::create(NA_REAL);
      }
      out[i - lag] = NA_REAL;
      continue;
    }
    out[i - lag] = x[i] - x[i - lag];
  }
  return out;
}
```

## Standard Template Library

To practice using the STL algorithms and data structures, implement the following using R functions in C++, using the hints provided:

__[Q1]{.Q}__: `median.default()` using `partial_sort`.

__[A]{.solved}__: The median is computed differently for even or odd vectors, which we allow for in the function below.

```{cpp}
#include <algorithm>
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double medianC(NumericVector x) {
  int n = x.size();
  double out;
  if (n % 2 == 0){
    std::partial_sort (x.begin(), x.begin() + n / 2 + 1, x.end());
    out = (x[n / 2 - 1] + x[n / 2]) / 2;
  } else {
    std::partial_sort (x.begin(), x.begin() + (n + 1) / 2, x.end());
    out = x[(n + 1) / 2 - 1];
  }
  
  return out;
}
```

__[Q2]{.Q}__: `%in%` using `unordered_set` and the `find()` or `count()` methods.

__[A]{.solved}__: We use the `find()` method and loop through the `unordered_set` until we find a match or have scanned the entire set.

```{cpp}
// [[Rcpp::plugins(cpp11)]]
#include <Rcpp.h>
#include <unordered_set>
using namespace Rcpp;

// [[Rcpp::export]]
LogicalVector inC(CharacterVector x, CharacterVector table) {
  std::unordered_set<String> seen;
  int n_x = x.size();
  std::vector<bool> out;
  
  seen.insert (table.begin(), table.end());
  
  for (int i = 0; i < n_x; ++i) {
    if (seen.find(x[i]) == seen.end()) {
      out.push_back(false);
      continue;
    } 
    out.push_back(true);
  }
  
  return wrap(out);
}
```

__[Q3]{.Q}__: `unique()` using an `unordered_set` (challenge: do it in one line!).

__[A]{.solved}__: The `insert()`-method will return if an equivalent element already existed. If a new element is inserted, we will add it to the (unique) return vector of our function.

```{Rcpp}
// [[Rcpp::plugins(cpp11)]]
#include <Rcpp.h>
#include <unordered_set>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector uniqueC(NumericVector x) {
  std::unordered_set<int> seen;
  int n = x.size();
  std::vector<double> out;
  
  for (int i = 0; i < n; ++i) {
    if (seen.insert(x[i]).second) out.push_back(x[i]);
  }
  
  return wrap(out);
}


// As a one-liner
// [[Rcpp::export]]
NumericVector uniqueCC(NumericVector x) {
  return wrap(
    *(new std::unordered_set<double>(x.begin(), x.end()))
  );
}
```


__[Q4]{.Q}__: `min()` using `std::min()`, or `max()` using `std::max()`.

__[A]{.solved}__: We will implement `min()` by iterating over the vector and recursively comparing each element to the current minimum value.

```{cpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double minC(NumericVector x){
  int n = x.size();
  double out = x[0];
  
  for (int i = 0; i < n; i++){
    out = std::min(out, x[i]);
  }
  
  return out;
}
```

__[Q5]{.Q}__: `which.min()` using `min_element`, or `which.max()` using `max_element`.

__[A]{.solved}__: To implement `which.min()`, we will first locate the `min_element()` and then compute the `distance()` to it (from the beginning of the vector).

```{cpp}
#include <Rcpp.h>
#include <algorithm>
#include <iterator>

using namespace Rcpp;

// [[Rcpp::export]]
double which_minC(NumericVector x){
  int out;
  out = std::distance(
    x.begin(), std::min_element(x.begin(), x.end())
  );
  out++;
  
  return out;
}
```

__[Q6]{.Q}__: `setdiff()`, `union()`, and `intersect()` for integers using sorted ranges and `set_union`, `set_intersection` and `set_difference`.

__[A]{.solved}__: The structure of the three functions will be very similar.

We first sort both input vectors. Then we apply the respective `set_union`, `set_intersection` or `set_difference` function. After that, the result will be between the iterators `tmp.begin()` and `out_end`. To retrieve the result we loop once through the range between `tmp.begin()` and `out_end` in the last part of each function.

One note on the naming of the functions: `union` is a special class in cpp so we may not use it for our function definition and rather use `setunion`.

```{Rcpp}
#include <Rcpp.h>
#include <unordered_set>
#include <algorithm>
using namespace Rcpp;

// [[Rcpp::plugins(cpp11)]]
// [[Rcpp::export]]
IntegerVector setunion(IntegerVector x, IntegerVector y) {
  int nx = x.size();
  int ny = y.size();
  
  IntegerVector tmp(nx + ny);
  
  std::sort(x.begin(), x.end());
  std::sort(y.begin(), y.end());
  
  IntegerVector::iterator out_end = std::set_union(
    x.begin(), x.end(), y.begin(), y.end(), tmp.begin()
  );
  
  IntegerVector out;
  for(IntegerVector::iterator it = tmp.begin();
      it != out_end; ++it) {
    out.push_back(*it);
  }
  
  return out;
}

// [[Rcpp::export]]
IntegerVector intersect(IntegerVector x, IntegerVector y) {
  int nx = x.size();
  int ny = y.size();
  
  IntegerVector tmp(std::min(nx, ny));
  
  std::sort(x.begin(), x.end());
  std::sort(y.begin(), y.end());
  
  IntegerVector::iterator out_end = std::set_intersection(
    x.begin(), x.end(), y.begin(), y.end(), tmp.begin()
  );
  
  IntegerVector out;
  for(IntegerVector::iterator it = tmp.begin();
      it != out_end; ++it) {
    out.push_back(*it);
  }
  
  return out;
}

// [[Rcpp::export]]
IntegerVector setdiff(IntegerVector x, IntegerVector y) {
  int nx = x.size();
  int ny = y.size();
  
  IntegerVector tmp(nx);
  
  std::sort(x.begin(), x.end());
  std::sort(y.begin(), y.end());
  
  IntegerVector::iterator out_end = std::set_difference(
    x.begin(), x.end(), y.begin(), y.end(), tmp.begin()
  );
  
  IntegerVector out;
  for(IntegerVector::iterator it = tmp.begin();
      it != out_end; ++it) {
    out.push_back(*it);
  }
  
  return out;
}
```

Let's verify, that these functions work as intended.

```{r}
x <- c(1, 2, 3, 3)
y <- c(3, 2, 5)

setunion(x, y)
intersect(x, y)
setdiff(x, y)

typeof(setdiff(x, y))
```
