---
title: "Cover Plot Bars Viridis"
author: "Tobias Stalder"
date: "17 12 2020"
output: html_document
---

```{r setup, include=FALSE}
```

## 1 Document Purpose
This document shows the code for creating the final cover plot for the new book of Henning Bumann and Malte Grosser "Advanced R Solutions".

**Load libraries and set working dir (and install with dependencies if needed)**
```{r, message = FALSE, warning = FALSE}

packages <- c("tidyverse", "Cairo", "ggpubr", "ggalt", "ggbump", "scales", "prismatic", "treemapify","GGally",
              "intergraph", "ggraph", "visNetwork", "threejs", "networkD3", "ndtv", "here", "igraph")

package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)

here()

```

**Load data from local dir and direct cleaning to long format**

The following snippet was copied from the book repo (originally written by Henning Bumann)
```{r}
df_long <- read.csv2(r"(progress_data.csv)", stringsAsFactors = FALSE, 
                     header = TRUE) %>% 
  as_tibble() %>% 
  mutate(status = "finalised") %>% 
  modify_at("status", factor,
            levels = c("open", "started", "solved", "reviewed", "finalised")) %>% 
  tibble::rowid_to_column("id") %>% 
  separate(chapter, "nr", sep = " ", remove = FALSE, extra = "drop") %>% 
  mutate(nr = as.integer(nr), 
         chapter = factor(chapter),
         chapter = fct_reorder(chapter, nr, .desc = FALSE),
         part = case_when(nr < 9 ~ "Foundations",
                          nr < 12 ~ "Functional\nProgramming",
                          nr < 17 ~ "Object Oriented\nProgramming",
                          nr < 22 ~ "Metaprogramming",
                          TRUE    ~ "Techniques"),
         part = factor(part, levels = c("Foundations", "Functional\nProgramming",
                                        "Object Oriented\nProgramming",
                                        "Metaprogramming", "Techniques"))) %>% 
  add_count(chapter)
df_long$subchapter = factor(df_long$subchapter)
head(df_long)
```



## 2 Requirement Catalogue
The vizualisation for a book cover has special requirements since it transports the very first message of the book. In alignment with the authors, I defined some requirements which shall guide through the creation process.

1: Viz should show some kind of "structure" in the design since it's plotted out of very well structured data of the exercises.

2: I'd like the viz to transport "structure" and "progress" as a message.

3: The viz should  invite the reader to read and engage, not push the reader away by transporting a message of complexity.

5: The viz should not be too creative since the book is about advanced R programming and not information design.

6: Potential readers should be able to relate with the viz in a sense of  ("I could do that", "This speaks to my abilities")


**1. Bar chart**
```{r, echo=TRUE, message=FALSE}
df_long %>% 
  group_by(chapter) %>% 
  mutate(exercise_in_chapter_id = row_number()) %>% 
  ungroup() %>% 
  ggplot(aes(x = chapter, fill = part, group = -id)) +
  geom_bar(color = "#2a0829", alpha = 0.8, aes(fill = part)) +
  # scale_y_continuous(expand = c(.05, 0)) +
  # scale_x_discrete(expand = c(.2,0)) +

  scale_fill_viridis_d(begin = .1, end = .8)+
  facet_grid(. ~ part, scales = "free",
             space = "free", switch = "x")+
  theme_void()+
  theme(legend.position = "none",
        plot.background = element_rect(fill = "transparent", color = "transparent"),
        panel.background = element_rect(fill = "transparent", color = "transparent"),
        strip.background = element_rect(fill = "transparent", color = "transparent"),
        strip.text = element_blank()) +
  
  coord_cartesian(ylim = c(0, 35)) -> bar_chart_viridis

bar_chart_viridis

ggsave(plot = bar_chart_viridis, filename = paste0(here(),"/bars_viridis.png"), units = "cm", width = 16, height = 8, dpi = 600, type = "cairo-png", bg = "transparent")
# ggsave(plot = bar_chart_viridis, filename = "plots/bars_viridis.svg", units = "cm", width = 16, height = 8, dpi = 600, bg = "transparent")


```




